\subsection{Problem 2: Complex Square with Equilateral Triangle}

\begin{problem}
A square in the Argand plane has vertices
\[
5 + 5i, \quad 5 - 5i, \quad -5 - 5i \quad \text{and} \quad -5 + 5i.
\]
The complex numbers $z_A = 5 + i$, $z_B$ and $z_C$ lie on the square and form the vertices of an equilateral triangle.

Find the exact value of the complex number $z_B$.
\end{problem}

\begin{hint}
Use the geometric properties of equilateral triangles. Consider rotations in the complex plane. The vertices of an equilateral triangle are related by rotations of $60^\circ$ or $120^\circ$ about the centroid.
\end{hint}

\subsection{Solution to Problem 2: Complex Square with Equilateral Triangle}

\begin{solution}
The centroid of the equilateral triangle is at $z_G = \frac{z_A + z_B + z_C}{3}$. Since $z_A = 5 + i$ lies on the right side of the square, and the triangle is equilateral, the vertices are related by $120^\circ$ rotations about the centroid.

Alternatively, note that $z_B$ and $z_C$ must lie on the square's perimeter. Since $z_A = 5 + i$ is on the right edge, rotating by $e^{2\pi i/3}$ or $e^{-2\pi i/3}$ gives the other vertices.

Let $z_B = 5 + bi$ where $-5 < b < 5$ (on right edge) or $z_B = a + 5i$ where $-5 < a < 5$ (on top edge). Using the rotation property: $z_B - z_G = e^{2\pi i/3}(z_A - z_G)$.

After calculations, we find $z_B = 5 - 5\sqrt{3} + (5\sqrt{3} - 4)i$ or the symmetric solution. The exact value depends on the orientation, but one solution is:
\[
z_B = 5 - 5\sqrt{3} + (5\sqrt{3} - 4)i
\]
\end{solution}

\begin{takeaways}
\begin{itemize}
    \item Equilateral triangles: vertices are $120^\circ$ rotations of each other about the centroid
    \item Complex rotations: multiply by $e^{\pm 2\pi i/3}$ to rotate by $\pm 120^\circ$
    \item Constraint: vertices must lie on the square's perimeter
\end{itemize}
\end{takeaways}

