\subsection{Problem 12: Bar Magnet and Falling Object}

\begin{problem}
A bar magnet is held vertically. An object that is repelled by the magnet is to be dropped from directly above the magnet and will maintain a vertical trajectory.

Let $x$ be the distance of the object above the magnet.

The object is subject to acceleration due to gravity, $g$, and an acceleration due to the magnet so that the total acceleration of the object is given by
\[
a = \frac{27g}{x^3} - g.
\]
The object is released from rest at $x=6$.

\begin{enumerate}
    \item[(i)] Show that $v^2 = g\left(\frac{51}{4} - 2x - \frac{27}{x^2}\right)$.
    
    \item[(ii)] Find where the object next comes to rest, giving your answer correct to 1 decimal place.
\end{enumerate}
\end{problem}

\begin{hint}
For part (i), use $a = v \frac{dv}{dx}$ and integrate. For part (ii), set $v = 0$ and solve the resulting equation. You may need to use numerical methods or factor the polynomial.
\end{hint}

\subsection{Solution to Problem 12: Bar Magnet and Falling Object}

\begin{solution}
\textbf{(i)} Using $a = v \frac{dv}{dx}$:
\[
v \frac{dv}{dx} = \frac{27g}{x^3} - g = g\left(\frac{27}{x^3} - 1\right)
\]
Separating variables and integrating:
\[
\int_0^v v \, dv = g \int_6^x \left(\frac{27}{x^3} - 1\right) dx
\]
\[
\frac{v^2}{2} = g\left[-\frac{27}{2x^2} - x\right]_6^x = g\left(-\frac{27}{2x^2} - x + \frac{27}{2 \cdot 36} + 6\right)
\]
\[
= g\left(-\frac{27}{2x^2} - x + \frac{3}{8} + 6\right) = g\left(\frac{51}{8} - x - \frac{27}{2x^2}\right)
\]
Therefore: $v^2 = g\left(\frac{51}{4} - 2x - \frac{27}{x^2}\right)$.

\textbf{(ii)} When the object comes to rest, $v = 0$:
\[
\frac{51}{4} - 2x - \frac{27}{x^2} = 0
\]
Multiplying by $4x^2$: $51x^2 - 8x^3 - 108 = 0$, or $8x^3 - 51x^2 + 108 = 0$.

We need to find the root where $x < 6$ (since object falls downward). Trying values:
- $x = 1.5$: $8(3.375) - 51(2.25) + 108 = 27 - 114.75 + 108 = 20.25$
- $x = 2$: $8(8) - 51(4) + 108 = 64 - 204 + 108 = -32$
- $x = 1.7$: $8(4.913) - 51(2.89) + 108 = 39.304 - 147.39 + 108 = -0.086$

By intermediate value theorem and refinement, $x \approx 1.7$ m (to 1 decimal place).
\end{solution}

\begin{takeaways}
\begin{itemize}
    \item Use $a = v \frac{dv}{dx}$ for position-dependent acceleration
    \item Integrate to find velocity as function of position
    \item Solve $v = 0$ to find rest positions
\end{itemize}
\end{takeaways}

