\subsection{Problem 16: Projectile with Quadratic Resistance}

\begin{problem}
A projectile of mass $M$ kg is launched vertically upwards from the origin with an initial speed $v_{0}$ m s$^{-1}$. The acceleration due to gravity is $g$ m s$^{-2}$.

The projectile experiences a resistive force of magnitude $kMv^{2}$ newtons, where $k$ is a positive constant and $v$ is the speed of the projectile at time $t$ seconds.

\begin{enumerate}
    \item[(i)] The maximum height reached by the particle is $H$ metres. Show that
    \[
    H=\frac{1}{2k}\ln\left(\frac{kv_{0}^{2}+g}{g}\right).
    \]
    
    \item[(ii)] When the projectile lands on the ground, its speed is $v_{1}$ m s$^{-1}$, where $v_{1}$ is less than the magnitude of the terminal velocity. Show that $g(v_{0}^{2}-v_{1}^{2})=kv_{0}^{2}v_{1}^{2}$.
\end{enumerate}
\end{problem}

\begin{hint}
For part (i), use $a = v \frac{dv}{dx}$ and integrate. The resistive force opposes motion, so the acceleration is $-g - kv^2$ on the way up. For part (ii), consider the energy or use the same technique for the downward motion.
\end{hint}

\subsection{Solution to Problem 16: Projectile with Quadratic Resistance}

\begin{solution}
\textbf{(i)} On the upward journey, the acceleration is $a = -g - kv^2$ (negative because both gravity and resistance oppose motion).

Using $a = v \frac{dv}{dx}$:
\[
v \frac{dv}{dx} = -g - kv^2
\]
Separating variables:
\[
\frac{v}{g + kv^2} \, dv = -dx
\]
Integrating from $v = v_0$ at $x = 0$ to $v = 0$ at $x = H$:
\[
\int_{v_0}^0 \frac{v}{g + kv^2} \, dv = -\int_0^H dx
\]
For the left integral, let $u = g + kv^2$, so $du = 2kv \, dv$, giving:
\[
\int_{v_0}^0 \frac{v}{g + kv^2} \, dv = \frac{1}{2k} \int_{g + kv_0^2}^g \frac{du}{u} = \frac{1}{2k} \ln\left(\frac{g}{g + kv_0^2}\right) = -\frac{1}{2k} \ln\left(\frac{g + kv_0^2}{g}\right)
\]
Therefore: $-H = -\frac{1}{2k} \ln\left(\frac{g + kv_0^2}{g}\right)$, so
\[
H = \frac{1}{2k}\ln\left(\frac{kv_0^2 + g}{g}\right).
\]

\textbf{(ii)} On the downward journey, acceleration is $a = g - kv^2$ (gravity assists, resistance opposes).

Using the same technique: $v \frac{dv}{dx} = g - kv^2$.

Integrating from $v = 0$ at $x = H$ to $v = v_1$ at $x = 0$:
\[
\int_0^{v_1} \frac{v}{g - kv^2} \, dv = \int_H^0 dx = -H
\]
Let $u = g - kv^2$, so $du = -2kv \, dv$:
\[
\int_0^{v_1} \frac{v}{g - kv^2} \, dv = -\frac{1}{2k} \int_g^{g - kv_1^2} \frac{du}{u} = -\frac{1}{2k} \ln\left(\frac{g - kv_1^2}{g}\right) = \frac{1}{2k} \ln\left(\frac{g}{g - kv_1^2}\right)
\]
So: $\frac{1}{2k} \ln\left(\frac{g}{g - kv_1^2}\right) = -H = -\frac{1}{2k}\ln\left(\frac{kv_0^2 + g}{g}\right)$.

Therefore: $\ln\left(\frac{g}{g - kv_1^2}\right) = -\ln\left(\frac{kv_0^2 + g}{g}\right)$, so $\frac{g}{g - kv_1^2} = \frac{g}{kv_0^2 + g}$.

Cross-multiplying: $(g - kv_1^2)(kv_0^2 + g) = g^2$.

Expanding: $g(kv_0^2 + g) - kv_1^2(kv_0^2 + g) = g^2$, so $gkv_0^2 + g^2 - kv_0^2v_1^2 - kgv_1^2 = g^2$.

Simplifying: $gkv_0^2 - kv_0^2v_1^2 - kgv_1^2 = 0$, so $gk(v_0^2 - v_1^2) = kv_0^2v_1^2$.

Therefore: $g(v_0^2 - v_1^2) = kv_0^2 v_1^2$.
\end{solution}

\begin{takeaways}
\begin{itemize}
    \item Quadratic resistance: Use $a = v \frac{dv}{dx}$ for position-dependent acceleration
    \item Integration: Use substitution $u = g \pm kv^2$ to integrate $\frac{v}{g \pm kv^2}$
    \item Energy approach: Can also use work-energy theorem for part (ii)
\end{itemize}
\end{takeaways}

