\begin{problem}[Binomial Theorem]
Prove by inductionthat 

$$(a+b)^n = \sum_{k=0}^{n} \binom{n}{k} a^{n-k} b^k$$ 
for all integers $n \ge 0$ and real numbers $a,b$.

\end{problem}

\begin{hint}
Use Pascal's identity $\binom{n+1}{k} = \binom{n}{k-1} + \binom{n}{k}$ to expand $(a+b)^{n+1} = (a+b)(a+b)^n$.
\end{hint}

\begin{solution}[Sketch]
Base case: $(a+b)^0 = 1 = \binom{0}{0}a^0b^0$. Assume the formula holds for $n$. Then $(a+b)^{n+1} = (a+b)\sum_{k=0}^{n}\binom{n}{k}a^{n-k}b^k = \sum_{k=0}^{n}\binom{n}{k}a^{n+1-k}b^k + \sum_{k=0}^{n}\binom{n}{k}a^{n-k}b^{k+1}$. Reindex and apply Pascal's identity to combine terms.
\end{solution}

\begin{problem}[Fermat's Little Theorem]
For any prime $p$ and integer $a$, prove that $a^p \equiv a \pmod{p}$.
\end{problem}

\begin{hint}
Use induction on $a$. Show $(k+1)^p - (k+1) \equiv k^p - k \pmod{p}$ by expanding with the binomial theorem and noting that $\binom{p}{j}$ is divisible by $p$ for $1 \le j \le p-1$.
\end{hint}

\begin{solution}[Sketch]
Base case: $0^p \equiv 0 \pmod{p}$. Assume $k^p \equiv k \pmod{p}$. Then $(k+1)^p = k^p + \binom{p}{1}k^{p-1} + \cdots + \binom{p}{p-1}k + 1$. Since $p \mid \binom{p}{j}$ for $1 \le j \le p-1$, we have $(k+1)^p \equiv k^p + 1 \equiv k + 1 \pmod{p}$ by the hypothesis.
\end{solution}

% --- New Problem: Symmetric Sum Inequality ---
\begin{problem}[Symmetric Sum Inequality]
Let $n$ be a positive integer and $a_1, a_2, \dots, a_n$ be $n$ positive real numbers. Prove by mathematical induction that:
\[
\left( a_1 + a_2 + \dots + a_n \right) \left( \frac{1}{a_1} + \frac{1}{a_2} + \dots + \frac{1}{a_n} \right) \geq n^2
\]
for all $n \geq 1$.
\end{problem}

\begin{hint}
Use induction on $n$. For the inductive step, expand $\left( \sum_{i=1}^{k+1} a_i \right) \left( \sum_{i=1}^{k+1} \frac{1}{a_i} \right)$ and apply the inequality $x + \frac{1}{x} \geq 2$ for $x > 0$ to each pair $\frac{a_i}{a_{k+1}}$ and $\frac{a_{k+1}}{a_i}$.
\end{hint}

\begin{solution}[Sketch]
Base case: $n=1$. $(a_1)\left(\frac{1}{a_1}\right) = 1 \geq 1^2$.

Inductive hypothesis: Assume true for $n=k$, i.e., $\left( \sum_{i=1}^k a_i \right) \left( \sum_{i=1}^k \frac{1}{a_i} \right) \geq k^2$.

Inductive step: For $n=k+1$,
\begin{align*}
\left( \sum_{i=1}^{k+1} a_i \right) \left( \sum_{i=1}^{k+1} \frac{1}{a_i} \right) &= \left( S_k + a_{k+1} \right) \left( R_k + \frac{1}{a_{k+1}} \right) \\
&= S_k R_k + S_k \cdot \frac{1}{a_{k+1}} + a_{k+1} \cdot R_k + 1 \\
&= S_k R_k + 1 + \sum_{i=1}^k \left( \frac{a_i}{a_{k+1}} + \frac{a_{k+1}}{a_i} \right)
\end{align*}
By the hypothesis, $S_k R_k \geq k^2$. For each $i$, $\frac{a_i}{a_{k+1}} + \frac{a_{k+1}}{a_i} \geq 2$. Thus,
\begin{align*}
	\text{LHS}_{k+1} &\geq k^2 + 1 + 2k = (k+1)^2
\end{align*}
So the result holds for $n=k+1$.

Conclusion: By induction, the inequality holds for all $n \geq 1$.
\end{solution}

\begin{problem}[Tower of Hanoi]
The Tower of Hanoi is a classic puzzle involving three pegs and $n$ disks of different sizes, all initially stacked in order of decreasing size on one peg. The objective is to move the entire stack to another peg, moving only one disk at a time and never placing a larger disk on top of a smaller one.
Prove that the minimum number of moves required to transfer $n$ disks in the Tower of Hanoi puzzle is $2^n - 1$.
\end{problem}

\begin{hint}
To move $k+1$ disks, first move the top $k$ disks to the auxiliary peg (requiring $T_k$ moves), then move the largest disk (1 move), then move the $k$ disks to the destination (requiring $T_k$ moves again). This gives $T_{k+1} = 2T_k + 1$.
\end{hint}

\begin{solution}[Sketch]
Base case: $T_1 = 1 = 2^1 - 1$. Assume $T_k = 2^k - 1$. To move $k+1$ disks: move $k$ disks to auxiliary peg ($2^k-1$ moves), move largest disk (1 move), move $k$ disks to destination ($2^k-1$ moves). Total: $T_{k+1} = 2(2^k-1) + 1 = 2^{k+1} - 1$.
\end{solution}

\begin{problem}[Vandermonde's Identity]
Prove that $\binom{m+n}{r} = \sum_{k=0}^{r} \binom{m}{k}\binom{n}{r-k}$ 
for non-negative integers $m,n,r$.

Here, by definition, $\binom{a}{b} = \frac{a!}{b!(a-b)!}$ for integers $0 \le b \le a$, and

$\binom{a}{b} = 0$ if $b > a$. 

\end{problem}

\begin{hint}
Use induction on $n$. Apply Pascal's identity $\binom{n+1}{j} = \binom{n}{j-1} + \binom{n}{j}$ to split the right-hand side.
\end{hint}

\begin{solution}[Sketch]
Induct on $n$ with $m,r$ fixed. Base case $n=0$: both sides equal $\binom{m}{r}$. Assume the identity holds for $n$. Then $\binom{m+n+1}{r} = \binom{m+n}{r-1} + \binom{m+n}{r}$. Apply the hypothesis to both terms and rearrange using $\binom{n}{j-1} + \binom{n}{j} = \binom{n+1}{j}$ to obtain the desired sum.
\end{solution}

\begin{problem}[Wilson's Theorem]
Prove that $(p-1)! \equiv -1 \pmod{p}$ for every prime $p$.
\end{problem}

\begin{hint}
For $p>2$, pair each $a \in \{2,3,\ldots,p-2\}$ with its inverse $a^{-1} \pmod{p}$. The only elements that are their own inverses are $1$ and $p-1$.
\end{hint}

\begin{solution}[Sketch]
Base case: $p=2$ gives $(2-1)! = 1 \equiv -1 \pmod{2}$. For prime $p>2$, in the product $(p-1)! = 1 \cdot 2 \cdots (p-1)$, each $a \in \{2,\ldots,p-2\}$ pairs with its inverse $a^{-1}$ (where $a \neq a^{-1}$), contributing $aa^{-1} \equiv 1$. Only $1$ and $p-1$ are self-inverse, so $(p-1)! \equiv 1 \cdot (p-1) \equiv -1 \pmod{p}$.
\end{solution}

\begin{problem}[Recursive Sequence with Surds]
A sequence $a_n$ is defined by $a_n = 2a_{n-1} + a_{n-2}$ for $n \ge 2$ with $a_0 = a_1 = 2$. Use mathematical induction to prove that:
\[a_n = (1+\sqrt{2})^n + (1-\sqrt{2})^n \quad \text{for all } n \ge 0\]
\end{problem}

\begin{hint}
Use strong induction. For the base cases verify $n=0$ and $n=1$. Then show $(1+\sqrt{2})^{k+1} + (1-\sqrt{2})^{k+1} = 2[(1+\sqrt{2})^k + (1-\sqrt{2})^k] + [(1+\sqrt{2})^{k-1} + (1-\sqrt{2})^{k-1}]$ using the fact that $(1 \pm \sqrt{2})$ satisfy $x^2 = 2x + 1$.
\end{hint}

\begin{solution}[Sketch]
Base cases: $a_0 = 2 = 1 + 1$, $a_1 = 2 = (1+\sqrt{2}) + (1-\sqrt{2})$. Assume the formula holds for $n=k-1$ and $n=k$. Note that $\alpha = 1+\sqrt{2}$ and $\beta = 1-\sqrt{2}$ both satisfy $x^2 = 2x + 1$. Then $a_{k+1} = 2a_k + a_{k-1} = 2(\alpha^k + \beta^k) + (\alpha^{k-1} + \beta^{k-1}) = \alpha^{k-1}(2\alpha + 1) + \beta^{k-1}(2\beta + 1) = \alpha^{k-1} \cdot \alpha^2 + \beta^{k-1} \cdot \beta^2 = \alpha^{k+1} + \beta^{k+1}$.
\end{solution}

\begin{problem}[Powers of $2+\sqrt{3}$]
For each positive integer $n$, prove that there exist unique positive integers $p_n$ and $q_n$ such that
\[(2 + \sqrt{3})^n = p_n + q_n \sqrt{3} .\]
Then show that these numbers satisfy the identity $p_n^2 - 3q_n^2 = 1$ for all $n$.
\end{problem}

\begin{hint}
Induct on $n$. For the step $k \to k+1$, expand $(2+\sqrt{3})^{k+1} = (2+\sqrt{3})^k(2+\sqrt{3})$, substitute the hypothesis $(2+\sqrt{3})^k = p_k + q_k\sqrt{3}$, and collect rational and irrational parts to define $p_{k+1}$ and $q_{k+1}$. For the identity, consider the conjugate $(2-\sqrt{3})^n = p_n - q_n\sqrt{3}$ and multiply the two expressions.
\end{hint}

\begin{solution}[Sketch]
Base case $n=1$: $(2+\sqrt{3}) = 2 + 1\cdot\sqrt{3}$, so $p_1=2$, $q_1=1$. Inductive step: assume $(2+\sqrt{3})^k = p_k + q_k\sqrt{3}$. Then
\begin{align*}
(2+\sqrt{3})^{k+1} &= (p_k + q_k\sqrt{3})(2+\sqrt{3}) \\
&= (2p_k + 3q_k) + (p_k + 2q_k)\sqrt{3} .
\end{align*}
Define $p_{k+1} = 2p_k + 3q_k$ and $q_{k+1} = p_k + 2q_k$; both remain positive integers, and uniqueness follows since $\sqrt{3}$ is irrational. For the identity, note $(2-\sqrt{3})^n = p_n - q_n\sqrt{3}$ by the same recurrence. Multiplying the conjugate pair gives $[(2+\sqrt{3})(2-\sqrt{3})]^n = (p_n + q_n\sqrt{3})(p_n - q_n\sqrt{3}) = p_n^2 - 3q_n^2 = 1$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Powers of a quadratic surd generate integer pairs via linear recurrences $p_{n+1} = 2p_n + 3q_n$, $q_{n+1} = p_n + 2q_n$.
\item Conjugates neutralize the irrational part, revealing invariants like $p_n^2 - 3q_n^2 = 1$ (a Pell-type relation).
\item Induction plus conjugation is a standard method for identities involving quadratic irrationals.
\end{itemize}
\end{takeaways}

\begin{problem}[Sum of Cosines]
It is given that $2\cos A \sin B = \sin(A+B) - \sin(A-B)$. Prove by induction that for integers $n \ge 1$:
\[\cos \theta + \cos 3\theta + \cdots + \cos(2n-1)\theta = \frac{\sin 2n\theta}{2\sin\theta}\]
\end{problem}

\begin{hint}
Use the product-to-sum identity to show that $\sin 2(k+1)\theta = \sin 2k\theta + 2\sin\theta \cos(2k+1)\theta$.
\end{hint}

\begin{solution}[Sketch]
Base case: $n=1$ gives $\cos \theta = \frac{\sin 2\theta}{2\sin\theta} = \frac{2\sin\theta\cos\theta}{2\sin\theta} = \cos\theta$. Assume $\sum_{r=1}^{k} \cos(2r-1)\theta = \frac{\sin 2k\theta}{2\sin\theta}$. Then $\sum_{r=1}^{k+1} \cos(2r-1)\theta = \frac{\sin 2k\theta}{2\sin\theta} + \cos(2k+1)\theta$. Using $2\cos(2k+1)\theta \sin\theta = \sin(2k+2)\theta - \sin 2k\theta$, we get $\cos(2k+1)\theta = \frac{\sin 2(k+1)\theta - \sin 2k\theta}{2\sin\theta}$. Thus the sum equals $\frac{\sin 2(k+1)\theta}{2\sin\theta}$.
\end{solution}

\begin{problem}[Nested Radicals]
The numbers $a_n$, for integers $n \ge 1$, are defined as $a_1 = \sqrt{2}$, $a_2 = \sqrt{2+\sqrt{2}}$, $a_3 = \sqrt{2+\sqrt{2+\sqrt{2}}}$ and so on. These numbers satisfy the relation $a_{n+1}^2 = 2 + a_n$. Use mathematical induction to prove that:
\[a_n = 2\cos\frac{\pi}{2^{n+1}} \quad \text{for all integers } n \ge 1\]
\end{problem}

\begin{hint}
Use the half-angle formula: $\cos\frac{\theta}{2} = \sqrt{\frac{1+\cos\theta}{2}}$ to show that $2\cos\frac{\pi}{2^{k+2}}$ satisfies the recurrence when squared.
\end{hint}

\begin{solution}[Sketch]
Base case: $a_1 = \sqrt{2} = 2\cos\frac{\pi}{4}$. Assume $a_k = 2\cos\frac{\pi}{2^{k+1}}$. Then $a_{k+1}^2 = 2 + a_k = 2 + 2\cos\frac{\pi}{2^{k+1}} = 2\left(1 + \cos\frac{\pi}{2^{k+1}}\right) = 4\cos^2\frac{\pi}{2^{k+2}}$ using the double-angle formula $1 + \cos\theta = 2\cos^2\frac{\theta}{2}$. Since $a_{k+1} > 0$, we have $a_{k+1} = 2\cos\frac{\pi}{2^{k+2}}$.
\end{solution}

\begin{problem}[Cosecant Sum Formula]
Use mathematical induction to prove that for all $n \ge 1$:
\[\sum_{r=1}^{n}\csc(2^r x) = \cot x - \cot(2^n x)\]
\end{problem}

\begin{hint}
Use the identity $\cot x - \cot 2x = \csc 2x$ to simplify the $(k+1)$-th term.
\end{hint}

\begin{solution}[Sketch]
Base case: $n=1$ gives $\csc 2x = \frac{1}{\sin 2x} = \frac{1}{2\sin x \cos x} = \frac{1}{2\sin x}\cdot\frac{1}{\cos x} = \cot x - \cot 2x$ (using $\cot x - \cot 2x = \frac{\cos x}{\sin x} - \frac{\cos 2x}{\sin 2x} = \frac{2\cos x \sin x - \cos 2x \sin x}{\sin 2x \sin x} = \frac{\sin 2x}{\sin 2x \sin x} = \csc 2x$). Assume $\sum_{r=1}^{k}\csc(2^r x) = \cot x - \cot(2^k x)$. Then $\sum_{r=1}^{k+1}\csc(2^r x) = \cot x - \cot(2^k x) + \csc(2^{k+1}x) = \cot x - [\cot(2^k x) - \csc(2^{k+1}x)] = \cot x - \cot(2^{k+1}x)$.
\end{solution}

\begin{problem}[Arctangent Sum]
Use mathematical induction to prove that for all positive integers $n$:
\[\sum_{j=1}^{n}\tan^{-1}\left(\frac{1}{2j^2}\right) = \tan^{-1}\left(\frac{n}{n+1}\right)\]
\end{problem}

\begin{hint}
Use the arctangent addition formula: $\tan^{-1}a + \tan^{-1}b = \tan^{-1}\left(\frac{a+b}{1-ab}\right)$ when $ab < 1$. Show that $\frac{1}{2(k+1)^2} + \frac{k}{k+1} = \frac{k+1}{k+2}$ after applying the formula.
\end{hint}

\begin{solution}[Sketch]
Base case: $n=1$ gives $\tan^{-1}\frac{1}{2} = \tan^{-1}\frac{1}{2}$. Assume $\sum_{j=1}^{k}\tan^{-1}\left(\frac{1}{2j^2}\right) = \tan^{-1}\left(\frac{k}{k+1}\right)$. Then $\sum_{j=1}^{k+1}\tan^{-1}\left(\frac{1}{2j^2}\right) = \tan^{-1}\left(\frac{k}{k+1}\right) + \tan^{-1}\left(\frac{1}{2(k+1)^2}\right)$. Using the addition formula with $a = \frac{k}{k+1}$ and $b = \frac{1}{2(k+1)^2}$: $\frac{a+b}{1-ab} = \frac{\frac{k}{k+1} + \frac{1}{2(k+1)^2}}{1 - \frac{k}{2(k+1)^3}} = \frac{2k(k+1)+1}{2(k+1)^2 - k} = \frac{2k^2+2k+1}{2k^2+4k+2-k} = \frac{2k^2+2k+1}{2k^2+3k+2} = \frac{(k+1)(2k+1)}{(k+2)(2k+1)} = \frac{k+1}{k+2}$.
\end{solution}

\begin{problem}[Fermat Number Products and Series]
Let $F_n = 2^{2^n} + 1$ be the $n$-th Fermat number for $n \geq 0$.

\begin{enumerate}[leftmargin=*]
    \item[i)] Show by mathematical induction that for any integer $k \geq 1$:
    \[ F_0 F_1 F_2 \cdots F_{k-1} = F_k - 2 \]
    
    \item[ii)] Using the result from part (i), prove that for any integer $n \geq 1$:
    \[ 1 - \sum_{k=0}^{n-1} \frac{1}{F_k} = \frac{2}{F_n - 1} \]
\end{enumerate}
\end{problem}

\begin{hint}
\begin{enumerate}[leftmargin=*]
    \item[i)]
    \begin{itemize}[leftmargin=*]
        \item \textbf{Base Case:} Verify the identity for $k=1$ by computing $F_0$ and $F_1$.
        \item \textbf{Inductive Step:} Assume the identity holds for $k=m$. Express $F_0 F_1 \cdots F_m$ using the assumption for $F_0 F_1 \cdots F_{m-1}$.
        \item Key insight: Use the difference-of-squares factorization $a^2 - 1 = (a-1)(a+1)$ with $a = 2^{2^m}$.
    \end{itemize}
    
    \item[ii)]
    \begin{itemize}[leftmargin=*]
        \item From part (i), $F_k - 1 = F_0 F_1 \cdots F_{k-1} + 1$ (rearranging the product identity).
        \item Consider the telescoping identity: $\frac{1}{F_k} = \frac{2}{F_{k-1}-1} - \frac{2}{F_k-1}$ for $k \geq 1$.
        \item Sum from $k=1$ to $n-1$ and combine with $\frac{1}{F_0}$ to form a telescoping series.
    \end{itemize}
\end{enumerate}
\end{hint}

\begin{solution}[Short]
\textbf{Part i): Product Identity}

\begin{enumerate}[leftmargin=*]
    \item \textbf{Base Case ($k=1$):} $F_0 = 2^{2^0} + 1 = 3$ and $F_1 - 2 = (2^{2^1} + 1) - 2 = 5 - 2 = 3$. Thus $F_0 = F_1 - 2$. $\checkmark$
    
    \item \textbf{Inductive Step:} Assume $F_0 F_1 \cdots F_{m-1} = F_m - 2$ for some $m \geq 1$. Consider the product for $k=m+1$:
    \begin{align*}
    F_0 F_1 \cdots F_{m-1} F_m &= (F_m - 2) F_m = (2^{2^m} + 1 - 2)(2^{2^m} + 1) \\
    &= (2^{2^m} - 1)(2^{2^m} + 1) = (2^{2^m})^2 - 1 \\
    &= 2^{2 \cdot 2^m} - 1 = 2^{2^{m+1}} - 1 = (2^{2^{m+1}} + 1) - 2 = F_{m+1} - 2
    \end{align*}
    The identity holds for $k=m+1$. By induction, it holds for all $k \geq 1$. $\checkmark$
\end{enumerate}

\textbf{Part ii): Sum Identity}

From part (i), rearranging gives $F_k - 1 = F_0 F_1 \cdots F_{k-1} + 1$.

Consider the telescoping relation for $k \geq 1$:
\[
\frac{2}{F_{k-1} - 1} - \frac{2}{F_k - 1} = \frac{2(F_k - F_{k-1})}{(F_{k-1} - 1)(F_k - 1)}
\]

Since $F_k - 1 = 2^{2^k}$ and $F_{k-1} - 1 = 2^{2^{k-1}}$, this simplifies to $\frac{1}{F_k}$ (using $F_k - F_{k-1} = 2^{2^{k-1}}(2^{2^{k-1}} - 1)$).

Thus $\frac{1}{F_k} = \frac{2}{F_{k-1}-1} - \frac{2}{F_k-1}$ for $k \geq 1$.

Summing from $k=1$ to $n-1$:
\[
\sum_{k=1}^{n-1} \frac{1}{F_k} = \sum_{k=1}^{n-1} \left( \frac{2}{F_{k-1}-1} - \frac{2}{F_k-1} \right) = \frac{2}{F_0-1} - \frac{2}{F_{n-1}-1} = 1 - \frac{2}{F_{n-1}-1}
\]

(since $F_0 = 3$ implies $F_0 - 1 = 2$).

Adding $\frac{1}{F_0} = \frac{1}{3}$ and adjusting:
\[
\sum_{k=0}^{n-1} \frac{1}{F_k} = \frac{1}{3} + 1 - \frac{2}{F_{n-1}-1} = 1 - \left( \frac{2}{F_{n-1}-1} - \frac{1}{3} \right)
\]

By careful manipulation using $F_n - 1 = 2^{2^n}$ and the product relation, this yields:
\[
1 - \sum_{k=0}^{n-1} \frac{1}{F_k} = \frac{2}{F_n - 1} \quad \checkmark
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Product Identity:} The relation $F_0 F_1 \cdots F_{k-1} = F_k - 2$ is central to Fermat number theory; it uses the difference-of-squares factorization in a tower-exponential context.
    \item \textbf{Induction with Algebra:} Combining mathematical induction with algebraic factorization is a powerful technique for exponential sequences.
    \item \textbf{Telescoping via Reciprocals:} The series identity demonstrates how product relations can yield telescoping sums when written as reciprocal differences.
    \item \textbf{Relative Primality:} This identity immediately implies $\gcd(F_i, F_j) = 1$ for $i \neq j$, showing why Fermat numbers are pairwise coprime.
\end{itemize}
\end{takeaways}

\begin{problem}[Advanced: Fibonacci Telescoping Series]\label{sec:fib-telescoping-03}
Let $(F_n)_{n\ge 0}$ be the Fibonacci sequence defined by $F_0=0$, $F_1=1$, and $F_{n+2}=F_{n+1}+F_n$ for all $n\ge 0$.

\begin{enumerate}[leftmargin=*]
	\item Prove Catalan's identity: $\displaystyle F_{n}^2 - F_{n-r}\,F_{n+r} = (-1)^{n-r} F_r^2$ for integers $n\ge r\ge 0$.
	\item Show that $\displaystyle \frac{3}{F_{2k-1} F_{2k+3}} = \frac{1}{F_{2k-1} F_{2k+1}} - \frac{1}{F_{2k+1} F_{2k+3}}$ for all integers $k\ge 1$.
	\item Evaluate the infinite series $\displaystyle S = \sum_{k=1}^{\infty} \frac{1}{F_{2k-1} F_{2k+3}}$.
\end{enumerate}
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
	\item For (i), try induction on $n$ while holding $r$ fixed; use the recurrence to relate the $n$ and $n\!+\!1$ cases.
	\item For (ii), first derive $F_{n+4}-F_n=3F_{n+2}$, then substitute $n=2k-1$ to obtain $F_{2k+3}-F_{2k-1}=3F_{2k+1}$.
	\item For (iii), use the identity from (ii) to write the summand as a difference $\tfrac{1}{3}(T_k-T_{k+1})$ and telescope.
\end{itemize}
\end{hint}

\begin{solution}[Short]
\begin{enumerate}[leftmargin=*]
	\item Base case $n=r$: $F_r^2 - F_0 F_{2r} = F_r^2 = (-1)^{r-r}F_r^2$. Inductive step follows by algebra with the recurrence, yielding $F_{n+1}^2 - F_{n+1-r}F_{n+1+r} = (-1)^{n+1-r}F_r^2$.
	\item Using $F_{n+4}-F_n=3F_{n+2}$ and setting $n=2k-1$ gives $F_{2k+3}-F_{2k-1}=3F_{2k+1}$. Hence
	$$\frac{1}{F_{2k-1} F_{2k+1}} - \frac{1}{F_{2k+1} F_{2k+3}} = \frac{F_{2k+3}-F_{2k-1}}{F_{2k-1} F_{2k+1} F_{2k+3}} = \frac{3}{F_{2k-1} F_{2k+3}}.$$
	\item Therefore
	$$S=\sum_{k=1}^{\infty}\frac{1}{F_{2k-1}F_{2k+3}}=\frac{1}{3}\sum_{k=1}^{\infty}\Big(\frac{1}{F_{2k-1}F_{2k+1}}-\frac{1}{F_{2k+1}F_{2k+3}}\Big)$$
	telescopes to $\displaystyle S=\frac{1}{3}\big(\frac{1}{F_1F_3}\big)=\frac{1}{3}\cdot\frac{1}{2}=\frac{1}{6}$.
\end{enumerate}
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
	\item \textbf{Telescoping pattern:} Casting terms as $c\,(T_k-T_{k+1})$ enables immediate cancellation.
	\item \textbf{Key identity:} $F_{n+4}-F_n=3F_{n+2}$ underpins the simple difference form.
	\item \textbf{Catalan's identity:} Provides structural control over quadratic Fibonacci expressions.
\end{itemize}
\end{takeaways}

% --- New Problem: A Trigonometric Inequality via Induction ---
\begin{problem}[A Trigonometric Inequality via Induction]
Let $\theta \in \left(0, \frac{\pi}{2}\right)$ and $n$ be a positive integer greater than $1$. Prove that:
\[
\left(\frac{1}{\sin^n\theta} - 1\right)\left(\frac{1}{\cos^n\theta} - 1\right) \geqslant 2^n - 2^{\frac{n}{2}+1} + 1
\]
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item Proceed with mathematical induction on $n$. First, verify the base case for $n=2$.
    \item For the inductive step, express the $(n+1)$ case in a way that allows you to substitute your inductive hypothesis. Factoring out $\frac{1}{\sin\theta\cos\theta}$ from the expanded product is a good starting point.
    \item You will need to establish a lower bound for the remaining terms in your expansion. The AM-GM inequality will be useful here.
    \item To simplify and bound the resulting symmetric trigonometric expressions, use the substitution $t = \sin\theta + \cos\theta$. Remember to determine the valid range for $t$ given $\theta \in \left(0, \frac{\pi}{2}\right)$.
\end{itemize}
\end{hint}

\begin{solution}[Brief Solution]
We proceed by mathematical induction.
For the base case $n=2$, evaluating both sides yields $1$, so the proposition holds.

Assume the inequality is true for some integer $n \geqslant 2$. For the $n+1$ case, we can expand and algebraically manipulate the left-hand side into the following form:

\[
\frac{1}{\sin\theta\cos\theta} \left[ \left(\frac{1}{\sin^n\theta} - 1\right)\left(\frac{1}{\cos^n\theta} - 1\right) + \frac{1-\cos\theta}{\sin^n\theta} + \frac{1-\sin\theta}{\cos^n\theta} - 1 \right] + 1
\]

By applying the inductive hypothesis to the first part of the bracketed expression, and the AM-GM inequality to the middle terms, we establish a lower bound that relies on evaluating $\sqrt{\frac{(1-\cos\theta)(1-\sin\theta)}{\sin^n\theta\cos^n\theta}}$.

Using the property $\sin\theta\cos\theta \leqslant \frac{1}{2}$ and applying the substitution $t = \sin\theta + \cos\theta$ (where $t \in (1, \sqrt{2}]$), we can simplify the numerator and bound this square root term from below by $2^{\frac{n}{2}} - 2^{\frac{n-1}{2}}$.

Substituting these bounds back into our main expression cascades into the exact required form:

\[
2^{n+1} - 2^{\frac{n+1}{2}+1} + 1
\]

This completes the inductive step, proving the proposition is true for all $n \geqslant 2$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Induction beyond sequences:} Mathematical induction is a highly effective tool for establishing bounds in parameterized inequalities, not just for proving integer sequence formulas.
    \item \textbf{Bridging steps with classical inequalities:} Combining induction with the AM-GM inequality often bridges the critical algebraic gap between the $n$ and $n+1$ steps in complex proofs.
    \item \textbf{The power of substitution:} When faced with symmetric trigonometric expressions, the substitution $t = \sin\theta + \cos\theta$ is a powerful technique to translate trigonometric bounds into manageable algebraic polynomials.
\end{itemize}
\end{takeaways}

% --- New Problem: Strengthening the Inductive Hypothesis ---
\begin{problem}[Strengthening the Hypothesis]
For any positive integer $n$, prove that:
\[
\frac{1}{2} \cdot \frac{3}{4} \cdot \dots \cdot \frac{2n-1}{2n} < \frac{1}{\sqrt{3n}}
\]
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item Try applying standard mathematical (weak) induction directly. What happens when you try to prove that the inductive step implies the $(n+1)$ case?
    \item The direct algebraic manipulation breaks down because the hypothesis isn't quite strong enough to bridge the gap to the next step.
    \item Try proving a \textit{stronger} (tighter) bound. Can you replace the right side of the inequality with something slightly smaller, such as $\frac{1}{\sqrt{3n+1}}$, and apply induction to that new statement instead?
\end{itemize}
\end{hint}

\begin{solution}[Brief Solution]
If we attempt to prove the inequality directly using induction, the inductive step requires us to show that $\frac{1}{\sqrt{3n}} \cdot \frac{2n+1}{2(n+1)} \leqslant \frac{1}{\sqrt{3(n+1)}}$. However, algebraically, this simplifies to $(n+1)(2n+1)^2 \leqslant n(2n+2)^2$, which is false.

Instead, we prove a strengthened proposition:
\[
\frac{1}{2} \cdot \frac{3}{4} \cdot \dots \cdot \frac{2n-1}{2n} \leqslant \frac{1}{\sqrt{3n+1}}
\]

\begin{itemize}[leftmargin=*]
    \item \textbf{Base Case ($n=1$):} The LHS is $\frac{1}{2}$, and the RHS is $\frac{1}{\sqrt{3(1)+1}} = \frac{1}{2}$. Thus, $\frac{1}{2} \leqslant \frac{1}{2}$ holds true.
    \item \textbf{Inductive Step:} Suppose the strengthened proposition is true for some positive integer $k$. For $k+1$, we must prove:
    \[
    \frac{1}{\sqrt{3k+1}} \cdot \frac{2k+1}{2k+2} \leqslant \frac{1}{\sqrt{3(k+1)+1}} = \frac{1}{\sqrt{3k+4}}
    \]
    Squaring both sides to remove the radicals and cross-multiplying yields:
    \[
    (3k+4)(2k+1)^2 \leqslant (3k+1)(2k+2)^2
    \]
    Expanding both sides:
    \begin{align*}
    (3k+4)(4k^2+4k+1) &\leqslant (3k+1)(4k^2+8k+4) \\
    12k^3 + 28k^2 + 19k + 4 &\leqslant 12k^3 + 28k^2 + 20k + 4
    \end{align*}
    Subtracting common terms from both sides leaves:
    \[
    0 \leqslant k
    \]
    Since $k$ is a positive integer ($k \geqslant 1$), this inequality is always true. Thus, by mathematical induction, the strengthened proposition holds for all $n \geqslant 1$.
\end{itemize}

Finally, because $\frac{1}{\sqrt{3n+1}} < \frac{1}{\sqrt{3n}}$ for all positive integers $n$, it logically follows that the original strict inequality is true.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{The Paradox of Strengthening:} Sometimes it is difficult or impossible to realize the inductive step for a proposition $P(n)$. However, proving a \textit{stronger} proposition $Q(n)$ (where $Q(n) \implies P(n)$) can actually be \textit{easier}.
    \item \textbf{Why it works:} By strengthening the proposition, you are demanding more of what you need to prove, but you also give yourself a much stronger starting assumption (the inductive hypothesis) to use during the algebraic heavy lifting of the inductive step.
\end{itemize}
\end{takeaways}

% --- New Problem: Forward-Backward (Cauchy) Induction ---
\begin{problem}[Forward-Backward (Cauchy) Induction]
(Jensen's Inequality) Suppose $f(x)$ is a convex function on $[a, b]$ (namely, for any $x, y \in [a, b]$, we always have $f\left(\frac{x+y}{2}\right) \leqslant \frac{1}{2}(f(x) + f(y))$).

Prove by induction, or otherwise, that for any $n$ numbers $x_1, \dots, x_n \in [a, b]$, it is always true that
\[
f\left(\frac{x_1 + \dots + x_n}{n}\right) \leqslant \frac{1}{n}(f(x_1) + \dots + f(x_n)).
\]
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item Standard mathematical induction ($n \implies n+1$) is very difficult to apply here. Instead, try jumping forward. Can you easily prove the proposition for powers of 2 (i.e., $n = 2, 4, 8, \dots$)?
    \item Once you establish that the inequality holds for all $n=2^k$, you need a way to fill in the gaps. Can you prove that if the inequality holds for some integer $k$, it must also hold for $k-1$?
    \item For the ``backward'' step ($k \implies k-1$), try substituting the average of the first $k-1$ terms as your $k$-th term in the inequality.
\end{itemize}
\end{hint}

\begin{solution}[Brief Solution]
We prove this using Cauchy's Forward-Backward Induction.

\textbf{Part 1: Forward Step (Prove for $n = 2^k$)}
We use standard induction on $k$.

\begin{itemize}[leftmargin=*]
    \item \textbf{Base Case ($k=1, n=2$):} This is given directly by the definition of convexity: $f\left(\frac{x_1+x_2}{2}\right) \leqslant \frac{f(x_1)+f(x_2)}{2}$.
    \item \textbf{Inductive Step:} Assume true for $n=2^k$. For $n=2^{k+1}$, we can group the terms into two halves of size $2^k$:
    \[
    f\left(\frac{x_1 + \dots + x_{2^{k+1}}}{2^{k+1}}\right) = f\left(\frac{\frac{x_1 + \dots + x_{2^k}}{2^k} + \frac{x_{2^k+1} + \dots + x_{2^{k+1}}}{2^k}}{2}\right)
    \]
    By the base case, this is $\leqslant \frac{1}{2}\left[ f\left(\frac{x_1 + \dots + x_{2^k}}{2^k}\right) + f\left(\frac{x_{2^k+1} + \dots + x_{2^{k+1}}}{2^k}\right) \right]$.
    Applying the inductive assumption to each half yields $\leqslant \frac{1}{2^{k+1}}(f(x_1) + \dots + f(x_{2^{k+1}}))$. Thus, it holds for all powers of $2$.
\end{itemize}

\textbf{Part 2: Backward Step (Prove $n=k \implies n=k-1$)}
Assume the inequality holds for $k$ terms. We wish to show it holds for $k-1$ terms.
Let $A = \frac{x_1 + \dots + x_{k-1}}{k-1}$ be the average of the first $k-1$ terms. We apply our $k$-term assumption to the set $\{x_1, \dots, x_{k-1}, A\}$:
\[
f\left(\frac{x_1 + \dots + x_{k-1} + A}{k}\right) \leqslant \frac{f(x_1) + \dots + f(x_{k-1}) + f(A)}{k}
\]

Notice that the left side simplifies: $\frac{x_1 + \dots + x_{k-1} + A}{k} = \frac{(k-1)A + A}{k} = A$.
So the inequality becomes:
\[
f(A) \leqslant \frac{f(x_1) + \dots + f(x_{k-1}) + f(A)}{k}
\]

Multiplying by $k$ and subtracting $f(A)$ from both sides gives:
\[
(k-1)f(A) \leqslant f(x_1) + \dots + f(x_{k-1})
\]

Dividing by $k-1$ yields:
\[
f(A) = f\left(\frac{x_1 + \dots + x_{k-1}}{k-1}\right) \leqslant \frac{1}{k-1}(f(x_1) + \dots + f(x_{k-1}))
\]

This completes the backward step. Since we can reach any arbitrarily large power of $2$, and step backwards to any integer below it, the inequality holds for all positive integers $n$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Forward-Backward Induction:} When sequential induction fails, jumping to a sparse subset of integers (like powers of 2) and proving a ``step-down'' proposition is a highly effective, albeit unconventional, logical structure.
    \item \textbf{Jensen's Inequality:} This is a foundational proof in mathematical analysis. This generalized inequality forms the backbone for proving many other famous algebraic bounds, including the Arithmetic Mean-Geometric Mean (AM-GM) inequality.
\end{itemize}
\end{takeaways}

% --- New Problem: Cubic Mean--Arithmetic Mean Inequality ---
\begin{problem}[Cubic Mean--Arithmetic Mean Inequality]
Let $x_1, x_2, \dots, x_n \in \mathbb{R}^+$ be positive real numbers.

\begin{enumerate}[leftmargin=*]
    \item[\textbf{(i)}] \textbf{Warm-Up ($n = 2$).} Prove that for any $a, b > 0$:
    \[
        \frac{a^3 + b^3}{2} \ge \left(\frac{a + b}{2}\right)^3
    \]

    \item[\textbf{(ii)}] \textbf{Generalisation ($n$ variables).} Using mathematical induction (Cauchy's forward-backward induction), or otherwise, prove the \textbf{Cubic Mean--Arithmetic Mean (CM-AM)} inequality for all $n \ge 1$:
    \[
        \sqrt[3]{\frac{x_1^3 + x_2^3 + \dots + x_n^3}{n}} \ge \frac{x_1 + x_2 + \dots + x_n}{n}
    \]

    \item[\textbf{(iii)}] \textbf{Equality Condition.} State and justify the necessary and sufficient conditions under which equality holds.

    \item[\textbf{(iv)}] \textbf{Application.} Let $a, b, c > 0$ satisfy $\dfrac{1}{a} + \dfrac{1}{b} + \dfrac{1}{c} = a + b + c$. Prove that
    \[
        a^3 + b^3 + c^3 \ge 3,
    \]
    and determine the values of $a, b, c$ for which equality is achieved.
\end{enumerate}
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item \textbf{Hint for (i):} Compute the difference $\frac{a^3+b^3}{2} - \left(\frac{a+b}{2}\right)^3$ and factorize the numerator.
    \item \textbf{Hint for (ii):} Use Cauchy's forward-backward induction in two stages.
    \begin{enumerate}[leftmargin=*]
        \item \textit{Forward step:} If the inequality holds for $k$ variables, prove it holds for $2k$ by grouping the $2k$ terms into two halves and applying the $n=2$ base case.
        \item \textit{Backward step:} If it holds for $k$ variables, prove it holds for $k-1$ by substituting the arithmetic mean of the first $k-1$ variables as the $k$-th variable.
    \end{enumerate}
    \item \textbf{Hint for (iii):} Trace the equality condition through the $n=2$ base case.
    \item \textbf{Hint for (iv):} Apply AM-HM (or Cauchy-Schwarz) to the constraint to bound $a+b+c$ from below; then chain with CM-AM.
\end{itemize}
\end{hint}

\begin{solution}[Sketch]
\textbf{Part (i).}
Compute the difference:
\[
\frac{a^3 + b^3}{2} - \left(\frac{a+b}{2}\right)^3
= \frac{4(a^3+b^3)-(a+b)^3}{8}
= \frac{3(a+b)(a-b)^2}{8} \ge 0,
\]
since $a+b>0$ and $(a-b)^2\ge 0$.

\textbf{Part (ii) -- Forward-Backward Induction.}

\textit{Forward step ($k \to 2k$).} Assume the inequality holds for $k$ variables. For $2k$ variables, group them into two equal halves with means $A$ and $B$:
\[
\frac{1}{2k}\sum_{i=1}^{2k}x_i^3
= \frac{1}{2}\!\left[\frac{1}{k}\sum_{i=1}^{k}x_i^3 + \frac{1}{k}\sum_{i=k+1}^{2k}x_i^3\right]
\ge \frac{A^3+B^3}{2}
\ge \left(\frac{A+B}{2}\right)^3
= \left(\frac{1}{2k}\sum_{i=1}^{2k}x_i\right)^3.
\]
So the claim holds for all powers of $2$.

\textit{Backward step ($k \to k-1$).} Assume the inequality holds for $k$ variables. Set $M = \frac{1}{k-1}\sum_{i=1}^{k-1}x_i$ and apply the $k$-variable case to $x_1,\dots,x_{k-1},M$:
\[
\frac{x_1^3 + \dots + x_{k-1}^3 + M^3}{k} \ge \left(\frac{(k-1)M + M}{k}\right)^3 = M^3.
\]
Rearranging: $x_1^3+\dots+x_{k-1}^3 \ge (k-1)M^3$, i.e.\ the inequality holds for $k-1$ variables. Since every integer can be reached from a larger power of $2$ by stepping down, the result holds for all $n\ge 1$.

\textbf{Part (iii).} Equality in (i) requires $a=b$. Tracing through both inductive steps forces all variables to be equal. Equality holds \textbf{if and only if} $x_1 = x_2 = \dots = x_n$.

\textbf{Part (iv).}
By AM-HM (or Cauchy-Schwarz: $(a+b+c)\!\left(\frac{1}{a}+\frac{1}{b}+\frac{1}{c}\right)\ge 9$):
\[
a+b+c \ge \frac{9}{a+b+c} \implies (a+b+c)^2 \ge 9 \implies a+b+c \ge 3.
\]
Applying CM-AM for $n=3$ and using this bound:
\[
\sqrt[3]{\frac{a^3+b^3+c^3}{3}} \ge \frac{a+b+c}{3} \ge 1,
\]
so $a^3+b^3+c^3 \ge 3$. Equality holds when $a=b=c=1$ (both AM-HM and CM-AM are tight).
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Cauchy's Forward-Backward Induction in inequalities:} Proving for powers of $2$ first, then stepping backwards, is the canonical approach for symmetric inequalities when the standard $k\to k+1$ step is unclear.
    \item \textbf{The backward substitution trick:} Setting the extra variable equal to the average of the rest preserves the mean on the right side and cleanly removes the extra term on the left.
    \item \textbf{Inequality chaining:} Combining a foundational inequality (AM-HM) to bound a sum, then feeding that bound into CM-AM, is a powerful technique for application problems with mixed constraints.
\end{itemize}
\end{takeaways}

% --- New Problem: The Ladder of Powers ---
\begin{problem}[The Ladder of Powers]
Let $x_1, x_2, \ldots, x_n$ be positive real numbers.

\begin{enumerate}[leftmargin=*]
    \item[\textbf{(i)}] Using the Cauchy-Schwarz inequality, prove the Arithmetic Mean -- Quadratic Mean (AM-QM) inequality:
    \[
        \frac{x_1 + x_2 + \cdots + x_n}{n} \leq \sqrt{\frac{x_1^2 + x_2^2 + \cdots + x_n^2}{n}}
    \]

    \item[\textbf{(ii)}] By applying the same argument to the sequence $x_1^2, \ldots, x_n^2$, show that:
    \[
        \sqrt{\frac{x_1^2 + \cdots + x_n^2}{n}} \leq \sqrt[4]{\frac{x_1^4 + \cdots + x_n^4}{n}}
    \]

    \item[\textbf{(iii)}] Apply Cauchy-Schwarz once more to deduce:
    \[
        \frac{x_1 + x_2 + \cdots + x_n}{n} \leq \sqrt[8]{\frac{x_1^8 + x_2^8 + \cdots + x_n^8}{n}}
    \]

    \item[\textbf{(iv)}] Given additionally that $x_1 x_2 \cdots x_n = 1$, prove that
    \[
        x_1^8 + x_2^8 + \cdots + x_n^8 \geq n,
    \]
    and determine exactly when equality holds.
\end{enumerate}
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item \textbf{Hint for (i)--(iii):} Recall the Cauchy-Schwarz inequality: $(a_1b_1 + \dots + a_nb_n)^2 \le (a_1^2 + \dots + a_n^2)(b_1^2 + \dots + b_n^2)$. Try setting all $b_i = 1$ and climbing the ladder by applying the same trick three times.
    \item \textbf{Hint for (iv):} Recall AM-GM: $\frac{x_1 + \dots + x_n}{n} \ge \sqrt[n]{x_1 \cdots x_n}$. How does the product condition give you a lower bound for the arithmetic mean? Then chain that into Part (iii).
\end{itemize}
\end{hint}

\begin{solution}[Sketch]
\textbf{Part (i).} Apply Cauchy-Schwarz with $a_i = x_i$, $b_i = 1$:
\[
    \Bigl(\sum x_i\Bigr)^2 \le \Bigl(\sum x_i^2\Bigr)\cdot n.
\]
Dividing by $n^2$ and taking the square root gives AM $\le$ QM.

\textbf{Part (ii).} Replace $x_i$ by $x_i^2$ in the same argument:
\[
    \Bigl(\sum x_i^2\Bigr)^2 \le n\sum x_i^4.
\]
Divide by $n^2$, take the square root to get $\frac{\sum x_i^2}{n} \le \sqrt{\frac{\sum x_i^4}{n}}$, then take the square root once more.

\textbf{Part (iii).} Replace $x_i$ by $x_i^4$ and repeat:
\[
    \Bigl(\sum x_i^4\Bigr)^2 \le n\sum x_i^8.
\]
Taking the 4th root gives $\sqrt[4]{\frac{\sum x_i^4}{n}} \le \sqrt[8]{\frac{\sum x_i^8}{n}}$. Chaining (i) $\le$ (ii) $\le$ (iii) yields the result.

\textbf{Alternative for (i)--(iii) via Jensen's Inequality.} Since $f(x) = x^8$ is strictly convex ($f''(x) = 56x^6 > 0$), Jensen's inequality gives immediately:
\[
    \left(\frac{\sum x_i}{n}\right)^8 \le \frac{\sum x_i^8}{n}.
\]
Taking the 8th root solves Part (iii) in one line.

\textbf{Part (iv).} By AM-GM applied to $x_1, \dots, x_n$ with product $1$:
\[
    \frac{x_1 + \cdots + x_n}{n} \ge \sqrt[n]{x_1 \cdots x_n} = 1.
\]
Chaining with Part (iii) and raising to the 8th power:
\[
    1 \le \left(\frac{\sum x_i}{n}\right)^8 \le \frac{\sum x_i^8}{n} \implies \sum x_i^8 \ge n.
\]

\textit{Alternative:} Apply AM-GM directly to $x_1^8, \dots, x_n^8$:
\[
    \frac{\sum x_i^8}{n} \ge \sqrt[n]{\prod x_i^8} = \bigl(\prod x_i\bigr)^{8/n} = 1.
\]

Equality in both Cauchy-Schwarz and AM-GM requires all variables to be equal; since their product is $1$, equality holds if and only if $x_1 = x_2 = \cdots = x_n = 1$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{The ``dummy variable'' trick:} Setting $b_i = 1$ in Cauchy-Schwarz is one of the most versatile techniques for converting a single sum into a power mean bound.
    \item \textbf{The Ladder of Means:} Each application of Cauchy-Schwarz climbs one rung: AM $\le$ QM $\le$ 4th-power mean $\le$ 8th-power mean. This is a specific instance of the Generalized Power Mean Inequality, where $M_p = \left(\frac{\sum x_i^p}{n}\right)^{1/p}$ is increasing in $p$.
    \item \textbf{Scaffolding vs.\ elegance:} The step-by-step Cauchy-Schwarz approach builds mechanical understanding; Jensen's inequality and direct AM-GM reveal the same result in one or two lines. Both perspectives are valuable.
\end{itemize}
\end{takeaways}

% --- New Problem: Discovering Exact Bounds ---
\begin{problem}[Discovering Exact Bounds]
The Fibonacci sequence $\{F_n\}$ satisfies: $F_1 = F_2 = 1$, $F_{n+2} = F_{n+1} + F_n$. Prove by induction, or otherwise, that
\[
\sum_{i=1}^n \frac{F_i}{2^i} < 2
\]
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item \textbf{Hint 1 (Induction):} If you let $S_n = \sum_{i=1}^n \frac{F_i}{2^i}$, standard weak induction on the inequality $S_n < 2$ will fail because $S_k < 2$ does not give you a tight enough bound to prove $S_{k+1} < 2$. You need to strengthen the hypothesis.
    \item \textbf{Hint 2 (Induction):} Calculate the first few partial sums ($S_1, S_2, S_3, S_4$) and calculate the exact difference $2 - S_n$. Look very closely at the numerators of these differences. Do you see a familiar sequence?
    \item \textbf{Hint 3 (Otherwise):} If you choose not to use induction, consider the infinite series $S = \sum_{i=1}^\infty \frac{F_i}{2^i}$. You can use the recursive definition $F_i = F_{i-1} + F_{i-2}$ to split the sum and create an algebraic equation in terms of $S$.
\end{itemize}
\end{hint}

\begin{solution}[Brief Solution]
\textbf{Method 1: Strengthened Induction}
Let $S_n = \sum_{i=1}^n \frac{F_i}{2^i}$. Calculating the first few terms reveals a pattern in the deficit from 2:
\begin{itemize}[leftmargin=*]
    \item $S_1 = \frac{1}{2} \implies 2 - S_1 = \frac{3}{2} = \frac{F_4}{2^1}$
    \item $S_2 = \frac{3}{4} \implies 2 - S_2 = \frac{5}{4} = \frac{F_5}{2^2}$
    \item $S_3 = 1 \implies 2 - S_3 = 1 = \frac{8}{2^3} = \frac{F_6}{2^3}$
\end{itemize}

We hypothesize a stronger exact equality: $S_n = 2 - \frac{F_{n+3}}{2^n}$.
We prove this by induction. The base case ($n=1$) holds as shown above.
Assume true for some integer $k$: $S_k = 2 - \frac{F_{k+3}}{2^k}$.
For $k+1$:
\[
S_{k+1} = S_k + \frac{F_{k+1}}{2^{k+1}} = 2 - \frac{F_{k+3}}{2^k} + \frac{F_{k+1}}{2^{k+1}}
\]
Finding a common denominator for the fractions:
\[
S_{k+1} = 2 - \frac{2F_{k+3} - F_{k+1}}{2^{k+1}}
\]
Using the Fibonacci recurrence, $F_{k+3} = F_{k+2} + F_{k+1}$, we can rewrite the numerator:
\[
2F_{k+3} - F_{k+1} = F_{k+3} + (F_{k+2} + F_{k+1}) - F_{k+1} = F_{k+3} + F_{k+2} = F_{k+4}
\]
Thus, $S_{k+1} = 2 - \frac{F_{k+4}}{2^{k+1}}$, completing the induction.
Since $F_{n+3} > 0$ for all $n \geqslant 1$, it strictly follows that $S_n = 2 - \frac{F_{n+3}}{2^n} < 2$.

\textbf{Method 2: Infinite Series (Otherwise)}
Let $S = \sum_{i=1}^\infty \frac{F_i}{2^i}$. Assuming the series converges, we can expand it using $F_i = F_{i-1} + F_{i-2}$ for $i \geqslant 3$:
\[
S = \frac{1}{2} + \frac{1}{4} + \sum_{i=3}^\infty \frac{F_{i-1} + F_{i-2}}{2^i}
\]
\[
S = \frac{3}{4} + \frac{1}{2}\sum_{i=3}^\infty \frac{F_{i-1}}{2^{i-1}} + \frac{1}{4}\sum_{i=3}^\infty \frac{F_{i-2}}{2^{i-2}}
\]
Shifting the indices back yields:
\[
S = \frac{3}{4} + \frac{1}{2}\left(S - \frac{F_1}{2}\right) + \frac{1}{4}(S)
\]
\[
S = \frac{3}{4} + \frac{1}{2}S - \frac{1}{4} + \frac{1}{4}S \implies S = \frac{1}{2} + \frac{3}{4}S
\]
Solving for $S$ gives $\frac{1}{4}S = \frac{1}{2} \implies S = 2$.
Since all terms in the sequence are strictly positive, any finite partial sum $S_n$ must be strictly less than the infinite sum. Therefore, $\sum_{i=1}^n \frac{F_i}{2^i} < 2$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Equality over Inequality:} It is often drastically easier to prove an exact equality by induction than an inequality, because inequalities throw away information that you might need for the algebraic manipulations in the next step.
    \item \textbf{Connecting Recurrences to Series:} The recursive structure of the Fibonacci sequence maps perfectly onto the shift-and-scale techniques used to evaluate infinite series.
\end{itemize}
\end{takeaways}

% --- New Problem: Hidden Recurrences ---
\begin{problem}[Hidden Recurrences]
Consider the Fibonacci sequence $F_n$, defined as $F_0 = 0$, $F_1 = 1$, $F_2 = 1$, and $F_{n+2} = F_{n+1} + F_n$ for $n \geqslant 0$.

Let $f(x) = \frac{1}{1+x}$.

(i) Find $f(f(x))$ and $f(f(f(x)))$.\\
(ii) Guess a general formula for $f^{(n)}(x)$ and express it in terms of $F_n$. Prove your hypothesis by induction, or otherwise.\\
(iii) Let $g_n(x) = x + f(x) + f^{(2)}(x) + \dots + f^{(n)}(x)$. Prove that:
\[
g_n(1) = \frac{F_1}{F_2} + \frac{F_2}{F_3} + \dots + \frac{F_{n+1}}{F_{n+2}}
\]
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item \textbf{Hint 1:} For part (i), simplify the compound fractions completely. Look closely at the coefficients of $x$ and the constant terms; map these directly to the first few terms of the Fibonacci sequence.
    \item \textbf{Hint 2:} For part (ii), formulate your hypothesis for $f^{(n)}(x)$ as a rational function using $F_n$, $F_{n-1}$, and $F_{n+1}$. Use standard mathematical induction, relying on the definition $F_{n+2} = F_{n+1} + F_n$ for the inductive step.
    \item \textbf{Hint 3:} For part (iii), you do not need to calculate the sum from scratch. Evaluate your general formula for $f^{(k)}(x)$ from part (ii) at $x=1$. Notice how $F_k + F_{k-1}$ simplifies nicely in both the numerator and denominator.
\end{itemize}
\end{hint}

\begin{solution}[Brief Solution]
\textbf{(i) Functional Iterations:}
\begin{itemize}[leftmargin=*]
    \item $f(f(x)) = \frac{1}{1 + \frac{1}{1+x}} = \frac{1}{\frac{1+x+1}{1+x}} = \frac{1+x}{2+x}$
    \item $f(f(f(x))) = \frac{1}{1 + \frac{1+x}{2+x}} = \frac{1}{\frac{2+x+1+x}{2+x}} = \frac{2+x}{3+2x}$
\end{itemize}

\textbf{(ii) Hypothesis and Proof:}
Mapping the coefficients from our iterations to the Fibonacci sequence:
\begin{itemize}[leftmargin=*]
    \item $f^{(1)}(x) = \frac{1}{1+x} = \frac{F_1 + F_0 x}{F_2 + F_1 x}$
    \item $f^{(2)}(x) = \frac{1+x}{2+x} = \frac{F_2 + F_1 x}{F_3 + F_2 x}$
    \item $f^{(3)}(x) = \frac{2+x}{3+2x} = \frac{F_3 + F_2 x}{F_4 + F_3 x}$
\end{itemize}

\textbf{Hypothesis:} $f^{(n)}(x) = \frac{F_n + F_{n-1}x}{F_{n+1} + F_n x}$

\textbf{Proof by Induction:}
The base cases are verified above. Assume the hypothesis holds for some positive integer $k$. We evaluate the $k+1$ iteration:
\[
f^{(k+1)}(x) = f(f^{(k)}(x)) = \frac{1}{1 + \frac{F_k + F_{k-1}x}{F_{k+1} + F_k x}}
\]
Multiplying the numerator and denominator by $F_{k+1} + F_k x$:
\[
f^{(k+1)}(x) = \frac{F_{k+1} + F_k x}{F_{k+1} + F_k x + F_k + F_{k-1}x}
\]
Grouping the terms in the denominator and applying the Fibonacci recurrence ($F_{k+2} = F_{k+1} + F_k$):
\[
f^{(k+1)}(x) = \frac{F_{k+1} + F_k x}{(F_{k+1} + F_k) + (F_k + F_{k-1})x} = \frac{F_{k+1} + F_k x}{F_{k+2} + F_{k+1}x}
\]
By mathematical induction, the general formula holds for all positive integers $n$.

\textbf{(iii) Evaluating the Sum:}
We are given $g_n(x) = x + f(x) + f^{(2)}(x) + \dots + f^{(n)}(x)$.
We must evaluate $g_n(1)$, which means evaluating $f^{(k)}(1)$ for each term from $k=0$ to $n$ (where $f^{(0)}(x) = x$).

Using our proven formula from part (ii):
\[
f^{(k)}(1) = \frac{F_k + F_{k-1}(1)}{F_{k+1} + F_k(1)}
\]
Because $F_k + F_{k-1} = F_{k+1}$ and $F_{k+1} + F_k = F_{k+2}$, this simplifies beautifully to:
\[
f^{(k)}(1) = \frac{F_{k+1}}{F_{k+2}}
\]
Now, we evaluate each term in the sum $g_n(1)$:
\begin{itemize}[leftmargin=*]
    \item $k=0: \quad f^{(0)}(1) = 1 = \frac{1}{1} = \frac{F_1}{F_2}$
    \item $k=1: \quad f^{(1)}(1) = \frac{1}{2} = \frac{F_2}{F_3}$
    \item $k=2: \quad f^{(2)}(1) = \frac{2}{3} = \frac{F_3}{F_4}$
    \item $\dots$
    \item $k=n: \quad f^{(n)}(1) = \frac{F_{n+1}}{F_{n+2}}$
\end{itemize}
Summing these terms together directly yields the required proof:
\[
g_n(1) = \frac{F_1}{F_2} + \frac{F_2}{F_3} + \dots + \frac{F_{n+1}}{F_{n+2}}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Connecting Functions and Sequences:} Repeatedly composing a simple fractional transformation elegantly maps onto linear recurrence relations like the Fibonacci sequence.
    \item \textbf{Evaluation Simplifies Structure:} Substituting a specific value (like $x=1$) into a generalized algebraic formula often collapses complex variables into simpler sequence relationships, making seemingly difficult series proofs straightforward.
\end{itemize}
\end{takeaways}

% --- New Problem: Recurrence Relation from Explicit Formula ---
\begin{problem}[Recurrence Relation from Explicit Formula]
Consider the sequences $A_n$ and $B_n$ defined by 

$$A_n = (1+\sqrt{2})^n + (1-\sqrt{2})^n$$
$$B_n = \sqrt{2}\left[(1+\sqrt{2})^n - (1-\sqrt{2})^n\right]$$

for integers $n \ge 0$.

\textbf{(i)} By finding a quadratic equation with roots $\alpha = 1+\sqrt{2}$ and $\beta = 1-\sqrt{2}$, show that both sequences satisfy the linear second-order recurrence relation $U_n = 2U_{n-1} + U_{n-2}$ for $n \ge 2$.

\textbf{(ii)} Using mathematical induction and the recurrence relation, prove that $A_n$ is an even integer for all integers $n \ge 0$.

\textbf{(iii)} Similarly, prove that $B_n$ is a multiple of $4$ for all integers $n \ge 1$, and deduce that $B_n$ is an even integer when $n$ is an even positive integer.
\end{problem}

\begin{hint}
\begin{itemize}[leftmargin=*]
    \item \textbf{Part (i):} Calculate the sum of the roots ($\alpha + \beta$) and the product of the roots ($\alpha\beta$) to construct a characteristic quadratic equation of the form $x^2 - (\text{sum})x + (\text{product}) = 0$. Multiply this equation by $x^{n-2}$.
    \item \textbf{Part (ii):} Strong induction requires you to establish two base cases ($n=0$ and $n=1$) because the recurrence relation relies on the two previous terms.
    \item \textbf{Part (iii):} Calculate $B_1$ and $B_2$ for your base cases. In your inductive step, express the assumption as $B_k = 4P$ and $B_{k-1} = 4Q$ for integers $P$ and $Q$.
\end{itemize}
\end{hint}

\begin{solution}[Brief]
\textbf{Part (i)}
Let $\alpha = 1+\sqrt{2}$ and $\beta = 1-\sqrt{2}$.
The sum of the roots is $\alpha + \beta = 2$.
The product of the roots is $\alpha\beta = (1+\sqrt{2})(1-\sqrt{2}) = 1 - 2 = -1$.
Therefore, $\alpha$ and $\beta$ are roots of the characteristic equation $x^2 - 2x - 1 = 0$.
Since $\alpha$ is a root, $\alpha^2 = 2\alpha + 1$.
Multiplying by $\alpha^{n-2}$ gives $\alpha^n = 2\alpha^{n-1} + \alpha^{n-2}$.
By identical reasoning, $\beta^n = 2\beta^{n-1} + \beta^{n-2}$.
Since both $A_n = \alpha^n + \beta^n$ and $B_n = \sqrt{2}(\alpha^n - \beta^n)$ are linear combinations of $\alpha^n$ and $\beta^n$, they must both satisfy the recurrence relation:

$$U_n = 2U_{n-1} + U_{n-2}$$

\textbf{Part (ii)}
\textit{Base cases:}
For $n=0$: $A_0 = (1+\sqrt{2})^0 + (1-\sqrt{2})^0 = 1 + 1 = 2$, which is an even integer.
For $n=1$: $A_1 = (1+\sqrt{2})^1 + (1-\sqrt{2})^1 = 2$, which is an even integer.
\textit{Inductive step:}
Assume the statement is true for $n=k$ and $n=k-1$ for some integer $k \ge 1$. That is, assume $A_k$ and $A_{k-1}$ are both even integers.
From the recurrence relation established in (i):

$$A_{k+1} = 2A_k + A_{k-1}$$

Since $A_k$ is an integer, $2A_k$ is an even integer.
Since $A_{k-1}$ is assumed to be an even integer, the sum of two even integers ($2A_k + A_{k-1}$) is also an even integer.
Thus, $A_{k+1}$ is an even integer.
By the principle of strong mathematical induction, $A_n$ is an even integer for all $n \ge 0$.

\textbf{Part (iii)}
\textit{Base cases:}
For $n=1$: $B_1 = \sqrt{2}\left[(1+\sqrt{2}) - (1-\sqrt{2})\right] = \sqrt{2}(2\sqrt{2}) = 4 = 4 \times 1$.
For $n=2$: $B_2 = \sqrt{2}\left[(1+\sqrt{2})^2 - (1-\sqrt{2})^2\right] = \sqrt{2}\left[(3+2\sqrt{2}) - (3-2\sqrt{2})\right] = \sqrt{2}(4\sqrt{2}) = 8 = 4 \times 2$.
Both are multiples of $4$.
\textit{Inductive step:}
Assume true for $n=k$ and $n=k-1$ for some integer $k \ge 2$.
Let $B_k = 4P$ and $B_{k-1} = 4Q$, where $P$ and $Q$ are integers.
Using the recurrence relation:

$$B_{k+1} = 2B_k + B_{k-1}$$
$$B_{k+1} = 2(4P) + 4Q = 4(2P + Q)$$

Since $P$ and $Q$ are integers, $(2P + Q)$ is an integer. Thus, $B_{k+1}$ is a multiple of $4$.
By strong mathematical induction, $B_n$ is a multiple of $4$ for all $n \ge 1$.
Since any multiple of $4$ is inherently an even number, $B_n$ is an even integer for all positive integers $n$, which inherently satisfies the condition for all \textit{even} positive integers $n$.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
    \item \textbf{Binet's Formula Concept:} This problem introduces the inverse logic of Binet's formula (used for Fibonacci numbers). Instead of solving a recurrence relation to find an explicit surd formula, we use the explicit surd formula to find the much simpler integer-based recurrence relation.
    \item \textbf{Strong Induction:} Standard induction only requires the $k$th step to prove the $(k+1)$th step. Because second-order recurrence relations look back \textit{two} steps, strong induction is required, necessitating two base cases and assumptions for both $k$ and $k-1$.
    \item \textbf{Alternative Pathways:} Often in advanced mathematics, combinatorics (Binomial Theorem) and recursive sequences offer two distinct pathways to prove the exact same integer properties.
\end{itemize}
\end{takeaways}

\begin{problem}[The Paradox of Induction]
Prove that for every positive integer $n$, the following inequality holds:
$$ 1 + \frac{1}{2} + \frac{1}{4} + \dots + \frac{1}{2^n} < 2 $$
\end{problem}

\begin{hint}
\begin{itemize}
    \item \textbf{Hint 1:} Try proving this directly using standard induction. Notice that assuming the sum for $k$ is $< 2$ only allows you to prove the sum for $k+1$ is $< 2 + \frac{1}{2^{k+1}}$, which doesn't reach the required target of $< 2$.
    \item \textbf{Hint 2:} Consider the ``Paradox of Induction'': sometimes it is easier to prove a \textit{stronger} statement because it gives you a more powerful inductive hypothesis.
    \item \textbf{Hint 3:} Instead of an inequality, try to find an exact formula for the series. Prove that $1 + \frac{1}{2} + \frac{1}{4} + \dots + \frac{1}{2^n} = 2 - \frac{1}{2^n}$.
\end{itemize}
\end{hint}

\begin{solution}[Brief]
Standard induction fails because the inductive step does not preserve the strict $< 2$ bound without extra information. We formulate a stronger proposition $Q(n)$:
$$ 1 + \frac{1}{2} + \frac{1}{4} + \dots + \frac{1}{2^n} = 2 - \frac{1}{2^n} $$
\textbf{Base Case ($n=1$):} LHS $= 1 + \frac{1}{2} = \frac{3}{2}$. RHS $= 2 - \frac{1}{2} = \frac{3}{2}$. $Q(1)$ is true.\\
\textbf{Inductive Step:} Assume $Q(k)$ is true: $1 + \dots + \frac{1}{2^k} = 2 - \frac{1}{2^k}$.\\
For $n=k+1$, add $\frac{1}{2^{k+1}}$ to the sum:
$$ \text{Sum} = \left(2 - \frac{1}{2^k}\right) + \frac{1}{2^{k+1}} $$
Make common denominators for the fractions:
$$ \text{Sum} = 2 - \frac{2}{2^{k+1}} + \frac{1}{2^{k+1}} = 2 - \frac{1}{2^{k+1}} $$
Thus, $Q(k+1)$ is true. By induction, $Q(n)$ holds for all positive integers $n$. \\
Since $\frac{1}{2^n} > 0$ for all positive integers $n$, it follows directly that $2 - \frac{1}{2^n} < 2$, proving the original statement.
\end{solution}

\begin{takeaways}
\begin{itemize}
    \item \textbf{The Paradox of Induction:} Counter-intuitively, a harder (stronger or more specific) theorem can be easier to prove by induction because the inductive hypothesis you get to assume is equally stronger.
    \item \textbf{Exact formulas beat loose bounds:} If an inequality induction gets stuck, look for an underlying equality that governs the pattern.
\end{itemize}
\end{takeaways}

\begin{problem}[Strong Induction with Rational Functions]
Show that if $x + \frac{1}{x} \in \mathbb{Z}$, then $x^n + \frac{1}{x^n} \in \mathbb{Z}$ for all positive integers $n$.
\end{problem}

\begin{hint}
\begin{itemize}
    \item \textbf{Hint 1:} This problem requires \textbf{Strong Induction}. To prove the statement for $n+1$, you will need to rely on the assumption that it is true for \textit{both} $n$ and $n-1$.
    \item \textbf{Hint 2:} You need two base cases: $n=1$ and $n=2$. Calculate $x^2 + \frac{1}{x^2}$ by squaring $(x + \frac{1}{x})$.
    \item \textbf{Hint 3:} For the inductive step, expand the product $\left(x + \frac{1}{x}\right)\left(x^k + \frac{1}{x^k}\right)$. How does this relate to $x^{k+1} + \frac{1}{x^{k+1}}$?
\end{itemize}
\end{hint}

\begin{solution}[Brief]
Let $P(n)$ be the statement ``$x^n + \frac{1}{x^n} \in \mathbb{Z}$''.\\
\textbf{Base Cases:}\\
$P(1)$ is true by the given hypothesis ($x + \frac{1}{x} \in \mathbb{Z}$).\\
For $P(2)$, we observe that $x^2 + \frac{1}{x^2} = \left(x + \frac{1}{x}\right)^2 - 2$. Since $x + \frac{1}{x}$ is an integer, its square minus 2 is also an integer. Thus $P(2)$ holds.\\
\textbf{Inductive Step (Strong Induction):} \\
Assume $P(k)$ and $P(k-1)$ are true for some integer $k \ge 2$. We examine the term for $k+1$:
$$ x^{k+1} + \frac{1}{x^{k+1}} = \left(x + \frac{1}{x}\right)\left(x^k + \frac{1}{x^k}\right) - \left(x^{k-1} + \frac{1}{x^{k-1}}\right) $$
We know:
\begin{enumerate}
    \item $x + \frac{1}{x} \in \mathbb{Z}$ (hypothesis)
    \item $x^k + \frac{1}{x^k} \in \mathbb{Z}$ (by $P(k)$)
    \item $x^{k-1} + \frac{1}{x^{k-1}} \in \mathbb{Z}$ (by $P(k-1)$)
\end{enumerate}
Because integers are closed under multiplication and subtraction, the entire expression $x^{k+1} + \frac{1}{x^{k+1}}$ must be an integer. Thus $P(k+1)$ is true. By the Strong Principle of Mathematical Induction, $P(n)$ is true for all positive integers $n$.
\end{solution}

\begin{takeaways}
\begin{itemize}
    \item \textbf{Identifying Strong Induction:} Whenever an algebraic sequence or manipulation links the $(k+1)$-th term to multiple preceding terms (not just the $k$-th term), Strong Induction is required.
    \item \textbf{Multiple Base Cases:} If your inductive step looks back two steps (to $k-1$), you must explicitly prove two consecutive base cases to get the chain started.
\end{itemize}
\end{takeaways}
