% Part 2 Problem 08 (Trap / Medium): Euclid's Lemma with "prime" omitted — pedagogical trap
% Technique: Disproof by Counterexample
% Topic: Divisibility / Logic

\begin{problem}[Divisibility of a Product]
Prove that for all positive integers $a, b,$ and $c$, if $a$ divides the product $bc$, then $a$ divides $b$ or $a$ divides $c$.
\end{problem}

\begin{hint}
Attempt to set up a direct proof using the definition of divisibility: $bc = ak$ for some integer $k$. Can you algebraically force $a$ to divide $b$ or $c$?

Wait—is this statement actually true? Try testing it with some small integer values. What happens if $a$ is a composite number (like $4$ or $6$) rather than a prime number?

\textbf{Exam Strategy:} If you ever encounter an exam question that seems fundamentally incorrect or impossible, do not leave the page blank! A blank page guarantees 0 marks. Instead, prove that the statement is false. The absolute best way to completely destroy a ``for all'' statement is to provide a single, concrete \textbf{counterexample}.
\end{hint}

\begin{solution}
The problem statement is strictly \textbf{false}. We will disprove it by providing a counterexample.

To disprove the statement ``if $a$ divides $bc$, then $a|b$ or $a|c$'', we must find specific positive integers $a, b,$ and $c$ such that $a$ divides $bc$, but $a \nmid b$ and $a \nmid c$.

Let $a = 6$, $b = 2$, and $c = 3$.

Check the condition (the premise):
\[
bc = 2 \times 3 = 6
\]
Since $6$ divides $6$, the premise that $a$ divides $bc$ is satisfied.

Check the conclusion:
\begin{itemize}
    \item Does $a$ divide $b$? No, $6$ does not divide $2$.
    \item Does $a$ divide $c$? No, $6$ does not divide $3$.
\end{itemize}

Since both parts of the conclusion are false, the entire statement is invalid.

Therefore, it is not true for all positive integers $a, b, c$ that $a|bc \implies a|b \text{ or } a|c$. $\qed$
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item \textbf{Trust Your Math, Not the Paper:} Examiners are human and occasionally make mistakes. If you leave a flawed question blank, you receive 0 marks. If you mathematically prove that the question is flawed, you demonstrate mastery and will be awarded marks.
\item \textbf{The Power of Counterexamples:} To prove a universal statement (``for all $x$\ldots'') you must prove it algebraically for every number in existence. But to \textit{disprove} it, you only need to find \textbf{one} single case where it fails.
\item \textbf{Euclid's Lemma:} The statement in this problem becomes completely true if we add one crucial condition: $a$ must be a \textit{prime} number. Composite numbers can have their factors split across $b$ and $c$, which is why $a=6$ failed above.
\end{itemize}
\end{takeaways}
