% Part 1, Problem 3: sqrt{23} is irrational - Proof by Contradiction
% Easy

\begin{problem}[Easy]
Prove that $\sqrt{23}$ is irrational.
\end{problem}

\begin{solution}
\textbf{Proof by Contradiction}

\textbf{Step 1: Assume the negation}

Assume, for contradiction, that $\sqrt{23}$ is rational. Then it can be written as:
\[ \sqrt{23} = \frac{p}{q} \]
where $p, q \in \mathbb{Z}$, $q \neq 0$, and $\gcd(p,q) = 1$ (the fraction is in lowest terms).

\textbf{Step 2: Square both sides}

\begin{align*}
    23 &= \frac{p^2}{q^2} \\
    p^2 &= 23q^2
\end{align*}

\textbf{Step 3: Deduce divisibility}

From $p^2 = 23q^2$, we see that $p^2$ is divisible by 23.

Since 23 is prime, if $23 \mid p^2$, then $23 \mid p$ (by fundamental theorem of arithmetic).

\textbf{Step 4: Substitute and derive contradiction}

Since $23 \mid p$, write $p = 23k$ for some integer $k$. Substituting into $p^2 = 23q^2$:
\begin{align*}
    (23k)^2 &= 23q^2 \\
    529k^2 &= 23q^2 \\
    23k^2 &= q^2
\end{align*}

This shows $q^2$ is divisible by 23. Since 23 is prime, $23 \mid q$.

\textbf{Step 5: Establish contradiction}

We have shown that both $p$ and $q$ are divisible by 23.

This contradicts our assumption that $\gcd(p,q) = 1$.

\textbf{Conclusion}

Therefore, $\sqrt{23}$ cannot be expressed as a fraction of integers.

Hence, $\sqrt{23}$ is irrational.

\hfill $\square$
\end{solution}

\begin{takeaways}
\begin{itemize}
\item \textbf{Standard Irrationality Template:} Assume rational form $\frac{p}{q}$ in lowest terms, square to get $p^2 = nq^2$, show both $p$ and $q$ divisible by prime factor, contradict lowest terms assumption
\item \textbf{Prime Divisibility:} If prime $p$ divides $a^2$, then $p$ divides $a$ (key property from fundamental theorem of arithmetic)
\item \textbf{Lowest Terms:} Always start with $\gcd(p,q) = 1$ to enable the contradiction via common factors
\item \textbf{Generalization:} This method proves $\sqrt{p}$ irrational for any prime $p$
\end{itemize}
\end{takeaways}
