% Part 2 Problem 03 (Easy): Irrational^Irrational can be Rational
% Source: samples/12.tex
% Technique: Proof by Cases (Non-constructive)
% Topic: Irrationality/Existence

\begin{problem}[Easy]
Prove that an irrational number raised to the power of an irrational number can be rational, by considering $\sqrt{2}^{\sqrt{2}}$. You may assume $\sqrt{2}$ is irrational.
\end{problem}

\begin{hint}
Consider $\alpha = \sqrt{2}^{\sqrt{2}}$. By the Law of Excluded Middle, $\alpha$ is either rational or irrational. Examine both cases:
\begin{itemize}
\item If $\alpha$ is rational, you're done immediately.
\item If $\alpha$ is irrational, compute $\alpha^{\sqrt{2}}$ and simplify.
\end{itemize}

This is a \textbf{non-constructive} existence proof---you prove something exists without determining which case actually holds!
\end{hint}

\begin{solutionsketch}
\textbf{Proof by Cases:}

Let $\alpha = \sqrt{2}^{\sqrt{2}}$. We consider two exhaustive cases:

\vspace{0.3cm}

\textbf{Case 1:} If $\alpha = \sqrt{2}^{\sqrt{2}}$ is rational, then we have found irrational base $\sqrt{2}$ and irrational exponent $\sqrt{2}$ such that $\sqrt{2}^{\sqrt{2}}$ is rational. Done.

\vspace{0.3cm}

\textbf{Case 2:} If $\alpha = \sqrt{2}^{\sqrt{2}}$ is irrational, consider:
\begin{align*}
\alpha^{\sqrt{2}} &= \left(\sqrt{2}^{\sqrt{2}}\right)^{\sqrt{2}} \\
&= \sqrt{2}^{\sqrt{2} \cdot \sqrt{2}} \quad \text{(exponent law)} \\
&= \sqrt{2}^{2} \\
&= 2
\end{align*}

Since $2$ is rational and both $\alpha$ (irrational by case assumption) and $\sqrt{2}$ (given as irrational) are irrational, we have found an example.

\vspace{0.3cm}

\textbf{Conclusion:} In either case, there exist irrational numbers $a$ and $b$ such that $a^b$ is rational. $\qed$

\vspace{0.3cm}

\textit{Note: This proof doesn't tell us whether $\sqrt{2}^{\sqrt{2}}$ is actually rational or irrational---and we don't need to know! This is the beauty of non-constructive existence proofs.}
\end{solutionsketch}
