% Part 1, Problem 21: Sum of surds irrational - Proof by Contradiction
% Hard (Question 16 style)

\begin{problem}[Irrationality of $\sqrt{2} + \sqrt{3} + \sqrt{5}$]
Prove by contradiction that the real number $x = \sqrt{2} + \sqrt{3} + \sqrt{5}$ is irrational.
\end{problem}

\begin{hint}
\begin{itemize}
    \item \textbf{The Setup:} Begin with the standard assumption for contradiction: let $x \in \mathbb{Q}$.
    \item \textbf{Strategic Isolation:} Do not square the expression as it is. Squaring a trinomial creates a mess of cross-terms. Isolate one of the surds (e.g., move $\sqrt{5}$ to the left) before squaring.
    \item \textbf{The Double Square:} Squaring once will leave you with a $\sqrt{5}$ term on the left and a $\sqrt{6}$ term on the right. Isolate the $\sqrt{6}$ term and square a second time to eliminate it.
    \item \textbf{The Contradiction Engine:} Rearrange your final equation to make $\sqrt{5}$ the subject. Rely on the closure property of rational numbers (addition, subtraction, multiplication, and non-zero division of rationals always yield a rational) to force the contradiction.
\end{itemize}
\end{hint}

\begin{solution}
\textbf{Proof by Contradiction}

\textbf{Step 1: Assume the negation}

Assume, for the purpose of contradiction, that $x$ is a rational number ($x \in \mathbb{Q}$).

\textbf{Step 2: Isolate one surd}

Rearrange the equation to isolate one surd:
\[
x - \sqrt{5} = \sqrt{2} + \sqrt{3}
\]

\textbf{Step 3: Square both sides}

\begin{align*}
(x - \sqrt{5})^2 &= (\sqrt{2} + \sqrt{3})^2 \\
x^2 - 2x\sqrt{5} + 5 &= 2 + 2\sqrt{6} + 3 \\
x^2 - 2x\sqrt{5} &= 2\sqrt{6}
\end{align*}

\textbf{Step 4: Square a second time}

\begin{align*}
(x^2 - 2x\sqrt{5})^2 &= (2\sqrt{6})^2 \\
x^4 - 4x^3\sqrt{5} + 20x^2 &= 24
\end{align*}

\textbf{Step 5: Make $\sqrt{5}$ the subject}

\begin{align*}
4x^3\sqrt{5} &= x^4 + 20x^2 - 24 \\
\sqrt{5} &= \frac{x^4 + 20x^2 - 24}{4x^3}
\end{align*}

Since $x > 0$, division by $4x^3$ is valid.

\textbf{Step 6: Derive the contradiction}

Because the rational numbers are closed under addition, subtraction, multiplication, and non-zero division, the right-hand side evaluates to a rational number. However, the left-hand side, $\sqrt{5}$, is known to be irrational.

\textbf{Conclusion}

A rational number cannot equal an irrational number. This is a contradiction. Therefore, the initial assumption is false, and $x = \sqrt{2} + \sqrt{3} + \sqrt{5}$ must be irrational.

\hfill $\square$
\end{solution}

\begin{takeaways}
\begin{itemize}
\item \textbf{Algebraic Foresight:} Recognising that squaring $(a + b + c)^2$ yields too many irrational cross-terms is the key separator for top-band students. Isolate one surd before squaring.
\item \textbf{The Power of Closure:} The entire proof hinges on knowing that manipulating rational $x$ through integer powers and coefficients guarantees a rational output on the right-hand side.
\item \textbf{Structured Argumentation:} The proof demands a clear logical flow from the initial assumption to an explicit statement of why the final line constitutes a mathematical contradiction.
\item \textbf{Double Squaring Technique:} This method extends the squaring strategy used in simpler surd proofs (such as comparing $\sqrt{3} + \sqrt{5}$ with $\sqrt{11}$) to irrationality arguments involving multiple surds.
\item You are encouraged to prove that $\sqrt{2} + \sqrt{3} + \sqrt{6}$ is irrational using a different method, which can be smarter thant the technique used here. Note that $6 = 2 \cdot 3$, which may be useful.
\end{itemize}
\end{takeaways}
