\documentclass[12pt]{article}

% Shared HSC style packages (see ../HSC-Common/).
\makeatletter
\def\input@path{{../HSC-Common/styles/}{./}}
\makeatother
\newcommand{\HSCBookletHeaderTitle}{Combinatorics Practice}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\newtheorem*{theorem}{Theorem}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
% Links without visible rectangles (TOC and cross-refs remain clickable).
\usepackage[hidelinks]{hyperref}
\hypersetup{hidelinks}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{calc}

\makeatletter
\newcommand\binomialCoefficient[2]{%
  \c@pgf@counta=#1% n
  \c@pgf@countb=#2% k
  \c@pgf@countc=\c@pgf@counta%
  \advance\c@pgf@countc by-\c@pgf@countb%
  \ifnum\c@pgf@countb>\c@pgf@countc%
    \c@pgf@countb=\c@pgf@countc%
  \fi%
  \c@pgf@countc=1%
  \c@pgf@countd=0%
  \pgfmathloop
    \ifnum\c@pgf@countd<\c@pgf@countb%
    \multiply\c@pgf@countc by\c@pgf@counta%
    \advance\c@pgf@counta by-1%
    \advance\c@pgf@countd by1%
    \divide\c@pgf@countc by\c@pgf@countd%
  \repeatpgfmathloop%
  \the\c@pgf@countc%
}
\makeatother

\usepackage{dl101-colors}
\usepackage{dl101-hints}
\usepackage[takeawaysitemize]{dl101-hsc-problems}
\usepackage{dl101-layout}

\providecolor{warmstone}{RGB}{208,184,168}

\newtcolorbox[auto counter,number within=section]{remarkbox}[1][]{
  title={\bfseries Remark~\thetcbcounter},
  enhanced,
  drop shadow={black!20!white},
  coltitle=bookwhite,
  colback=softivory,
  colframe=warmstone,
  boxrule=1.5pt,
  arc=4pt,
  left=12pt,right=12pt,top=12pt,bottom=12pt,
  attach boxed title to top left={xshift=1.5em,yshift=-\tcboxedtitleheight/2},
  boxed title style={size=small,colback=warmstone,coltitle=white},
  fonttitle=\bfseries,
  #1
}

\newenvironment{remark}[1][]{%
  \begin{remarkbox}[title={\bfseries Remark~\thetcbcounter\ifx\relax#1\relax\else\ (#1)\fi}]%
}{%
  \end{remarkbox}%
}

\author{Vu Hung Nguyen}
\title{HSC Mathematics Extension 1: Combinatorics Practice}
\date{}

\begin{document}

\input{assets/matte.tex}

% QR Code Page (Page 2)
\newpage
\input{../HSC-Common/assets/qrcode-page.tex}


\begingroup\small\noindent This work is licensed under CC BY 4.0, see the LICENSE file on the github page for more info.\par\endgroup

\newpage

\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
\itshape
Mathematics may be defined as the economy of counting. There is no problem in the whole of\\
mathematics which cannot be solved by direct counting.

\vspace{1em}
\normalfont Ernst Mach
\end{center}
\vspace*{\fill}

\newpage

\tableofcontents
\newpage

\section{Introduction}

\subsection{Project Overview}
This booklet is a growing collection of combinatorics problems for NSW HSC Mathematics Extension~1 students. It now covers a wider range of syllabus ideas including permutations, combinations, Pascal's triangle, circular arrangements, committee selection, simple probability, expected value, the pigeonhole principle, geometric counting, and proof-style binomial identities.

\subsection{Target Audience}
This resource is written for students who want extra practice with combinatorics in a clear, classroom-friendly format. The solutions are intentionally brief, but each one points out the key counting idea behind the answer.

\subsection{How to Use This Booklet}
Read the short summary of the core formulas first, then attempt each problem before reading the solution. As you work, decide whether each question is about selection or arrangement, look for structure such as gaps in circles, repeated objects, complementary counting, or balanced groups, and use the takeaways to build pattern recognition for future HSC questions.

\subsection{Extension 1 Knowledge Needed}
Most problems in this booklet can be solved using standard HSC Mathematics Extension~1 ideas rather than advanced university mathematics. The key background knowledge is:

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{|p{0.32\textwidth}|p{0.62\textwidth}|}
\hline
\textbf{Title} & \textbf{Explanation} \\
\hline
\textbf{Factorials and notation} & Understanding $n!$, permutation notation ${}^nP_r$, and combination notation $\binom{n}{r}$. For positive integer $n$, $n!=n(n-1)(n-2)\cdots 2\cdot 1$. Also, $0!=1$, $1!=1$, $2!=2$, $3!=6$, $4!=24$, $5!=120$. Here $\binom{n}{r}$ means the number of ways to choose $r$ objects from $n$ objects (order does not matter). \\
\hline
\textbf{The multiplication principle} & Breaking a problem into stages and multiplying the number of choices at each stage. \\
\hline
\textbf{Order versus selection} & Recognising when order matters (permutations) and when it does not (combinations). \\
\hline
\textbf{Repeated objects} & Dividing by repeated factorials when arranging words such as \texttt{COOKED} or \texttt{CONDOBOLIN}. \\
\hline
\textbf{Circular permutations} & Using $(n-1)!$ for arrangements around a circle and handling restrictions by fixing seats, using blocks, or using gaps. \\
\hline
\textbf{Complementary counting} & Counting all cases first, then subtracting the unwanted ones. \\
\hline
\textbf{Casework and restrictions} & Splitting problems into clean cases such as fixed numbers of girls and boys, men and women, or chosen leaders. \\
\hline
\textbf{Pascal's triangle and binomial coefficients} & Using identities such as $\binom{n}{r}=\binom{n-1}{r-1}+\binom{n-1}{r}$ and the symmetry rule $\binom{n}{r}=\binom{n}{n-r}$. \\
\hline
\end{tabular}
\end{center}

\subsection{Core Formulas and Ideas}

\textbf{Fundamental Counting Principle.} If a task has stages with $a$, $b$, and $c$ choices, then the total number of outcomes is
\[
a \times b \times c.
\]
\textit{Short example.} A meal has $2$ drink choices, $3$ main choices, and $2$ dessert choices, so the number of meals is
\[
2\times 3\times 2=12.
\]

\textbf{Generalized counting principle.} If $k$ experiments are performed, and experiment $i$ has $n_i$ outcomes, then the total number of outcomes is
\[
n_1\cdot n_2\cdot \dots \cdot n_k
\]
for all $k$ experiments together.
\textit{Short example.} A code has $4$ characters, each chosen from $10$ digits. The total number of codes is
\[
10\cdot 10\cdot 10\cdot 10=10^4=10000.
\]

\textbf{What is a set?} A \textbf{set} is a collection of distinct objects. In a set, order does not matter, and repeated copies are not listed separately. For example,
\[
\{1,2,3\}=\{3,2,1\},
\]
and writing $\{1,1,2\}$ does not create a new $3$-element set; as a set, it is just $\{1,2\}$.
\textit{Short example.} The selections $\{\text{red, blue}\}$ and $\{\text{blue, red}\}$ are the same set.

In counting problems, this usually means:
\begin{itemize}[leftmargin=*]
  \item treat objects as distinct unless the question says some are identical
  \item do not assume repetition is allowed
  \item if repetition is allowed, the question should state it.
\end{itemize}

\textbf{Permutations versus combinations.} A \textbf{permutation} is an arrangement, so order matters. A \textbf{combination} is a selection, so order does not matter.

\textit{Short example.} Suppose $3$ students are Alice, Ben, and Chloe.
\begin{itemize}[leftmargin=*]
  \item Captain and vice-captain: order matters, so this is a permutation:
  \[
  {}^3P_2 = 3\times 2 = 6.
  \]
  ``Alice then Ben'' is different from ``Ben then Alice''.

  \item Team of $2$: order does not matter, so this is a combination:
  \[
  \binom{3}{2}=3.
  \]
  The team $\{\text{Alice, Ben}\}$ is the same team as $\{\text{Ben, Alice}\}$.
\end{itemize}

The following formulas are useful to keep together as a quick reference:

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|p{0.34\textwidth}|p{0.56\textwidth}|}
\hline
\textbf{Name of formula} & \textbf{Formula} \\
\hline
\textbf{Permutations without repetition} & ${}^{n}P_r = \dfrac{n!}{(n-r)!}$ \\
\hline
\textbf{Combinations} & $\binom{n}{r} = \dfrac{n!}{r!(n-r)!}$ \\
\hline
\textbf{Multinomial coefficients} & If $n=a+b+c+\cdots$, then the number of distinct arrangements is $\dfrac{n!}{a!\,b!\,c!\cdots}$. \\
\hline
\textbf{Multinomial expansion note} & In $(x_1+x_2+\cdots+x_k)^m$, the coefficient of $x_1^{a_1}x_2^{a_2}\cdots x_k^{a_k}$ where $a_1+a_2+\cdots+a_k=m$ is $\dfrac{m!}{a_1!a_2!\cdots a_k!}$. \\
\hline
\textbf{Circular permutations} & $(n-1)!$ \\
\hline
\textbf{Pascal's triangle identity} & $\binom{n}{r}=\binom{n-1}{r-1}+\binom{n-1}{r}$ \\
\hline
\textbf{Binomial symmetry} & $\binom{n}{r}=\binom{n}{n-r}$ \\
\hline
\end{tabular}
\end{center}

\subsection{Pascal's Triangle}
Pascal's triangle is a triangular arrangement of binomial coefficients. 
It is useful both for quick coefficient reading and for proving combinatorial identities.

% For $n=0,1,2,\dots,9$ (10 rows):
% , Pascal's triangle can be shown with stacked hexagons:
\begin{center}
\def\HexR{0.43cm}
\begin{tikzpicture}[line width=0.75pt]
  \foreach \k in {0,...,9}{
    \begin{scope}[shift={(-60:{sqrt(3)*\HexR*\k})}]
      \pgfmathtruncatemacro\ystart{9-\k}
      \foreach \n in {0,...,\ystart}{
        \pgfmathtruncatemacro\row{\n+\k}
        \def\hexbase{softivory}

        % Use the shared booklet palette for each diagonal.
        \ifnum\k=0\relax\def\hexbase{warmstone}\fi
        \ifnum\k=1\relax\def\hexbase{bookred}\fi
        \ifnum\k=2\relax\def\hexbase{lawpurple}\fi
        \ifnum\k=3\relax\def\hexbase{bookpurple}\fi
        \ifnum\k=4\relax\def\hexbase{deepcharcoal}\fi
        \ifnum\k=5\relax\def\hexbase{warmstone}\fi
        \ifnum\k=6\relax\def\hexbase{bookred}\fi
        \ifnum\k=7\relax\def\hexbase{lawpurple}\fi
        \ifnum\k=8\relax\def\hexbase{bookpurple}\fi
        \ifnum\k=9\relax\def\hexbase{deepcharcoal}\fi

        \begin{scope}[shift={(-120:{sqrt(3)*\HexR*\n})}]
          \draw[draw=deepcharcoal!80,
                top color=\hexbase!18!bookwhite,
                bottom color=\hexbase!40!softivory]
            (30:\HexR)
            \foreach \x in {90,150,...,330} { -- (\x:\HexR) }
            -- cycle
            node at (0,0) {\scriptsize\bfseries\color{deepcharcoal}\binomialCoefficient{\row}{\k}};
        \end{scope}
      }
    \end{scope}
  }
\end{tikzpicture}
\end{center}

% \textbf{Three facts to remember:}

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{|p{0.24\textwidth}|p{0.68\textwidth}|}
\hline
\textbf{Pascal rule} &
$\binom{n}{r}=\binom{n-1}{r-1}+\binom{n-1}{r}$. Each interior entry is the sum of the two entries above it. \\
\hline
\textbf{Symmetry} &
$\binom{n}{r}=\binom{n}{n-r}$. Each row reads the same from left to right and from right to left. \\
\hline
\textbf{Row sum} &
$\binom{n}{0}+\binom{n}{1}+\cdots+\binom{n}{n}=2^n$. The sum of the entries in row $n$ is a power of $2$. \\
\hline
\end{tabular}
\end{center}

\subsection{The Binomial Theorem}
For a positive integer $n$,
\[
(x+y)^n=\binom{n}{0}x^n+\binom{n}{1}x^{n-1}y+\binom{n}{2}x^{n-2}y^2+\cdots+\binom{n}{n}y^n.
\]

This theorem tells us how to expand powers such as $(x+y)^n$. The coefficients come directly from row $n$ of Pascal's triangle.

\textbf{Worked example.} Expand $(x+2)^4$.

Using row $4$ of Pascal's triangle, the coefficients are $1,4,6,4,1$:
\begin{align*}
(x+2)^4
&= \binom{4}{0}x^4 + \binom{4}{1}x^3(2) + \binom{4}{2}x^2(2^2) + \binom{4}{3}x(2^3) + \binom{4}{4}(2^4) \\
&= x^4 + 4x^3(2) + 6x^2(4) + 4x(8) + 16 \\
&= x^4 + 8x^3 + 24x^2 + 32x + 16.
\end{align*}

This is especially useful later in the booklet when proving identities by comparing coefficients of matching powers of $x$.

\textbf{Expected value.} If a random variable takes values $x_1,x_2,\dots$ with probabilities $p_1,p_2,\dots$, then
\[
E(X)=x_1p_1+x_2p_2+\cdots.
\]

\subsection{Binomial Probability}
Binomial probability is used when the same trial is repeated a fixed number of times and each trial has only two possible outcomes, often called \textit{success} and \textit{failure}.

\textbf{The main features are:}
\begin{itemize}[leftmargin=*]
  \item there are only two possible outcomes on each trial
  \item the probability of success is the same on every trial
  \item the number of trials is fixed
  \item the trials are independent
\end{itemize}

If $X$ is the number of successes in $n$ trials, and the probability of success on each trial is $p$, then
\[
P(X=r)=\binom{n}{r}p^r(1-p)^{n-r}.
\]

Here $\binom{n}{r}$ counts the number of ways to arrange the $r$ successes among the $n$ trials.

\textit{Example.} A green counter is selected with probability $0.2$ on each draw, and the draw is repeated $5$ times with replacement. The probability of getting exactly $2$ green counters is
\[
P(X=2)=\binom{5}{2}(0.2)^2(0.8)^3.
\]

This idea appears in repeated-draw questions later in the booklet.

\subsection{The Pigeonhole Principle and Remainders}
\begin{remark}[Syllabus Note]
From 2026, the Pigeonhole Principle is excluded from NESA and will not be tested in the HSC examinations for the Class of 2026 (or any cohort after). However, this is an important concept for your proofs, counting, and further study.
\end{remark}

Suppose that $n$ objects are placed into $d$ boxes. If we divide $n$ by $d$, we can write
\[
n=qd+r,
\qquad \text{where } 0\le r<d.
\]

This helps us read the pigeonhole principle more precisely:
\begin{itemize}[leftmargin=*]
  \item if $r=0$, then at least one box contains at least $q$ objects
  \item if $r>0$, then at least one box contains at least $q+1$ objects.
\end{itemize}

In particular, the smallest number of objects that guarantees at least one box contains at least $q+1$ objects is
\[
qd+1.
\]

This point of view is especially useful in ``must happen'' questions, where the boxes may be teams, remainders after division, topic choices, or vote totals.

\subsection{Problem-Solving Habits for This Booklet}
\begin{itemize}[leftmargin=*]
  \item Ask first: \emph{Does order matter?}
  \item Check whether repetition is allowed or forbidden.
  \item In circle problems, decide whether to use a fixed reference point, a block, or a gap method.
  \item In identity questions, look for Pascal's triangle, symmetry, coefficient matching, or telescoping.
  \item In probability questions, decide whether the experiment is with replacement or without replacement.
  \item In ``must happen'' questions, test whether the pigeonhole principle or a contradiction argument applies.
\end{itemize}

\begin{remark}
The most common combinatorics mistake is using a permutation when the problem is actually a combination, or vice versa. A good first question is always: \emph{Does order matter?}
\end{remark}

\begin{remark}[How fast counting grows]
A $19\times 19$ Go board has $361$ points. If each point is allowed to be black, white, or empty, then a rough count of possible board states is
\[
3^{361}\approx 10^{172}.
\]
Even after restricting to legal positions, the number is still about $10^{170}$, which is unimaginably large. This is one of the best reminders that counting problems can grow far beyond ordinary intuition.
\end{remark}

\section{Part 1: Worked Problems}

\subsection{Basic}
\input{problems/part1-basic.tex}

\subsection{Medium}
\input{problems/part1-medium.tex}

\subsection{Advanced}
\input{problems/part1-advanced.tex}

\section{Part 2: Practice and Challenge Bank}

\subsection{Warm-Up Drills}
\input{problems/part2-basic.tex}

\subsection{Stretch Problems}
\input{problems/part2-medium.tex}

\subsection{Challenge Corner}
\input{problems/part2-advanced.tex}

\section{Conclusion}

Combinatorics rewards careful reading as much as technical skill. Many HSC questions become much easier once you decide whether the task is an arrangement, a selection, or a probability question built from counting. As this booklet grows, it can become a compact reference for the most common combinatorics patterns students meet in Extension~1.

\vspace{2em}
\noindent\textbf{Contact Information:}\\[0.5em]
LinkedIn: \href{https://www.linkedin.com/in/nguyenvuhung/}{https://www.linkedin.com/in/nguyenvuhung/}\\[0.3em]
GitHub: \href{https://github.com/vuhung16au/}{https://github.com/vuhung16au/}\\[0.3em]
Repository: \href{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Combinatorics}{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Combinatorics}

\newpage
\appendix
\section{Appendices}

\subsection{Gamma and Beta Functions}
Although this booklet focuses on HSC Mathematics Extension~1, it is worth knowing that factorials and combinations have important extensions in higher mathematics.

\textbf{Gamma function.} The Gamma function is defined for suitable real or complex inputs by
\[
\Gamma(z)=\int_0^\infty t^{z-1}e^{-t}\,dt.
\]
It satisfies the recurrence
\[
\Gamma(z+1)=z\Gamma(z),
\]
which mirrors the factorial rule $n!=n(n-1)!$. In fact,
\[
\Gamma(n)=(n-1)!,
\qquad\text{so}\qquad
n!=\Gamma(n+1).
\]
This means the Gamma function is the natural continuous extension of the factorial function.

\textbf{Beta function.} The Beta function is closely related to Gamma and is defined by
\[
B(x,y)=\frac{\Gamma(x)\Gamma(y)}{\Gamma(x+y)}.
\]
Since combinations are built from factorials,
\[
\binom{n}{r}=\frac{n!}{r!(n-r)!}
=\frac{\Gamma(n+1)}{\Gamma(r+1)\Gamma(n-r+1)}.
\]
So Gamma and Beta functions provide a continuous version of the factorial and binomial-coefficient ideas that appear throughout combinatorics.

\textit{Why this matters.} These functions form a bridge between discrete counting and continuous mathematics. They are useful in advanced probability, statistics, asymptotics, and areas such as machine learning.

\subsection{Stars and Bars}
Stars and bars is a method for counting integer solutions to equations such as
\[
x_1+x_2+\cdots+x_k=n.
\]

The idea is to represent the total $n$ by $n$ stars and use $k-1$ bars to divide the stars into $k$ groups.

\begin{center}
\begin{tikzpicture}[baseline=(current bounding box.center),scale=1]
  \foreach \x in {0,0.45,0.9,1.35,1.8,2.25,2.7} {
    \node at (\x,0) {$*$};
  }
  \draw[thick] (0.65,-0.35) -- (0.65,0.35);
  \draw[thick] (2.05,-0.35) -- (2.05,0.35);
  \node at (1.35,-0.9) {$2\mid 3\mid 2$};
\end{tikzpicture}
\end{center}

If the variables are required to be positive, then each group must contain at least one star, so the number of solutions is
\[
\binom{n-1}{k-1}.
\]

If the variables are allowed to be zero, then empty groups are allowed, so the number of solutions is
\[
\binom{n+k-1}{k-1}.
\]

Stars and bars is a standard combinatorial technique for turning algebraic equations into counting problems. It is directly connected to the two Part~2 challenge problems on positive and non-negative integer solutions.

\textit{Exam note.} Recent HSC Mathematics Extension~1 exams do not usually ask this type of problem directly. However, it is still very good to know, because it is a powerful counting technique to keep as a mathematical ``weapon'' for olympiad-style problems, extension work, and future university study.

\subsection{Algorithmic Counting and Complexity}
Algorithmic counting is the study of methods for counting objects systematically. Complexity asks how difficult a counting problem becomes as the size of the problem grows.

Some counting problems have simple closed formulas, such as
\[
\binom{n}{r},\qquad n!,\qquad 2^n.
\]
However, many counting problems grow extremely quickly and cannot be solved efficiently by checking every possibility one by one.

For example:
\[
\begin{array}{c|c|c}
n & n! & 2^n \\ \hline
5 & 120 & 32 \\
10 & 3628800 & 1024 \\
20 & 2432902008176640000 & 1048576
\end{array}
\]

So even a simple arrangement problem can become enormous very quickly.

This is one reason why combinatorics is important: good counting methods, symmetry arguments, recurrences, and identities help us avoid brute force.

\textit{Syllabus note.} This topic is outside the HSC Mathematics Extension~1 syllabus. However, it is very useful for higher education, especially in areas such as computer science, discrete mathematics, and algorithm design.

\subsection{The Inclusion-Exclusion Principle}
The inclusion-exclusion principle is used when counting objects that may belong to overlapping sets.

For two sets $A$ and $B$,
\[
|A\cup B|=|A|+|B|-|A\cap B|.
\]

For three sets $A$, $B$, and $C$,
\[
|A\cup B\cup C|
=
|A|+|B|+|C|
-|A\cap B|-|A\cap C|-|B\cap C|
+|A\cap B\cap C|.
\]

The idea is simple: when we add the sizes of sets, elements in overlaps are counted more than once, so we must correct for this.

Also note that this is an important topic when you learn probability.

This principle is very useful in restricted counting problems, probability, and Venn diagram questions.

\begin{center}
\begin{tikzpicture}
  \begin{scope}[blend group=soft light]
    \fill[red!40!white]   (270:1.1) circle (1.8);
    \fill[blue!40!white]  (30:1.1)  circle (1.8);
    \fill[green!40!white] (150:1.1) circle (1.8);
  \end{scope}
  \draw (270:1.1) circle (1.8);
  \draw (30:1.1)  circle (1.8);
  \draw (150:1.1) circle (1.8);
  % Set labels (outside)
  \node at (270:3.2)  {$B$};
  \node at (30:3.2)   {$C$};
  \node at (150:3.2)  {$A$};
  % Pairwise intersections
  \node[font=\small] at ($(270:1.1)!0.5!(30:1.1)+(0,-0.55)$)  {$B\cap C$};
  \node[font=\small] at ($(270:1.1)!0.5!(150:1.1)+(-0.55,0)$) {$A\cap B$};
  \node[font=\small] at ($(30:1.1)!0.5!(150:1.1)+(0.55,0)$)   {$A\cap C$};
  % Triple intersection
  \node[font=\small] at (0,0) {$A\cap B\cap C$};
\end{tikzpicture}
\end{center}

For four sets $A, B, C,$ and $D$, we have:

\begin{align*}
& |A \cup B \cup C \cup D| =
\sum |A| - \sum |A \cap B| + \sum |A \cap B \cap C| - |A \cap B \cap C \cap D|\\
& = |A| + |B| + |C| + |D| -(|A \cap B| + |A \cap C| + |A \cap D| + |B \cap C| + |B \cap D| + |C \cap D|)\\
& \quad +( |A \cap B \cap C| + |A \cap B \cap D| + |A \cap C \cap D| + |B \cap C \cap D|) - |A \cap B \cap C \cap D|.
\end{align*}
  
\subsubsection*{General Case: Theorem}

If $A_1, A_2, \dots, A_n$ are finite sets, then the cardinality of their union is given by:

\begin{align*}
|A_1 \cup A_2 \cup \dots \cup A_n| &= \sum_{i=1}^n |A_i| - \sum_{1 \le i < j \le n} |A_i \cap A_j| + \dots \\
&\quad + (-1)^{k-1} \sum_{1 \le i_1 < i_2 < \dots < i_k \le n} |A_{i_1} \cap A_{i_2} \dots \cap A_{i_k}| + \dots + (-1)^{n-1} |A_1 \cap \dots \cap A_n|
\end{align*}

\subsubsection*{Proof by Mathematical Induction}

\textbf{Step 1: Proposition Definition}\\
Let $P(n)$ be the statement: "For any sets $A_1 \dots A_n$, we have
\[
|A_1 \cup \dots \cup A_n| = \sum_{i} |A_i| - \sum_{i<j} |A_i \cap A_j| + \dots \pm |A_1 \cap A_2 \cap \dots \cap A_n|
\]
"

\textbf{Step 2: Base Cases}
\begin{itemize}
    \item $P(1)$ is trivially true, as $|A_1| = |A_1|$.
    \item $P(2)$ is also true, since by basic properties of sets:
\[
|A_1 \cup A_2| = |A_1| + |A_2| - |A_1 \cap A_2|
\]
\end{itemize}

\textbf{Step 3: Inductive Hypothesis and Step}\\
Assume now that the result holds for all positive integers less than or equal to some integer $n \ge 2$. We must show that $P(n+1)$ holds for $n+1$ sets.

Given sets $A_1 \dots A_{n+1}$, let us define a new collection of intersections:
Let $B_i = A_i \cap A_{n+1}$ for $1 \le i \le n$.

From this definition, observe how the intersections of our new $B$ sets correspond to higher-order intersections of our original $A$ sets:
\begin{itemize}
    \item $B_i \cap B_j = A_i \cap A_j \cap A_{n+1}$
    \item $B_i \cap B_j \cap B_k = A_i \cap A_j \cap A_k \cap A_{n+1}$
\end{itemize}

By treating the first $n$ sets as a single unioned set, and $A_{n+1}$ as the second set, we can apply the $P(2)$ base case:
\[
|A_1 \cup A_2 \cup \dots \cup A_{n+1}| = |(A_1 \cup \dots \cup A_n) \cup A_{n+1}|
\]
\[
|A_1 \cup A_2 \cup \dots \cup A_{n+1}| = |A_1 \cup \dots \cup A_n| + |A_{n+1}| - |(A_1 \cup \dots \cup A_n) \cap A_{n+1}|
\]

Using the distributive law of sets, the subtracted term becomes:
\[
(A_1 \cup \dots \cup A_n) \cap A_{n+1} = B_1 \cup \dots \cup B_n
\]

Substitute this back into our equation:
\[
|A_1 \cup A_2 \cup \dots \cup A_{n+1}| = |A_1 \cup \dots \cup A_n| + |A_{n+1}| - |B_1 \cup \dots \cup B_n|
\]

Now, apply the induction hypothesis to both $|A_1 \cup \dots \cup A_n|$ (which has $n$ sets) and $|B_1 \cup \dots \cup B_n|$ (which also has $n$ sets):
\[
= \sum_{i \le n} |A_i| - \sum_{i < j \le n} |A_i \cap A_j| + \dots + |A_{n+1}| - \sum_{i \le n} |B_i| + \sum_{i < j \le n} |B_i \cap B_j| - \dots \quad \text{(Equation 1)}
\]

\textbf{Step 4: Algebraic Recombination}\\
Notice that the sum of the single $B$ sets perfectly supplements the pairwise intersections of the $A$ sets that are strictly up to $n$:
\[
\sum_{i \le n} |B_i| = \sum_{i \le n} |A_i \cap A_{n+1}|
\]

Hence, adding this to the existing pairwise sum gives us the complete set of pairwise intersections up to $n+1$:
\[
\sum_{i < j \le n} |A_i \cap A_j| + \sum_{i \le n} |B_i| = \sum_{i < j \le n+1} |A_i \cap A_j|
\]

This same supplementary logic applies perfectly for all higher-order intersections. Substituting these combined terms back into Equation (1), we obtain:
\[
|A_1 \cup A_2 \dots \cup A_{n+1}| = \sum_{i \le n+1} |A_i| - \sum_{i < j \le n+1} |A_i \cap A_j| + \dots
\]

This shows that $P(n+1)$ holds. By the principle of mathematical induction, the Inclusion-Exclusion Principle is true for any finite number of sets $n$.

\subsection{Concepts: Injective, Surjective, and Bijective}

Understanding these concepts is incredibly important for combinatorics, particularly when answering ``how many'' or proving that two sets have the same number of elements without actually counting them one by one. In combinatorics, formalizing a counting process as a function provides powerful insights:
\begin{itemize}[leftmargin=*]
  \item \textbf{Injective (One-to-one) tracking:} If you can construct an injective mapping from set $X$ to set $Y$, you instantly prove that $|X| \le |Y|$ (the size of $X$ is less than or equal to $Y$). If you try to map more elements into fewer targets, an injection is impossible---this is the formal basis of the \textbf{Pigeonhole Principle}. When counting arrangements, injections describe ways to assign objects to distinct slots without repetition (like ${}^nP_r$).
  \item \textbf{Surjective (Onto) coverage:} If you can construct a surjective mapping from set $X$ to set $Y$, you prove that $|X| \ge |Y|$. In counting, surjections are used when we want to count the number of ways to distribute objects into boxes where \emph{no box is left empty}. 
  \item \textbf{Bijective (Perfect pairing) counting:} If you can construct a bijection between set $X$ and set $Y$, you prove that $|X| = |Y|$. This is called \textbf{Bijective Proof} and it is one of the most elegant techniques in combinatorics. If set $X$ is very hard to count, but you can link it perfectly to set $Y$ which is easy to count, then you have solved the problem! (For instance, stars and bars is a bijection between dividing items and placing dividers).
\end{itemize}

Here is the complete breakdown of how these mappings work mathematically.

\begin{remark}[Syllabus Note]
These concepts are outside the HSC Mathematics Extension 1 and 2 syllabus, but they are very good to know for higher education. In combinatorics, they help formalize counting by pairing elements perfectly between sets. In functions, bijections define exactly when an inverse function exists.
\end{remark}

\subsubsection*{Injective / One-to-one}
A function is injective if every output is mapped by \emph{at most} one input. In other words, no two different inputs will ever give you the same output. 

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
  % Domain
  \draw[thick, fill=blue!10] (-2,0) ellipse (1 and 2);
  \node at (-2,2.5) {Domain $X$};
  \node[circle,fill,inner sep=1.5pt, label=left:$a$] (x1) at (-2,1) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$b$] (x2) at (-2,0) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$c$] (x3) at (-2,-1) {};
  
  % Codomain
  \draw[thick, fill=red!10] (2,0) ellipse (1 and 2.5);
  \node at (2,3) {Codomain $Y$};
  \node[circle,fill,inner sep=1.5pt, label=right:$1$] (y1) at (2,1.5) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$2$] (y2) at (2,0.5) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$3$] (y3) at (2,-0.5) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$4$] (y4) at (2,-1.5) {};
  
  % Mappings
  \draw[->, thick, shorten >=2pt] (x1) -- (y1);
  \draw[->, thick, shorten >=2pt] (x2) -- (y3);
  \draw[->, thick, shorten >=2pt] (x3) -- (y4);
\end{tikzpicture}
\end{center}

\begin{itemize}[leftmargin=*]
  \item \textbf{Formal Definition:} If $f(a) = f(b)$, then it must be true that $a = b$.
  \item \textbf{Visualizing it:} Imagine a group of people (domain) and a group of chairs (codomain). An injective mapping means no two people are sitting in the same chair. There might be empty chairs left over, but no chair is double-occupied.
  \item \textbf{Math Example:} $f(x) = 2x$ is injective. Every unique $x$ produces a unique $y$. However, $f(x) = x^2$ is \emph{not} injective because $f(2) = 4$ and $f(-2) = 4$ (two different inputs hit the same output).
\end{itemize}

\subsubsection*{Surjective / Onto}

A function is surjective if every possible output in the target set (codomain) is hit by \emph{at least} one input. There are no ``leftover'' or unreachable outputs.

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
  % Domain
  \draw[thick, fill=blue!10] (-2,0) ellipse (1 and 2.5);
  \node at (-2,3) {Domain $X$};
  \node[circle,fill,inner sep=1.5pt, label=left:$a$] (x1) at (-2,1.5) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$b$] (x2) at (-2,0.5) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$c$] (x3) at (-2,-0.5) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$d$] (x4) at (-2,-1.5) {};
  
  % Codomain
  \draw[thick, fill=red!10] (2,0) ellipse (1 and 2);
  \node at (2,2.5) {Codomain $Y$};
  \node[circle,fill,inner sep=1.5pt, label=right:$1$] (y1) at (2,1) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$2$] (y2) at (2,0) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$3$] (y3) at (2,-1) {};
  
  % Mappings
  \draw[->, thick, shorten >=2pt] (x1) -- (y1);
  \draw[->, thick, shorten >=2pt] (x2) -- (y1);
  \draw[->, thick, shorten >=2pt] (x3) -- (y2);
  \draw[->, thick, shorten >=2pt] (x4) -- (y3);
\end{tikzpicture}
\end{center}

\begin{itemize}[leftmargin=*]
  \item \textbf{Visualizing it:} Every chair in the room has at least one person sitting in it. (Some chairs might have two people squeezing into them, but no chair is empty).
\end{itemize}

\subsubsection*{Bijective / One-to-one correspondence}
A function is bijective if it is \textbf{both} injective and surjective. Every input maps to a unique output, and every possible output is mapped to by an input. 

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
  % Domain
  \draw[thick, fill=blue!10] (-2,0) ellipse (1 and 2);
  \node at (-2,2.5) {Domain $X$};
  \node[circle,fill,inner sep=1.5pt, label=left:$a$] (x1) at (-2,1) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$b$] (x2) at (-2,0) {};
  \node[circle,fill,inner sep=1.5pt, label=left:$c$] (x3) at (-2,-1) {};
  
  % Codomain
  \draw[thick, fill=red!10] (2,0) ellipse (1 and 2);
  \node at (2,2.5) {Codomain $Y$};
  \node[circle,fill,inner sep=1.5pt, label=right:$1$] (y1) at (2,1) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$2$] (y2) at (2,0) {};
  \node[circle,fill,inner sep=1.5pt, label=right:$3$] (y3) at (2,-1) {};
  
  % Mappings
  \draw[->, thick, shorten >=2pt] (x1) -- (y2);
  \draw[->, thick, shorten >=2pt] (x2) -- (y1);
  \draw[->, thick, shorten >=2pt] (x3) -- (y3);
\end{tikzpicture}
\end{center}

\begin{itemize}[leftmargin=*]
  \item \textbf{Visualizing it:} It is a perfect pairing. There is exactly one person for every chair, and exactly one chair for every person. No empty chairs, no double-occupied chairs. 
  \item \textbf{Why it matters:} Bijective functions are incredibly important in mathematics because they are the \textbf{only} functions that have a perfect \textbf{inverse function} ($f^{-1}$). Because the mapping is a perfect two-way street, you can flawlessly travel backward from the output to the exact input. 
  \item \textbf{Math Example:} $f(x) = 2x + 1$ (where $x$ is any real number) is a bijection. You can easily find its inverse: $f^{-1}(x) = \frac{x - 1}{2}$.
\end{itemize}

\newtcolorbox{problembox*}[1][]
{
  title={\bfseries \ifstrempty{#1}{Problem}{#1}},
  enhanced,
  drop shadow={black!30!white},
  coltitle=bookwhite,
  colback=softivory,
  colframe=bookpurple,
  boxrule=1.5pt,
  arc=4pt,
  left=12pt,right=12pt,top=12pt,bottom=12pt,
  attach boxed title to top left={xshift=1.5em,yshift=-\tcboxedtitleheight/2},
  boxed title style={size=small,colback=bookpurple,coltitle=white},
  fonttitle=\bfseries,
  before skip=15pt,
  after skip=10pt
}
\newtcolorbox{solutionbox*}[1][]
{
  title={\bfseries \ifstrempty{#1}{Solution}{#1}},
  enhanced,
  drop shadow={black!20!white},
  coltitle=bookwhite,
  colback=softivory,
  colframe=bookred,
  boxrule=1.5pt,
  arc=4pt,
  left=10pt,right=10pt,top=10pt,bottom=10pt,
  attach boxed title to top left={xshift=1.5em,yshift=-\tcboxedtitleheight/2},
  boxed title style={size=small,colback=bookred,coltitle=white},
  fonttitle=\bfseries,
  fontupper=\footnotesize,
  before upper={
    \setlength{\abovedisplayskip}{4pt}%
    \setlength{\belowdisplayskip}{4pt}%
    \setlength{\abovedisplayshortskip}{3pt}%
    \setlength{\belowdisplayshortskip}{3pt}%
    \setlength{\jot}{2pt}%
  },
  before skip=15pt,
  after skip=10pt
}

\subsection{Multinomial Theorem}
\begin{remark}[Syllabus Note]
This is an enrichment topic and is outside the NESA syllabus.
\end{remark}

The Multinomial Theorem is a fundamental principle in algebra and combinatorics. It acts as the direct generalization of the well-known Binomial Theorem. While the Binomial Theorem provides a shortcut for expanding a sum of two terms raised to a power, like $(x + y)^n$, the Multinomial Theorem provides the formula for expanding a sum of any number of terms raised to a power, such as $(x_1 + x_2 + \dots + x_m)^n$.

\begin{theorem}[Multinomial Theorem]

For any positive integer $n$ and a sum of $m$ terms, the algebraic expansion is given by:
\[
(x_1 + x_2 + \dots + x_m)^n = \sum_{k_1+k_2+\dots+k_m=n} \frac{n!}{k_1!k_2!\dots k_m!} x_1^{k_1} x_2^{k_2} \dots x_m^{k_m}
\]


\end{theorem}

Here, $\sum_{k_1+k_2+\dots+k_m=n}$ means all possible combinations of 
non-negative integers $k_1, k_2, \dots, k_m$ that sum to $n$.


\begin{problembox*}[Worked example]
Find the last 3 digits of the coefficient of $x^{17}$ in the expansion of $(1 + x^3 + x^5 + x^7)^{97}$.
\end{problembox*}

\begin{solutionbox*}[Solution]
\textbf{1. Set up the General Term}\\
Using the multinomial theorem, the general term is:
\[
T = \frac{97!}{a! b! c! d!} x^{3b + 5c + 7d}
\]

\textbf{2. Establish Constraints}\\
Total Power: $a + b + c + d = 97$, Target Exponent: $3b + 5c + 7d = 17$.

\textbf{3. Find Valid $(a, b, c, d)$ Combinations}\\
Solve $3b + 5c + 7d = 17$ for non-negative integers by testing values of $d$:
\begin{itemize}[leftmargin=*]
\item \textbf{If $d = 2$:} $3b + 5c = 3 \implies b = 1, c = 0$. Solution: \textbf{$(94, 1, 0, 2)$}
\item \textbf{If $d = 1$:} $3b + 5c = 10 \implies b = 0, c = 2$. Solution: \textbf{$(94, 0, 2, 1)$}
\item \textbf{If $d = 0$:} $3b + 5c = 17 \implies b = 4, c = 1$. Solution: \textbf{$(92, 4, 1, 0)$}
\end{itemize}

\textbf{4. Calculate Coefficients Modulo 1000}\\
Evaluate $\frac{97!}{a! b! c! d!}$ for each valid set and find the remainder modulo 1000:

For $(94, 1, 0, 2)$:
\[
\frac{97!}{94! 1! 0! 2!} = \frac{97 \times 96 \times 95}{2} = 442320 \equiv \mathbf{320} \pmod{1000}
\]
For $(94, 0, 2, 1)$:
\[
\frac{97!}{94! 0! 2! 1!} = \frac{97 \times 96 \times 95}{2} = 442320 \equiv \mathbf{320} \pmod{1000}
\]
For $(92, 4, 1, 0)$:
\[
\frac{97!}{92! 4! 1! 0!} = \frac{97 \times 96 \times 95 \times 94 \times 93}{24} = 97 \times 4 \times 95 \times 94 \times 93
\]
*(Note: $97 \times 4 \times 95 = 36860 \equiv 860$, and $94 \times 93 = 8742 \equiv 742$. Then $860 \times 742 = 638120$)*
\[
T_3 \equiv \mathbf{120} \pmod{1000}
\]

\textbf{5. Final Sum}\\
Add the modulo results together:
\[
320 + 320 + 120 = \mathbf{760}
\]
The last 3 digits are \textbf{760}.
\end{solutionbox*}

\begin{problembox*}[Worked example]
Find the constant term (the term independent of $x$) in the expansion of:
\[
\left(2x^2 - \frac{1}{x} + 3\right)^8
\]
\end{problembox*}

\begin{solutionbox*}[Solution]
\textbf{1. Set up the General Term}\\
Using the multinomial theorem $(A + B + C)^n$, the general term is:
\[
T = \left[ \frac{8!}{a! b! c!} 2^a (-1)^b 3^c \right] x^{2a - b}
\]

\textbf{2. Establish Constraints}\\
Total Power: $a + b + c = 8$, Constant Term (Exponent = 0): $2a - b = 0 \implies b = 2a$.

\textbf{3. Find Valid $(a, b, c)$ Combinations}\\
Substitute $b = 2a$ into the total power constraint to get $3a + c = 8$. Test integers for $a \ge 0$ to find non-negative integer solutions for $c$:
\begin{itemize}[leftmargin=*]
\item If $a = 0$, then $c = 8$. Solution: \textbf{$(0, 0, 8)$}
\item If $a = 1$, then $c = 5$. Solution: \textbf{$(1, 2, 5)$}
\item If $a = 2$, then $c = 2$. Solution: \textbf{$(2, 4, 2)$}
\item (If $a \ge 3$, $c$ becomes negative, which is \textbf{invalid}.)
\end{itemize}

\textbf{4. Calculate Coefficients and Sum}\\
Substitute the valid combinations into the coefficient formula: $\frac{8!}{a! b! c!} 2^a (-1)^b 3^c$.

For $(0, 0, 8)$:
\[
\frac{8!}{0! 0! 8!} (2)^0 (-1)^0 (3)^8 = 6561
\]
For $(1, 2, 5)$:
\[
\frac{8!}{1! 2! 5!} (2)^1 (-1)^2 (3)^5 = 168 \times 2 \times 243 = 81648
\]
For $(2, 4, 2)$:
\[
\frac{8!}{2! 4! 2!} (2)^2 (-1)^4 (3)^2 = 420 \times 4 \times 9 = 15120
\]

\textbf{Final Answer:}\\
\[
6561 + 81648 + 15120 = \mathbf{103329}
\]
\end{solutionbox*}

\end{document}
