\begin{problem}[Asymmetric Cubic Inequality]
For any positive real numbers $a, b,$ and $c$, prove that:
\[
a^3 + b^3 + c^3 \ge a^2b + b^2c + c^2a
\]
\end{problem}

\begin{hint}
\begin{enumerate}
\item Standard AM-GM on $a^3 + b^3 + c^3$ will just give you $3\sqrt[3]{abc}$, which doesn't help you reach the right-hand side. You need to break the sum apart and attack it in smaller, strategic blocks.
\item Look at the target term $a^2b$. If you were to use the 3-variable AM-GM inequality, what combination of $a^3$ and $b^3$ terms would you need to put under the cube root to perfectly output an $a^2b$?
\item Try writing out an AM-GM inequality for the sum $a^3 + a^3 + b^3$. Once you see what happens, ``cycle'' the variables to create two more identical inequalities for the other terms, and sum them all vertically.
\end{enumerate}
\end{hint}

\begin{solution}
\textbf{Step 1: Construct the Lopsided AM-GM.}
Apply the 3-variable AM-GM inequality to the terms $a^3, a^3,$ and $b^3$:
\begin{align*}
    \frac{a^3 + a^3 + b^3}{3} &\ge \sqrt[3]{a^3 \cdot a^3 \cdot b^3} = \sqrt[3]{a^6b^3} = a^2b
\end{align*}
Multiply by 3:
\[
2a^3 + b^3 \ge 3a^2b
\]

\textbf{Step 2: Cycle the Variables.}
Repeat this exact same logical grouping for the other two target terms on the right-hand side ($b^2c$ and $c^2a$):
\begin{align*}
    2b^3 + c^3 &\ge 3b^2c \\
    2c^3 + a^3 &\ge 3c^2a
\end{align*}

\textbf{Step 3: Sum and Simplify.}
Add the three inequalities together vertically:
\begin{align*}
    (2a^3 + b^3) + (2b^3 + c^3) + (2c^3 + a^3) &\ge 3a^2b + 3b^2c + 3c^2a \\
    3a^3 + 3b^3 + 3c^3 &\ge 3(a^2b + b^2c + c^2a)
\end{align*}
Divide by 3 to complete the proof:
\[
    a^3 + b^3 + c^3 \ge a^2b + b^2c + c^2a
\]
\end{solution}

\begin{takeaways}
\item \textbf{Majorising / Cycling AM-GM:} When the target inequality looks asymmetric (like $a^2b$), you can intentionally ``weight'' the left-hand side of your AM-GM setup by doubling up on one variable ($a^3 + a^3 + b^3$) to manufacture the exact powers you need. This is a powerful technique for Extension 2.
\item \textbf{The Vertical Sum:} Proving a massive inequality in one shot is often impossible. Building three smaller, identical-logic inequalities and stacking them is one of the most reliable proof architectures.
\item \textbf{Equality Condition:} Equality holds when $a = b = c$ (when all three inequalities have equality simultaneously).
\item \textbf{Weighted AM-GM:} is what actually used in the solution.
\end{takeaways}

\begin{remark}[The Rearrangement Inequality]
This problem can be solved in one step using the out-of-syllabus \textbf{Rearrangement Inequality}. 

\textbf{Theorem:} If two sequences are similarly ordered (both increasing or both decreasing), then the sum obtained by pairing them in order is maximal.

\textbf{Proof outline:}
\begin{itemize}
    \item Without loss of generality, assume $a \ge b \ge c > 0$.
    \item Then $a^2 \ge b^2 \ge c^2$ (sequences $(a, b, c)$ and $(a^2, b^2, c^2)$ are similarly ordered).
    \item The Rearrangement Inequality states that the ``order-preserving'' pairing $a^2 \cdot a + b^2 \cdot b + c^2 \cdot c = a^3 + b^3 + c^3$ is maximal.
    \item Any permutation, such as the cyclic shift $a^2 \cdot b + b^2 \cdot c + c^2 \cdot a$, must be less than or equal to the perfectly matched sum.
    \item Therefore $a^3 + b^3 + c^3 \ge a^2b + b^2c + c^2a$.
\end{itemize}
\end{remark}

\begin{remark}[H\"older's Inequality]
For an alternative approach using more sophisticated machinery, we can apply \textbf{H\"older's Inequality}, which generalizes Cauchy-Schwarz to multiple sequences.

\textbf{Theorem:} For positive reals $a_i, b_i, c_i$ and exponents $p, q, r$ with $\frac{1}{p} + \frac{1}{q} + \frac{1}{r} = 1$:
\[
\sum_{i=1}^n a_i b_i c_i \le \left(\sum_{i=1}^n a_i^p\right)^{1/p} \left(\sum_{i=1}^n b_i^q\right)^{1/q} \left(\sum_{i=1}^n c_i^r\right)^{1/r}
\]


\textbf{Proof outline:}
\begin{itemize}
    \item Set $p = q = r = 3$ (so $\frac{1}{3} + \frac{1}{3} + \frac{1}{3} = 1$).
    \item Consider sequences: $(a, b, c)$, $(a, b, c)$, and $(b, c, a)$ (cyclic shift).
    \item Applying H\"older's inequality:
    \[
    a^2b + b^2c + c^2a = \sum ab \cdot a \cdot b_{\text{shifted}} \le (a^3+b^3+c^3)^{1/3} (a^3+b^3+c^3)^{1/3} (b^3+c^3+a^3)^{1/3}
    \]
    \item Simplifying: $a^2b + b^2c + c^2a \le a^3 + b^3 + c^3$.
\end{itemize}

\textit{Note:} H\"older's inequality is beyond Extension 2 but provides elegant connections to higher mathematics.
\end{remark}

\begin{remark}[Using Lagrange Multipliers]
Another method to solve this problem is to use \textbf{Lagrange Multipliers}.
\end{remark}

\begin{remark}[Using Cauchy-Schwarz inequality]
This problem can also be solved using the Cauchy-Schwarz inequality.
\end{remark}

\begin{remark}[Practice Problem]
Try proving the following related inequality for positive real numbers $a, b,$ and $c$:
\[
\frac{a^3}{b} + \frac{b^3}{c} + \frac{c^3}{a} \ge a^2 + b^2 + c^2
\]

\textbf{Hint:} Use a similar weighted AM-GM approach. Apply AM-GM to three carefully chosen terms for each target term on the right-hand side:
\begin{align*}
    \frac{a^3}{b} + \frac{a^3}{b} + b^2 &\ge 3a^2 \\
    \frac{b^3}{c} + \frac{b^3}{c} + c^2 &\ge 3b^2 \\
    \frac{c^3}{a} + \frac{c^3}{a} + a^2 &\ge 3c^2
\end{align*}
Then sum all three inequalities and simplify.
\end{remark}
