\begin{problem}[Power Mean, Young's, and AM-GM]

% \section*{Problem 3.26: Power Mean, Young's, and AM-GM}

Let $p, q,$ and $s$ be fixed positive real numbers with $p > 1$ and $q > 1$. \\
These constants satisfy the relationship $\frac{1}{p} + \frac{1}{q} = 1$.

\begin{enumerate}
    \item[(i)] Consider the function defined for $t > 0$:
    \[ f(t) = \frac{s^p}{p} + \frac{t^q}{q} - st \]
    Find the positive value of $t$ (in terms of $s$) that minimises $f(t)$.

    \item[(ii)] Hence, show that for all $t > 0$:
    \[ \frac{s^p}{p} + \frac{t^q}{q} \ge st \]
    \textit{(This result is known as Young's Inequality).}

    \item[(iii)] Prove by mathematical induction that:
    \[ (x_1 x_2 \cdots x_n)^{\frac{1}{n}} \le \frac{x_1 + x_2 + \cdots + x_n}{n} \]
    for all positive real numbers $x_1, x_2, \dots, x_n$.
    \textit{(This result is known as the Arithmetic Mean - Geometric Mean Inequality).}

    \item[(iv)] Using the result from part (iii), deduce that for all positive real numbers $y_1, y_2, \dots, y_n$:
    \[ \frac{y_1}{y_2} + \frac{y_2}{y_3} + \cdots + \frac{y_{n-1}}{y_n} + \frac{y_n}{y_1} \ge n \]
\end{enumerate}

% The numbers $p, q$ and $s$ are fixed and positive constants. 
% Also $p > 1, q > 1$ and  $1/p + 1/q = 1$.
% % $p = \frac{q}{q-1}$.
% \begin{enumerate}[label=(\roman*)]
%     \item What positive value of $t$ minimises the expression
%     \[ f(t) = \frac{s^p}{p} + \frac{t^q}{q} - st \ ? \]
%     \item Show that for all $t > 0$,
%     \[ \frac{s^p}{p} + \frac{t^q}{q} \geq st . \]
%     \item Prove by induction that
%     \[ (x_1 x_2 \cdots x_n)^{\frac{1}{n}} \leq \frac{x_1 + x_2 + \cdots + x_n}{n} \]
%     for all $x_1, \dots, x_n > 0$.
%     \item Deduce that, for all $y_1, y_2, \dots, y_n > 0$,
%     \[ \frac{y_1}{y_2} + \frac{y_2}{y_3} + \cdots + \frac{y_{n-1}}{y_n} + \frac{y_n}{y_1} \geq n . \]
% \end{enumerate}

\end{problem}

\begin{hint}
For (i), find the stationary point of $f(t)$. For (ii), evaluate $f$ at the minimum. For (iii), use induction on $n$ for AM-GM. For (iv), apply AM-GM to the cyclic ratios.
\end{hint}

\begin{solution}
	\textbf{(i)} $f'(t) = t^{q-1} - s = 0 \implies t = s^{1/(q-1)} = s^{p-1}$. $f''(t) > 0$ for $q > 1$, so this is a minimum.

	\textbf{(ii)} At $t = s^{p-1}$:
    \[ f(s^{p-1}) = \frac{s^p}{p} + \frac{s^p}{q} - s^p = s^p\left(\frac{1}{p} + \frac{1}{q} - 1\right) = 0 \]
    So $f(t) \geq 0$ for all $t > 0$.

	\textbf{(iii)} Induction for AM-GM: Base case $n=1$ is trivial. Assume for $k$, prove for $k+1$ by grouping and using the hypothesis.

	\textbf{(iv)} Apply AM-GM to $\frac{y_1}{y_2}, \dots, \frac{y_n}{y_1}$:
    \[ \sum_{cyc} \frac{y_i}{y_{i+1}} \geq n \left(1\right)^{1/n} = n \]
\end{solution}

\begin{takeaways}
\item Young's inequality generalizes AM-GM
\item Induction is a powerful tool for inequalities. And convexity applied for Inequalities. 
\item Cyclic sums often reduce to AM-GM
\item Pay close attention to how the product of cyclic terms ($\frac{y_1}{y_2} \dots \frac{y_n}{y_1}$) cancels to 1; this is a common technique in harder inequality proofs.
\item \textbf{{General Young's Inequality}:} 

Let $a_1, a_2, \dots, a_n$ be non-negative real numbers. Let $p_1, p_2, \dots, p_n$ be positive real numbers satisfying the condition:
\[
\sum_{i=1}^{n} \frac{1}{p_i} = 1
\]
Then the product of these numbers is bounded by the sum of their powers:
\[
\prod_{i=1}^{n} a_i \le \sum_{i=1}^{n} \frac{a_i^{p_i}}{p_i}
\]
Or written explicitly:
\[
a_1 a_2 \cdots a_n \le \frac{a_1^{p_1}}{p_1} + \frac{a_2^{p_2}}{p_2} + \cdots + \frac{a_n^{p_n}}{p_n}
\]
\textit{Equality holds if and only if } $a_1^{p_1} = a_2^{p_2} = \cdots = a_n^{p_n}$.

Link this inequality to weighted AM-GM inequalities.
\end{takeaways}
