\begin{problem}[Comparing coefficients in $(1+x)^{m+n}$]
Use
\[
(1+x)^{m+n}=(1+x)^m(1+x)^n
\]
and compare coefficients of $x^r$ to show that
\[
\binom{m+n}{r}=\binom{m}{0}\binom{n}{r}+\binom{m}{1}\binom{n}{r-1}+\cdots+\binom{m}{r}\binom{n}{0},
\]
where terms with impossible binomial coefficients are taken to be $0$.
\end{problem}

\begin{hint}
Find the coefficient of $x^r$ on the left first. Then, on the right, think about all the ways to obtain $x^r$ by multiplying a term from $(1+x)^m$ with a term from $(1+x)^n$.
\end{hint}

\begin{solution}
From the binomial theorem,
\[
(1+x)^{m+n}=\sum_{k=0}^{m+n}\binom{m+n}{k}x^k,
\]
so the coefficient of $x^r$ on the left is
\[
\binom{m+n}{r}.
\]

Also,
\[
(1+x)^m=\sum_{i=0}^m \binom{m}{i}x^i,
\qquad
(1+x)^n=\sum_{j=0}^n \binom{n}{j}x^j.
\]
To obtain an $x^r$ term in the product, we must choose $x^i$ from the first factor and $x^{r-i}$ from the second. Hence the coefficient of $x^r$ on the right is
\[
\sum_{i=0}^r \binom{m}{i}\binom{n}{r-i}.
\]

Since both expressions are the coefficient of $x^r$ in the same polynomial,
\[
\binom{m+n}{r}=\sum_{i=0}^r \binom{m}{i}\binom{n}{r-i}.
\]
That is,
\[
\boxed{
\binom{m+n}{r}
=\binom{m}{0}\binom{n}{r}
+\binom{m}{1}\binom{n}{r-1}
+\cdots
+\binom{m}{r}\binom{n}{0}
}.
\]
\end{solution}

\begin{takeaways}
\item Expanding the same expression in two ways is a standard way to prove combinatorial identities.
\item Comparing coefficients is especially useful when a problem involves powers of $(1+x)$.
\item This identity is often called Vandermonde's identity.
\end{takeaways}

\begin{problem}[Grid paths and binomial identities]
The diagram shows a $6\times 8$ grid. The aim is to walk from the point $A$ in the top left-hand corner to the point $B$ in the bottom right-hand corner by moving along the grid lines either downwards or to the right. Let
\[
C
\]
be the grid point that is $3$ steps to the right and $2$ steps down from $A$.

\begin{center}
\begin{tikzpicture}[scale=0.8]
    \draw[step=1cm, black] (0,0) grid (8,6);
    \filldraw (0,6) circle (2pt);
    \node[above left] at (0,6) {$A$};
    \filldraw (8,0) circle (2pt);
    \node[below right] at (8,0) {$B$};
    \filldraw (3,4) circle (2pt);
    \node[right] at (3.1,4) {$C$};
\end{tikzpicture}
\end{center}

\begin{enumerate}[label=\alph*)]
\item Find how many different routes are possible:
\begin{enumerate}[label=\roman*)]
\item without restriction
\item if you must pass through $C$.
\end{enumerate}

\item Consider the diagonal of grid points
\[
(0,6),(1,5),(2,4),\dots,(6,0).
\]
Explain why every route from $A$ to $B$ passes through exactly one of these points, and hence prove that
\[
\binom{14}{6}
=
\binom{8}{0}\binom{6}{6}
+\binom{8}{1}\binom{6}{5}
+\cdots
+\binom{8}{6}\binom{6}{0}.
\]

\item Generalise part (b) to prove that
\[
\binom{m+n}{n}
=
\binom{m}{0}\binom{n}{n}
+\binom{m}{1}\binom{n}{n-1}
+\cdots
+\binom{m}{n}\binom{n}{0},
\]
for integers $m\ge n\ge 0$.

\item By considering an $n\times n$ square grid and the corresponding diagonal, prove that
\[
\binom{2n}{n}
=
\binom{n}{0}^2+\binom{n}{1}^2+\cdots+\binom{n}{n}^2.
\]
\end{enumerate}
\end{problem}

\begin{hint}
For part (a), count the number of right and down moves. For part (b), split each route at the unique point where it meets the diagonal. Parts (c) and (d) are the corresponding general forms of the same idea.
\end{hint}

\begin{solution}
\textbf{a)(i)} A route from $A$ to $B$ uses $8$ right moves and $6$ down moves, so
\[
\binom{14}{6}=\binom{14}{8}.
\]

\textbf{a)(ii)} Split the route at $C$. From $A$ to $C$ we need $3$ right and $2$ down moves:
\[
\binom{5}{2}=10.
\]
From $C$ to $B$ we need $5$ right and $4$ down moves:
\[
\binom{9}{4}=126.
\]
Hence
\[
\binom{5}{2}\binom{9}{4}=10\cdot 126=1260.
\]

\textbf{b)} Every move decreases $x+y$ by $1$, so each route meets the diagonal
\[
x+y=6
\]
exactly once, at one of
\[
(0,6),(1,5),\dots,(6,0).
\]
If the meeting point is $(k,6-k)$, then the number of paths from $A$ to this point is
\[
\binom{6}{k}.
\]
From $(k,6-k)$ to $B$, the number of paths is
\[
\binom{8}{k}.
\]
So the number through this point is
\[
\binom{6}{k}\binom{8}{k}
=
\binom{8}{k}\binom{6}{6-k}.
\]
Summing over $k=0,1,\dots,6$ gives
\[
\binom{14}{6}
=
\binom{8}{0}\binom{6}{6}
+\binom{8}{1}\binom{6}{5}
+\cdots
+\binom{8}{6}\binom{6}{0}.
\]

\textbf{c)} In an $n\times m$ grid with $m\ge n$, the total number of routes is
\[
\binom{m+n}{n}.
\]
Split routes by the unique point where they meet
\[
x+y=n.
\]
If this occurs after $k$ right moves, then there are
\[
\binom{n}{k}
\]
ways to reach the diagonal point and
\[
\binom{m}{k}
\]
ways to complete the route. Hence
\[
\binom{m+n}{n}
=
\sum_{k=0}^{n}\binom{m}{k}\binom{n}{n-k}.
\]

\textbf{d)} For an $n\times n$ square grid,
\[
\binom{2n}{n}
=
\sum_{k=0}^{n}\binom{n}{k}\binom{n}{n-k}.
\]
Using the symmetry identity
\[
\binom{n}{n-k}=\binom{n}{k},
\]
we obtain
\[
\boxed{
\binom{2n}{n}
=
\binom{n}{0}^2+\binom{n}{1}^2+\cdots+\binom{n}{n}^2.
}
\]
\end{solution}

\begin{takeaways}
\item Grid-path counting gives a very natural combinatorial proof of Vandermonde-type identities.
\item Splitting paths at a suitable diagonal point is a standard and powerful technique.
\item The square-grid case leads directly to the classic identity
\[
\binom{2n}{n}=\sum_{k=0}^n \binom{n}{k}^2.
\]
\end{takeaways}

\begin{problem}[A coefficient identity for $\binom{4n}{2n}$]
\begin{enumerate}
  \item[(i)] Write down the coefficient of $x^{2n}$ in the binomial expansion of $(1+x)^{4n}$.

  \item[(ii)] Show that
  \[
  (1+x^2+2x)^{2n}=\sum_{k=0}^{2n}\binom{2n}{k}x^{2n-k}(x+2)^{2n-k}.
  \]

  \item[(iii)] It is known that
  \[
  x^{2n-k}(x+2)^{2n-k}
  =
  \binom{2n-k}{0}2^{2n-k}x^{2n-k}
  +\binom{2n-k}{1}2^{2n-k-1}x^{2n-k+1}
  +\cdots
  +\binom{2n-k}{2n-k}2^0x^{4n-2k}.
  \]
  Do \emph{not} prove this.

  Show that
  \[
  \binom{4n}{2n}
  =\sum_{k=0}^{n}2^{2n-2k}\binom{2n}{k}\binom{2n-k}{k}.
  \]
\end{enumerate}
\end{problem}

\begin{hint}
For part (ii), rewrite $1+x^2+2x$ as $1+x(x+2)$. For part (iii), compare the coefficient of $x^{2n}$ on both sides and identify which term in the inner expansion produces that power.
\end{hint}

\begin{solution}
\textbf{(i)} From the binomial theorem,
\[
\left(1+x\right)^{4n}=\sum_{r=0}^{4n}\binom{4n}{r}x^r,
\]
so the coefficient of $x^{2n}$ is $\binom{4n}{2n}$.
\[
\boxed{\binom{4n}{2n}}.
\]

\textbf{(ii)} Since $1+x^2+2x=1+x(x+2)$,
\[
\left(1+x^2+2x\right)^{2n}
=\left(1+x(x+2)\right)^{2n}
=\sum_{k=0}^{2n}\binom{2n}{k}x^{2n-k}(x+2)^{2n-k}.
\]

\textbf{(iii)} Also,
\[
(1+x^2+2x)^{2n}=((1+x)^2)^{2n}=(1+x)^{4n},
\]
so its coefficient of $x^{2n}$ is again $\binom{4n}{2n}$.

Now in
\[
\sum_{k=0}^{2n}\binom{2n}{k}x^{2n-k}(x+2)^{2n-k},
\]
a typical term from
\[
x^{2n-k}(x+2)^{2n-k}
\]
is
\[
\binom{2n-k}{r}2^{2n-k-r}x^{2n-k+r}.
\]
For an $x^{2n}$ term we need
\[
2n-k+r=2n,
\qquad \text{so } r=k.
\]
Hence the contribution from the $k$th summand is
\[
\binom{2n}{k}\binom{2n-k}{k}2^{2n-2k}.
\]
This is possible only when $k\le 2n-k$, i.e. $k\le n$. Therefore
\[
\binom{4n}{2n}
=\sum_{k=0}^{n}2^{2n-2k}\binom{2n}{k}\binom{2n-k}{k}.
\]
\end{solution}

\begin{takeaways}
\item This is another example of proving an identity by expanding the same expression in two different ways.
\item The key step is to compare coefficients of a carefully chosen power of $x$.
\item Bounds matter: the upper limit drops from $2n$ to $n$ because terms with $k>n$ cannot contribute to the coefficient of $x^{2n}$.
\end{takeaways}

\begin{problem}[Repeated draws of green counters]
A bag contains counters, some of which are green.

One hundred trials of an experiment are run. In each trial, one counter is selected from the bag at random and its colour noted. The counter is then returned to the bag.

Let $X$ be the random variable representing the number of times that a green counter is selected.

Given that $E(X)=20$, identify the distribution of $X$ and find the probability that a green counter is selected exactly $20$ times.
\end{problem}

\begin{hint}
The experiment consists of repeated independent trials with only two outcomes on each trial: green or not green.
\end{hint}

\begin{solution}
Since each trial is independent and each draw is with replacement, $X$ follows a binomial distribution:
\[
X\sim \operatorname{Bin}(100,p),
\]
where $p$ is the probability of selecting a green counter on any one draw.

For a binomial random variable,
\[
E(X)=np.
\]
Given that $E(X)=20$, we have
\[
100p=20 \quad \Rightarrow \quad p=0.2.
\]

So
\[
X\sim \operatorname{Bin}(100,0.2).
\]
Therefore,
\[
P(X=20)=\binom{100}{20}(0.2)^{20}(0.8)^{80}.
\]

Hence the required probability is
\[
\boxed{\binom{100}{20}(0.2)^{20}(0.8)^{80}}.
\]
\end{solution}

\begin{takeaways}
\item Repeated draws with replacement lead naturally to a binomial model.
\item For $X\sim \operatorname{Bin}(n,p)$, the mean is $E(X)=np$.
\item Once $p$ is known, exact probabilities are found using the binomial formula.
\end{takeaways}

\begin{problem}[Arrangement formulas for $x$ people]
Find the number of arrangements possible if $x$ people are
\begin{enumerate}[label=\alph*)]
\item in a straight line
\item in a circle
\item in a circle with $2$ particular people together
\item in a straight line with $3$ particular people together
\item in a circle with $2$ particular people not together.
\end{enumerate}
\end{problem}

\begin{hint}
Use $x!$ for a line and $(x-1)!$ for a circle. For ``together'' questions, bind people into a block. For ``not together'', subtract from the total.
\end{hint}

\begin{solution}
\textbf{a)} In a straight line:
\[
\boxed{x!}
\]

\textbf{b)} In a circle:
\[
\boxed{(x-1)!}
\]

\textbf{c)} In a circle with $2$ particular people together:
Treat the $2$ people as one block. Then there are $(x-1)$ objects in a circle, giving
\[
(x-2)!
\]
arrangements, and the pair can switch places in $2!$ ways. So
\[
\boxed{2(x-2)!}
\]

\textbf{d)} In a straight line with $3$ particular people together:
Treat the $3$ people as one block. Then there are $(x-2)$ objects in a line, giving
\[
(x-2)!
\]
arrangements, and the $3$ people inside the block can be arranged in $3!$ ways. So
\[
\boxed{3!(x-2)! = 6(x-2)!}
\]

\textbf{e)} In a circle with $2$ particular people not together:
Start with all circular arrangements:
\[
(x-1)!.
\]
Subtract the arrangements where the two particular people are together:
\[
2(x-2)!.
\]
Hence
\[
\boxed{(x-1)! - 2(x-2)!}.
\]
\end{solution}

\begin{takeaways}
\item The standard counts are $x!$ for a line and $(x-1)!$ for a circle.
\item ``Together'' usually suggests treating people as a single block.
\item ``Not together'' is often easiest by complementary counting.
\end{takeaways}

\begin{problem}[A pair summing to $21$]
Eleven numbers are randomly chosen from the set of integers
\[
S=\{1,2,3,4,\dots,20\}.
\]

Prove that the sum of two of the eleven numbers selected must equal $21$.
\end{problem}

\begin{hint}
Pair the numbers in the set so that each pair sums to $21$.
\end{hint}

\begin{solution}
The set $S$ can be split into the following $10$ pairs:
\[
(1,20),\ (2,19),\ (3,18),\ \dots,\ (10,11).
\]
Each pair sums to $21$.

Now suppose $11$ numbers are chosen from the $20$ numbers in $S$.

There are only $10$ such pairs, but we have chosen $11$ numbers. By the pigeonhole principle, at least one complete pair must be selected.

Therefore, among the $11$ chosen numbers, there must be two numbers whose sum is $21$.
\end{solution}

\begin{takeaways}
\item A strong first step in pigeonhole problems is to group objects into meaningful boxes.
\item Here, the natural boxes are the pairs that sum to $21$.
\item Choosing more objects than there are boxes forces one box to contain two chosen objects.
\end{takeaways}

\begin{problem}[Positive integer solutions]
Show that there are
\[
\binom{n-1}{k-1}
\]
distinct positive integer-valued vectors $(x_1,\dots,x_k)$ satisfying
\[
x_1+\cdots+x_k=n,
\qquad x_i>0 \text{ for all } i=1,\dots,k.
\]
\end{problem}

\begin{hint}
Think of writing $n$ as a row of $n$ identical objects, then inserting dividers to split them into $k$ positive groups.
\end{hint}

\begin{solution}
Represent $n$ as $n$ identical stars:
\[
\underbrace{*\,*\,*\,\cdots\,*}_{n \text{ stars}}.
\]
To split these into $k$ positive parts, we must place $k-1$ dividers into the gaps between consecutive stars.

There are
\[
n-1
\]
gaps between the $n$ stars, and we choose
\[
k-1
\]
of them for the dividers.

Hence the number of positive integer solutions is
\[
\boxed{\binom{n-1}{k-1}}.
\]
\end{solution}

\begin{takeaways}
\item This is the standard stars-and-bars count for positive integer solutions.
\item Positivity means every group must contain at least one star, so dividers can only go in the internal gaps.
\end{takeaways}

\begin{problem}[Non-negative integer solutions]
Show that there are
\[
\binom{n+k-1}{k-1}
\]
distinct non-negative integer-valued vectors $(x_1,\dots,x_k)$ satisfying
\[
x_1+\cdots+x_k=n,
\qquad x_i\ge 0 \text{ for all } i=1,\dots,k.
\]
\end{problem}

\begin{hint}
Use the same stars-and-bars idea, but now allow some groups to be empty.
\end{hint}

\begin{solution}
Again represent $n$ as $n$ identical stars. This time we want to split them into $k$ groups, but some groups may be empty.

So we place $k-1$ dividers among the $n$ stars, allowing dividers to appear at the ends or next to each other.

This is equivalent to arranging
\[
n \text{ stars and } k-1 \text{ dividers}
\]
in a row. The total number of symbols is
\[
n+k-1.
\]
We simply choose the positions of the $k-1$ dividers:
\[
\boxed{\binom{n+k-1}{k-1}}.
\]
\end{solution}

\begin{takeaways}
\item This is the standard stars-and-bars count for non-negative integer solutions.
\item Allowing zero means dividers may be adjacent or may appear at the ends.
\end{takeaways}

\begin{problem}[Three identities from $(1+x)^n$]
Consider the binomial expansion
\[
(1+x)^n=\binom{n}{0}+\binom{n}{1}x+\binom{n}{2}x^2+\cdots+\binom{n}{n}x^n.
\]

\begin{enumerate}[label=\roman*)]
\item Deduce that
\[
\binom{2n}{0}+\binom{2n}{2}+\cdots+\binom{2n}{2n}
=
\binom{2n}{1}+\binom{2n}{3}+\cdots+\binom{2n}{2n-1}.
\]

\item Integrate both sides with respect to $x$ from $0$ to $1$, and deduce that
\[
\sum_{r=0}^n \frac{1}{r+1}\binom{n}{r}=\frac{2^{n+1}-1}{n+1}.
\]

\item Differentiate both sides with respect to $x$, then let $x=1$, and deduce that
\[
\binom{n}{1}+2\binom{n}{2}+3\binom{n}{3}+\cdots+n\binom{n}{n}=n2^{n-1}.
\]
\end{enumerate}
\end{problem}

\begin{hint}
For each part, start from the same binomial expansion and apply the requested operation directly to both sides.
\end{hint}

\begin{solution}
\textbf{i)} Apply the identity
\[
(1+x)^{2n}=\sum_{r=0}^{2n} \binom{2n}{r}x^r
\]
and let $x=-1$. Then
\[
\sum_{r=0}^{2n}\binom{2n}{r}(-1)^r=0.
\]
So
\[
\binom{2n}{0}-\binom{2n}{1}+\binom{2n}{2}-\binom{2n}{3}+\cdots+\binom{2n}{2n}=0.
\]
Hence the sum of the even-index terms equals the sum of the odd-index terms:
\[
\boxed{
\binom{2n}{0}+\binom{2n}{2}+\cdots+\binom{2n}{2n}
=
\binom{2n}{1}+\binom{2n}{3}+\cdots+\binom{2n}{2n-1}.
}
\]

\textbf{ii)} Integrating from $0$ to $1$,
\[
\int_0^1 (1+x)^n\,dx=\int_0^1 \sum_{r=0}^n \binom{n}{r}x^r\,dx.
\]
The left-hand side is
\[
\left[\frac{(1+x)^{n+1}}{n+1}\right]_0^1=\frac{2^{n+1}-1}{n+1}.
\]
The right-hand side is
\[
\sum_{r=0}^n \binom{n}{r}\int_0^1 x^r\,dx
=\sum_{r=0}^n \binom{n}{r}\frac{1}{r+1}.
\]
Hence
\[
\boxed{\sum_{r=0}^n \frac{1}{r+1}\binom{n}{r}=\frac{2^{n+1}-1}{n+1}.}
\]

\textbf{iii)} Differentiating,
\[
n(1+x)^{n-1}=\binom{n}{1}+2\binom{n}{2}x+3\binom{n}{3}x^2+\cdots+n\binom{n}{n}x^{n-1}.
\]
Now let $x=1$:
\[
n2^{n-1}=\binom{n}{1}+2\binom{n}{2}+3\binom{n}{3}+\cdots+n\binom{n}{n}.
\]
So
\[
\boxed{\binom{n}{1}+2\binom{n}{2}+3\binom{n}{3}+\cdots+n\binom{n}{n}=n2^{n-1}.}
\]
\end{solution}

\begin{takeaways}
\item A single binomial expansion can produce several useful identities by substitution, integration, and differentiation.
\item Setting $x=-1$ often produces alternating-sign identities.
\item Integrating or differentiating term-by-term can turn coefficients into new weighted sums.
\end{takeaways}

\begin{problem}[Making an amount of money with Australian denominations]
Suppose we use the simplified Australian denominations (coins less than $\$1$ are not used): 
\[
\$1,\ \ \$2,\ \ \$5,\ \ \$10,\ \ \$20,\ \ \$50,\ \ \$100.
\]

For a fixed positive integer amount $M$, let $N(M)$ be the number of ways to express $\$M$ using these coins and notes, where order does \emph{not} matter.

\begin{enumerate}[label=\roman*)]
\item Write down a generating function whose coefficient of $x^M$ is $N(M)$.
\item Hence explain how the problem of finding $N(M)$ can be turned into a coefficient problem.
\end{enumerate}

\smallskip
\noindent\textit{Enrichment note:} This problem is outside the HSC Mathematics Extension~1 syllabus, but it is a very useful example of how generating functions turn a counting problem into algebra.
\end{problem}

\begin{hint}
For each denomination $d$, think about the possible contributions $0,d,2d,3d,\dots$.
\end{hint}

\begin{solution}
For each denomination, we write a generating function for how many of that denomination may be used:
\[
1+x+x^2+x^3+\cdots \qquad \text{for } \$1,
\]
\[
1+x^2+x^4+x^6+\cdots \qquad \text{for } \$2,
\]
and similarly for the other notes.

So the full generating function is
\[
G(x)=
(1+x+x^2+\cdots)
(1+x^2+x^4+\cdots)
(1+x^5+x^{10}+\cdots)
(1+x^{10}+x^{20}+\cdots)
\]
\[
\qquad\qquad
(1+x^{20}+x^{40}+\cdots)
(1+x^{50}+x^{100}+\cdots)
(1+x^{100}+x^{200}+\cdots).
\]

The coefficient of $x^M$ in $G(x)$ is exactly the number of ways to make $\$M$, because each chosen power of $x$ records how much each denomination contributes, and multiplying the choices combines them.

So
\[
N(M)=[x^M]\,G(x),
\]
where $[x^M]\,G(x)$ means ``the coefficient of $x^M$ in $G(x)$.'' 

\textbf{Outline of method.} To solve a specific case, such as $M=20$ or $M=50$, expand only as far as needed and collect the coefficient of $x^M$. In practice, this is often done using recurrence methods or computer algebra, because the coefficients grow quickly.
\end{solution}

\begin{takeaways}
\item A generating function is a power series in which the coefficient of $x^M$ stores the number of ways to make the amount $M$.
\item For a sequence of numbers $a_0,a_1,a_2,\dots$, the generating function is
\[a_0+a_1x+a_2x^2+\cdots.
\]
\item For the $\$1$ denomination, the generating function is
\[
1+x+x^2+x^3+\cdots
\]
because we may use $0,1,2,3,\dots$ one-dollar coins, contributing $\$0,\$1,\$2,\$3,\dots$ respectively.

\item The full generating function is the product of the generating functions for each denomination, because we can choose how many of each denomination to use independently.
\item Each other denomination contributes a similar series, and multiplying these series combines the choices from all denominations.
\item This is a standard higher-level combinatorics technique and is excellent enrichment beyond the HSC course.
\end{takeaways}

\begin{remark}[Algorithmic note]
We can also compute $N(M)$ using dynamic programming. This is the algorithmic version of finding the coefficient of $x^M$ in the generating function.
\end{remark}

\begin{lstlisting}[language=Python,numbers=left,basicstyle=\ttfamily\small,keywordstyle=\color{bookpurple}\bfseries,commentstyle=\color{bookred},stringstyle=\color{blue!60!black},showstringspaces=false,breaklines=true]
def calculate_coin_combinations(target_amount, denominations):
    """
    Calculates the number of ways to make a target amount using given denominations.
    This is the algorithmic equivalent of finding the coefficient of x^M
    in our generating function.
    """
    # dp is an array where dp[i] will store the number of ways to make amount i.
    # We initialize an array of zeros from 0 up to our target_amount.
    dp = [0] * (target_amount + 1)

    # Base case: There is exactly 1 way to make $0 (by choosing no coins)
    dp[0] = 1

    # Outer loop: We process one denomination bracket at a time.
    # This is exactly like multiplying the next bracket in our generating function:
    # (1 + x^2 + x^4...) * (1 + x^5 + x^10...)
    for coin in denominations:

        # Inner loop: We update the combinations for every amount
        # from 'coin' up to the target.
        for amount in range(coin, target_amount + 1):

            # The number of ways to make the current 'amount' increases by the
            # number of ways we could make the amount WITHOUT this current coin.
            dp[amount] += dp[amount - coin]

    # The final answer is stored at the index of our target amount
    return dp[target_amount]


# --- Example Usage ---
australian_denominations = [1, 2, 5, 10, 20, 50, 100]
M = 10

ways = calculate_coin_combinations(M, australian_denominations)
print(f"To make ${M}, there are {ways} combinations.")
# Output: To make $10, there are 47 combinations.
\end{lstlisting}

\begin{problem}[Unlocking a phone by trial and error]
To unlock a phone, we need to enter a $4$-digit PIN.

We know that the four digits are $1,2,3,4$, used distinctly, but we do not know the correct order.

If we enter a wrong PIN:
\begin{itemize}[leftmargin=*]
  \item after the first wrong attempt, the phone locks the next input for $4$ seconds
  \item after the second wrong attempt, the phone locks the next input for $4\times 2$ seconds
  \item after the third wrong attempt, the phone locks the next input for $4\times 2\times 2$ seconds
  \item and so on.
\end{itemize}

Find the average time needed to unlock the phone by trial and error, assuming each possible PIN order is equally likely to be the correct one and the time to type each PIN is negligible.
\end{problem}

\begin{hint}
How many possible PINs are there? If the correct PIN is found on the $m$th attempt, how much total waiting time has occurred before that attempt?
\end{hint}

\begin{solution}
The digits $1,2,3,4$ can be arranged in
\[
4!=24
\]
different ways, so if we test PINs one by one, the correct PIN is equally likely to appear in positions
\[
1,2,3,\dots,24.
\]

If the correct PIN is found on the $m$th attempt, then there have been $(m-1)$ wrong attempts before success.

The waiting times after the wrong attempts are
\[
4,\ 8,\ 16,\ \dots,\ 4\cdot 2^{m-2}.
\]
So the total waiting time before the $m$th attempt is
\[
4+8+16+\cdots+4\cdot 2^{m-2}
=4(2^{m-1}-1).
\]

Therefore the average waiting time is
\[
\frac{1}{24}\sum_{m=1}^{24}4(2^{m-1}-1).
\]
Now
\[
\sum_{m=1}^{24}2^{m-1}=2^{24}-1,
\]
so
\begin{align*}
\frac{1}{24}\sum_{m=1}^{24}4(2^{m-1}-1)
&= \frac{4}{24}\left((2^{24}-1)-24\right) \\
&= \frac{1}{6}(2^{24}-25) \\
&= \frac{16777191}{6}.
\end{align*}

Hence the average time needed is
\[
\boxed{\frac{16777191}{6}\text{ seconds}}
\]
which is about
\[
\boxed{2796198.5\text{ seconds}}.
\]
\end{solution}

\begin{takeaways}
\item When all possible arrangements are equally likely, the correct one is equally likely to occur in any position of your trial order.
\item Exponentially growing waiting times often lead to a geometric series.
\item Expected-value questions can often be solved by averaging over all possible positions of success.
\end{takeaways}

\begin{remark}
\begin{center}
\itshape
``Space is big. Really big. You just won't believe how vastly, hugely, mind-bogglingly big it is.
I mean, you may think it's a long way down the road to the chemist, but that's just peanuts to
space.''\\
\normalfont Douglas Adams
\end{center}
\end{remark}

\begin{problem}[Polynomial derivatives and a power-series expansion]
Show that if
\[
f(x)=a_nx^n+a_{n-1}x^{n-1}+\cdots+a_1x+a_0
\]
is any polynomial, then
\[
f^{(k)}(0)=
\begin{cases}
a_k\,k!, & \text{for } k=0,1,2,\dots,n,\\
0, & \text{for } k>n.
\end{cases}
\]

Hence explain why
\[
f(x)=\frac{f(0)}{0!}+\frac{f'(0)}{1!}x+\frac{f''(0)}{2!}x^2+\frac{f^{(3)}(0)}{3!}x^3+\cdots.
\]

\noindent\textit{Enrichment note:} This problem is outside the HSC Mathematics Extension~1 syllabus, but it is a very useful first step toward power series and Taylor series.
\end{problem}

\begin{hint}
Differentiate a general term $a_r x^r$ exactly $k$ times, then substitute $x=0$. Most terms vanish.
\end{hint}

\begin{solution}
Write
\[
f(x)=\sum_{r=0}^{n} a_r x^r.
\]
If we differentiate $k$ times, then
\[
\frac{d^k}{dx^k}(a_r x^r)=
\begin{cases}
a_r\,r(r-1)\cdots (r-k+1)x^{r-k}, & r\ge k,\\
0, & r<k.
\end{cases}
\]

Now substitute $x=0$. Every term with $r>k$ still contains a factor of $x$ and becomes $0$. The only surviving term is the one with $r=k$, giving
\[
f^{(k)}(0)=a_k\,k!
\]
for $k=0,1,2,\dots,n$.

If $k>n$, every term has already vanished after enough differentiations, so
\[
f^{(k)}(0)=0.
\]

Therefore
\[
a_k=\frac{f^{(k)}(0)}{k!},
\]
so the polynomial can be rewritten as
\[
f(x)=a_0+a_1x+a_2x^2+\cdots+a_nx^n
=\frac{f(0)}{0!}+\frac{f'(0)}{1!}x+\frac{f''(0)}{2!}x^2+\cdots+\frac{f^{(n)}(0)}{n!}x^n.
\]
Since $f$ is a polynomial, this series has only finitely many nonzero terms.
\end{solution}

\begin{takeaways}
\item Repeated derivatives at $x=0$ recover the coefficients of a polynomial.
\item The factor of $k!$ appears because differentiating $x^k$ exactly $k$ times gives $k!$.
\item This is the polynomial version of the idea behind Taylor series, so it is a natural enrichment topic beyond the HSC course.
\end{takeaways}

\begin{problem}[Binary strings and a binomial identity]
Computing is based on binary strings, which are sequences of $0$s and $1$s, such as $00101$ and $1001110$.

The problem in this question is to find the number $N$ of binary strings that contain exactly $a$ zeros and at most $b$ ones, where $a$ and $b$ are whole numbers. We will count such strings in two different ways, and hence prove an identity.

\begin{enumerate}[label=\alph*)]
\item Using ${}^nC_r$ notation, find how many binary strings with exactly $a$ zeros contain:
\begin{enumerate}[label=\roman*)]
\item no $1$s
\item one $1$
\item two $1$s
\item three $1$s
\item $r$ $1$s.
\end{enumerate}

\item Hence prove that
\[
N=\binom{a}{0}+\binom{a+1}{1}+\binom{a+2}{2}+\cdots+\binom{a+b}{b}.
\]

\item Let $S$ be a binary string consisting of $a$ zeros and $r$ ones, where $r\le b$. Extend $S$ to a longer binary string $S01\cdots 1$ by adding $0$ on the right, and then adding $b-r$ ones. How many $0$s and $1$s are there in the resulting string, and what is its total length?

\item Describe the inverse process by which a string of $(a+1)$ zeros and $b$ ones can be converted to a string of $a$ zeros and at most $b$ ones, and show that the two processes are one-to-one correspondences.

\item Hence prove that
\[
N=\binom{a+b+1}{b}.
\]

\item Deduce that
\[
\binom{a}{0}+\binom{a+1}{1}+\binom{a+2}{2}+\cdots+\binom{a+b}{b}
=
\binom{a+b+1}{b}.
\]
\end{enumerate}
\end{problem}

\begin{hint}
For part (a), if a string has exactly $a$ zeros and $r$ ones, then its length is $a+r$, and you only need to choose where the $r$ ones go. For parts (c) and (d), build a bijection between the two families of strings.
\end{hint}

\begin{solution}
\textbf{a)} If a binary string has exactly $a$ zeros and exactly $r$ ones, then its total length is
\[
a+r.
\]
We may choose the positions of the $r$ ones (or equivalently the $a$ zeros), so the number of such strings is
\[
\binom{a+r}{r}.
\]
Hence:
\[
\text{i) } \binom{a}{0},\qquad
\text{ii) } \binom{a+1}{1},\qquad
\text{iii) } \binom{a+2}{2},\qquad
\text{iv) } \binom{a+3}{3},\qquad
\text{v) } \binom{a+r}{r}.
\]

\textbf{b)} Since the number of ones can be $0,1,2,\dots,b$, summing these cases gives
\[
N=\binom{a}{0}+\binom{a+1}{1}+\binom{a+2}{2}+\cdots+\binom{a+b}{b}.
\]

\textbf{c)} Start with a binary string $S$ containing $a$ zeros and $r$ ones, where $r\le b$.
Append one extra $0$, then append $(b-r)$ ones. The resulting string has
\[
a+1
\]
zeros and
\[
r+(b-r)=b
\]
ones, so its total length is
\[
a+b+1.
\]

\textbf{d)} Conversely, given a binary string with $(a+1)$ zeros and $b$ ones, locate the last zero in the string and delete it together with all the ones to its right.

What remains is a binary string with exactly $a$ zeros and at most $b$ ones.

This process is the inverse of the construction in part (c), so the two sets of strings are in one-to-one correspondence.

\textbf{e)} Therefore, the number $N$ is equal to the number of binary strings with $(a+1)$ zeros and $b$ ones.
Such strings have length
\[
a+b+1,
\]
and we choose the positions of the $b$ ones, so
\[
N=\binom{a+b+1}{b}.
\]

\textbf{f)} Equating the two expressions for $N$ gives
\[
\boxed{
\binom{a}{0}+\binom{a+1}{1}+\binom{a+2}{2}+\cdots+\binom{a+b}{b}
=
\binom{a+b+1}{b}.
}
\]
\end{solution}

\begin{takeaways}
\item This is a nice example of proving an identity by counting the same set in two different ways.
\item Strings with fixed numbers of zeros and ones are counted by choosing positions.
\item Bijections are powerful: if two families of objects can be matched one-to-one, then they have the same size.
\end{takeaways}

\begin{problem}[Derangements]
A \emph{derangement} of $n$ distinct letters is a permutation of them so that no letter appears in its original position. For example, DABC is a derangement of ABCD, but DACB is not. Denote the number of derangements of $n$ letters by $D(n)$.

\begin{enumerate}[label=\alph*)]
\item By listing all the derangements of A, AB, ABC and ABCD, find the values of $D(1)$, $D(2)$, $D(3)$ and $D(4)$.

\item Suppose that we have formed a derangement of the five letters ABCDE. Let the last letter in the derangement be $X$, and exchange $X$ with E. This puts E back into its original position. Either $X$ is now also in its original position so that three letters are away from their original positions, or $X$ is not in its original position so that four letters are away from their original positions. Hence explain why
\[
D(5)=4D(4)+4D(3).
\]

\item Use this formula to evaluate $D(5)$. Then apply the corresponding arguments and formulae to evaluate $D(6)$, $D(7)$ and $D(8)$.

\item Explain how this can be generalised to
\[
D(n)=n!\left(\frac{1}{0!}-\frac{1}{1!}+\frac{1}{2!}-\frac{1}{3!}+\cdots+\frac{(-1)^n}{n!}\right).
\]

\item The alternating sum in part (d) approaches $\dfrac1e$ as $n\to\infty$. Explain what this means for the ratio
\[
\frac{D(n)}{n!}
\]
and give both a combinatorial and a probabilistic interpretation of this result.
\end{enumerate}

\noindent\textit{Enrichment note:} Derangements are a classic higher-level counting topic. They are not part of the standard HSC Mathematics Extension~1 syllabus, but they are an excellent example of recurrence relations in combinatorics.
\end{problem}

\begin{hint}
For part (a), list small cases directly. For parts (b) and (c), fix where E goes, then split into the two cases described in the question. For part (d), think about inclusion-exclusion: start from all permutations and subtract those with one or more fixed points.
\end{hint}

\begin{solution}
\textbf{a)} By direct listing:
\[
D(1)=0,\qquad D(2)=1,\qquad D(3)=2,\qquad D(4)=9.
\]

\textbf{b)} In a derangement of ABCDE, the letter E cannot stay in the last position, so it has $4$ possible positions. Let the last letter be $X$, and swap E with $X$.
Either $X$ lands in its own position, giving $D(3)$ possibilities, or it does not, giving $D(4)$ possibilities. Hence
\[
D(5)=4D(4)+4D(3).
\]

\textbf{c)} Using $D(4)=9$ and $D(3)=2$,
\[
D(5)=4\cdot 9+4\cdot 2=44.
\]

More generally,
\[
D(n)=(n-1)\bigl(D(n-1)+D(n-2)\bigr).
\]
Thus

% \[
% D(6)=5(D(5)+D(4))=5(44+9)=265,
% \]
% \[
% D(7)=6(D(6)+D(5))=6(265+44)=1854,
% \]
% \[
% D(8)=7(D(7)+D(6))=7(1854+265)=14833.
% \]
\[
\boxed{D(5)=44,\quad D(6)=265,\quad D(7)=1854,\quad D(8)=14833.}
\]

\textbf{d)} Let $A_i$ be the set of permutations in which the $i$th letter stays fixed. Then
\[
|A_i|=(n-1)!,
\qquad
|A_i\cap A_j|=(n-2)!,
\]
and in general fixing $k$ specified letters leaves
\[
(n-k)!
\]
permutations. Hence, by inclusion-exclusion,
\[
D(n)=n!-\binom{n}{1}(n-1)!+\binom{n}{2}(n-2)!-\cdots+(-1)^n\binom{n}{n}(n-n)!.
\]
Using $\binom{n}{k}(n-k)!=\dfrac{n!}{k!}$, we can conclude that,

% \[
% D(n)=n!\left(1-\frac1{1!}+\frac1{2!}-\frac1{3!}+\cdots+\frac{(-1)^n}{n!}\right),
% \]
% that is,
\[
\boxed{
D(n)=n!\left(\frac{1}{0!}-\frac{1}{1!}+\frac{1}{2!}-\frac{1}{3!}+\cdots+\frac{(-1)^n}{n!}\right).
}
\]

\textbf{e)} Dividing by $n!$ gives
\[
\frac{D(n)}{n!}
=
\frac{1}{0!}-\frac{1}{1!}+\frac{1}{2!}-\frac{1}{3!}+\cdots+\frac{(-1)^n}{n!}.
\]
As $n\to\infty$, the right-hand side approaches $\dfrac1e$, so
\[
\boxed{\frac{D(n)}{n!}\to \sum_{n=0}^{\infty} \frac{(-1)^n}{n!} = \frac1e.}
\]
\end{solution}

\begin{takeaways}
\item \textit{Combinatorially,} about a fraction $\dfrac1e$ of all permutations are derangements for large $n$.
\item \textit{Probabilistically,} a random permutation has no fixed points with probability approaching $\dfrac1e$.
\item A derangement is a permutation with no fixed points.
\item Small cases are often best handled by direct listing before a general pattern appears.
\item This problem is a good example of a recurrence relation arising from careful casework.
\item Inclusion-exclusion gives a closed formula for $D(n)$ by counting permutations with fixed points.
\item The ratio $\dfrac{D(n)}{n!}$ approaching $\dfrac1e$ means derangements make up about $\dfrac1e$ of all permutations when $n$ is large.

\item The Taylor series for $e^x$ is
\[
e^{x}= \sum_{n=0}^{\infty} \frac{x^n}{n!}.
\]

\end{takeaways}

\begin{problem}[A monochromatic triangle on six vertices]
A regular octahedron has $6$ vertices. Label them $A_0,A_1,\dots,A_5$. Each pair of vertices is joined by a rod, and each rod is coloured either yellow or blue.

\begin{enumerate}[label=\alph*)]
\item Explain why there are $15$ rods.
\item Each set of three vertices, together with the rods joining them, forms a triangle. Explain why there are $20$ such triangles.
\item Explain why there must be at least one triangle whose rods all have the same colour.
\end{enumerate}

\begin{center}
\begin{tikzpicture}[scale=1.15]
    \coordinate (A0) at (0,2.2);
    \coordinate (A1) at (-1.8,0.9);
    \coordinate (A2) at (1.8,0.9);
    \coordinate (A3) at (-1.8,-0.9);
    \coordinate (A4) at (1.8,-0.9);
    \coordinate (A5) at (0,-2.2);

    % All rods
    \draw[yellow!70!orange, line width=1.1pt] (A0) -- (A1);
    \draw[blue!70!black,   line width=1.1pt] (A0) -- (A2);
    \draw[yellow!70!orange, line width=1.1pt] (A0) -- (A3);
    \draw[blue!70!black,   line width=1.1pt] (A0) -- (A4);
    \draw[yellow!70!orange, line width=1.1pt] (A0) -- (A5);

    \draw[blue!70!black,   line width=1.1pt] (A1) -- (A2);
    \draw[yellow!70!orange, line width=1.1pt] (A1) -- (A3);
    \draw[blue!70!black,   line width=1.1pt] (A1) -- (A4);
    \draw[yellow!70!orange, line width=1.1pt] (A1) -- (A5);

    \draw[yellow!70!orange, line width=1.1pt] (A2) -- (A3);
    \draw[blue!70!black,   line width=1.1pt] (A2) -- (A4);
    \draw[yellow!70!orange, line width=1.1pt] (A2) -- (A5);

    \draw[blue!70!black,   line width=1.1pt] (A3) -- (A4);
    \draw[yellow!70!orange, line width=1.1pt] (A3) -- (A5);
    \draw[blue!70!black,   line width=1.1pt] (A4) -- (A5);

    % Vertices
    \foreach \P/\Lbl in {A0/A_0,A1/A_1,A2/A_2,A3/A_3,A4/A_4,A5/A_5}{
        \filldraw[black] (\P) circle (1.8pt);
        \node[fill=white, inner sep=1pt] at ($(\P)+(0.25,0.22)$) {$\Lbl$};
    }
\end{tikzpicture}
\end{center}

\noindent\textit{Enrichment note:} This is a small and beautiful example of a Ramsey-type idea. It is beyond the standard HSC Mathematics Extension~1 syllabus, but it is a great illustration of how counting and the pigeonhole principle can force structure.
\end{problem}

\begin{hint}
For part (c), focus on one vertex, say $A_0$, and look at the $5$ rods from that vertex. Use the pigeonhole principle on their colours.
\end{hint}

\begin{solution}
\textbf{a)} A rod joins a pair of distinct vertices. Since there are $6$ vertices, the number of rods is
\[
\binom{6}{2}=15.
\]

\textbf{b)} Any choice of $3$ vertices determines a triangle. So the number of triangles is
\[
\binom{6}{3}=20.
\]

\textbf{c)} Focus on the vertex $A_0$. It is joined to the other $5$ vertices by $5$ rods, and each rod is coloured either yellow or blue.

By the pigeonhole principle, at least $3$ of these $5$ rods must have the same colour. Without loss of generality, suppose
\[
A_0A_1,\ A_0A_2,\ A_0A_3
\]
are all yellow.

Now consider the triangle formed by $A_1,A_2,A_3$.
\begin{itemize}[leftmargin=*]
  \item If one of the rods $A_1A_2$, $A_1A_3$, or $A_2A_3$ is yellow, then together with $A_0$ it forms a yellow triangle.
  \item If none of those rods is yellow, then all three are blue, so $A_1A_2A_3$ is a blue triangle.
\end{itemize}

In either case, there is at least one triangle whose rods all have the same colour.
\end{solution}

\begin{takeaways}
\item This is a classic example of combining counting with the pigeonhole principle.
\item Looking at all edges from one vertex is the key step that forces a repeated colour.
\item This is the first nontrivial case of a famous Ramsey-theory fact: in any two-colouring of the edges of $K_6$, there is a monochromatic triangle.
\end{takeaways}

\begin{problem}[A more general form of the binomial theorem]
The classical binomial theorem expands
\[
(1+x)^n
\]
when $n$ is a non-negative integer. In more advanced mathematics, a similar expansion can be used when $n$ is negative, fractional, or any real number, provided we interpret the right-hand side as an infinite power series.

\begin{enumerate}[label=\alph*)]
\item Show that for a non-negative integer $n$, the binomial expansion can be written as
\[
(1+x)^n
=
1+nx+\frac{n(n-1)}{2!}x^2+\frac{n(n-1)(n-2)}{3!}x^3+\cdots.
\]

\item Assuming the same pattern continues for more general values of $n$, generate the power-series expansions of:
\begin{enumerate}[label=\roman*)]
\item $\dfrac{1}{1-x}$
\item $\dfrac{1}{(1-x)^2}$
\item $\dfrac{1}{(1+x)^2}$
\item $\sqrt{1+x}$.
\end{enumerate}

\item Verify, using your expansions in parts (b)(i) and (b)(ii), that
\[
\frac{d}{dx}\left(\frac{1}{1-x}\right)=\frac{1}{(1-x)^2},
\]
assuming that the power series may be differentiated term-by-term.
\end{enumerate}

\noindent\textit{Enrichment note:} This problem is outside the HSC Mathematics Extension~1 syllabus. It connects the classical binomial theorem to infinite power series, Taylor series, and later to Gamma-function ideas that allow binomial coefficients to be extended beyond whole numbers.
\end{problem}

\begin{hint}
In part (a), rewrite the usual binomial coefficients in factorial form. In part (b), substitute values such as $n=-1$, $n=-2$, and $n=\tfrac12$ into the general pattern.
\end{hint}

\begin{solution}
\textbf{a)} For a non-negative integer $n$, the usual binomial theorem gives
\[
(1+x)^n=\sum_{r=0}^{n}\binom{n}{r}x^r.
\]
Now
\[
\binom{n}{r}=\frac{n!}{r!(n-r)!}
=\frac{n(n-1)(n-2)\cdots (n-r+1)}{r!},
\]
so the expansion may be written as
\[
(1+x)^n
=
1+nx+\frac{n(n-1)}{2!}x^2+\frac{n(n-1)(n-2)}{3!}x^3+\cdots.
\]

\textbf{b)(i)} Take $n=-1$:
\[
(1-x)^{-1}=1+x+x^2+x^3+\cdots.
\]
So
\[
\boxed{\frac{1}{1-x}=1+x+x^2+x^3+\cdots.}
\]

\textbf{b)(ii)} Take $n=-2$:
\[
(1-x)^{-2}=1+2x+3x^2+4x^3+\cdots.
\]
So
\[
\boxed{\frac{1}{(1-x)^2}=1+2x+3x^2+4x^3+\cdots.}
\]

\textbf{b)(iii)} Replace $x$ by $-x$ in part (ii):
\[
\boxed{\frac{1}{(1+x)^2}=1-2x+3x^2-4x^3+\cdots.}
\]

\textbf{b)(iv)} Take $n=\tfrac12$:
\[
(1+x)^{1/2}
=1+\frac12x+\frac{\frac12(-\frac12)}{2!}x^2+\frac{\frac12(-\frac12)(-\frac32)}{3!}x^3+\cdots.
\]
So
\[
\boxed{
\sqrt{1+x}
=1+\frac12x-\frac18x^2+\frac1{16}x^3-\frac{5}{128}x^4+\cdots
}.
\]

\textbf{c)} From part (b)(i),
\[
\frac{1}{1-x}=1+x+x^2+x^3+\cdots.
\]
Differentiating term-by-term gives
\[
\frac{d}{dx}\left(\frac{1}{1-x}\right)=1+2x+3x^2+4x^3+\cdots.
\]
But this is exactly the expansion found in part (b)(ii), so
\[
\boxed{
\frac{d}{dx}\left(\frac{1}{1-x}\right)=\frac{1}{(1-x)^2}.
}
\]
\end{solution}

\begin{takeaways}
\item The usual finite binomial theorem is the beginning of a much more general power-series expansion.
\item For any real number $n$, and for $|x|<1$,
\[
(1+x)^n
=
\sum_{r=0}^{\infty}\binom{n}{r}x^r
=
1+nx+\frac{n(n-1)}{2!}x^2+\frac{n(n-1)(n-2)}{3!}x^3+\cdots.
\]
\item When $n$ is not a non-negative integer, the binomial expansion usually becomes an infinite Taylor-style series rather than a finite polynomial.
\item The generalized coefficient may be written using Gamma functions:
\[
\binom{n}{r}
=
\frac{\Gamma(n+1)}{\Gamma(r+1)\Gamma(n-r+1)},
\]
which extends the factorial formula to non-integer values of $n$.
\item This topic is outside HSC Mathematics Extension~1, but it is excellent enrichment for students interested in university mathematics.
\end{takeaways}

\begin{problem}[Isomers --- a possible computing project]
Because of branching, it is difficult to count the number of possible hydrocarbons with, say, $20$ carbon atoms.

For example, for butane $\mathrm{C}_4\mathrm{H}_{10}$ there are two carbon chains, and for pentane $\mathrm{C}_5\mathrm{H}_{12}$ there are three carbon chains.

Investigate the number of possible carbon chains for hydrocarbons with increasing numbers of carbon atoms.

\begin{enumerate}[label=\alph*)]
\item Explain why this is a counting problem about \emph{connected graphs} with no cycles, where each carbon atom has degree at most $4$.
\item Find the numbers of possible carbon chains for $n=1,2,3,4,5,6,7$ carbon atoms.
\item If you can write code, design a program to generate all possible carbon chains for a given $n$ and remove duplicates up to relabelling.
\end{enumerate}

\noindent\textit{Enrichment note:} This is a hard problem. Writing code is strongly recommended. It is outside the HSC Mathematics Extension~1 syllabus, but it is an excellent project linking combinatorics, graph theory, chemistry, and computing.
\end{problem}

\begin{hint}
Model each carbon chain as a tree: vertices are carbon atoms and edges are carbon-carbon bonds. Two drawings represent the same chain if they are the same up to relabelling of the vertices.
\end{hint}

\begin{solution}
\textbf{a)} A carbon chain can be modelled by a graph in which:
\begin{itemize}[leftmargin=*]
  \item each vertex represents a carbon atom
  \item each edge represents a carbon-carbon bond.
\end{itemize}

For an alkane chain, the carbon skeleton is connected and has no cycles, so it is a tree. Also, each carbon atom can make at most $4$ bonds, so each vertex has degree at most $4$.

\textbf{b)} The numbers of possible carbon chains for small $n$ are:
\[
\begin{array}{c|ccccccc}
n & 1 & 2 & 3 & 4 & 5 & 6 & 7\\ \hline
\text{number of chains} & 1 & 1 & 1 & 2 & 3 & 5 & 9
\end{array}
\]

These are the counts of non-isomorphic trees on $n$ vertices with maximum degree at most $4$. For small $n$, the degree restriction does not yet exclude any tree.

\textbf{c)} A practical computational approach is:
\begin{itemize}[leftmargin=*]
  \item generate trees on $n$ vertices by adding one new vertex at a time
  \item reject any graph with a vertex of degree more than $4$
  \item identify graphs that are the same up to relabelling by converting them to a canonical form
  \item count the distinct canonical forms.
\end{itemize}
\end{solution}

\begin{remark}[Python idea]
The following Python snippet generates non-isomorphic trees on $n$ vertices by adding one new vertex at a time and keeping one canonical representative of each isomorphism class. It is suitable for small values such as $n\le 7$.
\end{remark}

\begin{lstlisting}[language=Python,numbers=left,basicstyle=\ttfamily\small,keywordstyle=\color{bookpurple}\bfseries,commentstyle=\color{bookred},stringstyle=\color{blue!60!black},showstringspaces=false,breaklines=true]
from itertools import permutations


def canonical_form(adj):
    n = len(adj)
    best = None
    for perm in permutations(range(n)):
        bits = []
        for i in range(n):
            for j in range(i + 1, n):
                bits.append("1" if perm[j] in adj[perm[i]] else "0")
        code = "".join(bits)
        if best is None or code < best:
            best = code
    return best


def generate_trees(n):
    if n == 1:
        return [tuple([frozenset()])]

    previous = generate_trees(n - 1)
    seen = {}

    for tree in previous:
        base = [set(neighbours) for neighbours in tree]
        for v in range(n - 1):
            if len(base[v]) >= 4:
                continue

            graph = [set(neighbours) for neighbours in base] + [set()]
            graph[v].add(n - 1)
            graph[n - 1].add(v)

            key = canonical_form(graph)
            seen[key] = tuple(frozenset(x) for x in graph)

    return list(seen.values())


for n in range(1, 8):
    print(n, len(generate_trees(n)))
\end{lstlisting}

\begin{takeaways}
\item This problem is better viewed as a graph-theory and computing project than as a routine exam question.
\item \textbf{Combinatorial explosion} means the number of possible structures grows extremely quickly as $n$ increases, so direct listing soon becomes impractical.
\item Because of this explosion, writing code is often the most realistic way to investigate the problem.
\item This is outside the HSC Mathematics Extension~1 syllabus, but it is an excellent example of how combinatorics connects to chemistry and computer science.
\end{takeaways}
