\documentclass[12pt]{article}

% Core math packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{booktabs}

% Custom styles
\usepackage{styles/dl101-colors}
\usepackage{styles/dl101-hints}
\usepackage{styles/dl101-hsc-problems}

\geometry{
  a4paper,
  total={170mm,257mm},
  left=20mm,
  top=20mm,
}

\author{Vu Hung Nguyen}
\title{HSC Math Extension 2: The Nature of Proof}
\date{}

\begin{document}

\maketitle

\section{Introduction}

\subsection{Project Overview}
This booklet presents a comprehensive collection of proof problems for HSC Mathematics Extension~2 students. The 30 carefully selected problems cover all essential proof techniques: Direct Proof, Proof by Contradiction, Mathematical Induction, and Proof by Cases. Topics include divisibility, irrationality, modular arithmetic, parity, and logical reasoning.

The collection is divided into two parts:
\begin{itemize}[leftmargin=*]
  \item \textbf{Part 1 (15 problems):} Detailed step-by-step solutions showing complete reasoning, algebraic manipulation, and justification of each step. Each solution is designed to fit within one page, modeling clear and concise proof writing.
  \item \textbf{Part 2 (15 problems):} Strategic hints and solution sketches using flipped hint environments, encouraging independent problem-solving while providing guidance on key steps.
\end{itemize}

\subsection{Target Audience}
This booklet is designed for HSC Mathematics Extension~2 students who want to:
\begin{itemize}[leftmargin=*]
  \item Master fundamental proof techniques required for the HSC examination
  \item Develop rigorous mathematical reasoning and proof-writing skills
  \item Build confidence with both standard and non-standard proof problems
  \item Understand connections between different proof methods
  \item Practice problems at Easy, Medium, and Hard difficulty levels
\end{itemize}

Teachers and tutors will also find this collection valuable for:
\begin{itemize}[leftmargin=*]
  \item Demonstrating worked examples with clear pedagogical progression
  \item Selecting problems at appropriate difficulty levels for students
  \item Providing comprehensive coverage of HSC Extension~2 proof topics
  \item Preparing students for examination standards in mathematical proof
\end{itemize}

\subsection{How to Use This Booklet}
\textbf{For Students:}
\begin{itemize}[leftmargin=*]
  \item Begin with the Proof Primer below to review fundamental concepts and techniques.
  \item Work through Part~1 problems first, attempting each problem before reading the detailed solution.
  \item Compare your proof structure and reasoning with the provided solutions.
  \item Pay attention to the ``Takeaways'' section after each solution---these highlight key techniques and common pitfalls.
  \item Move to Part~2 problems, using the upside-down hints only after making a genuine attempt.
  \item Revisit challenging problems after a few days to reinforce understanding and technique.
\end{itemize}

\textbf{For Tutors and Teachers:}
\begin{itemize}[leftmargin=*]
  \item Use Part~1 problems as worked examples in lessons, highlighting proof structure and key techniques.
  \item Assign Part~2 problems for homework or practice, encouraging students to attempt problems before revealing hints.
  \item Select problems by proof technique or topic to target specific areas of the syllabus.
  \item Use the variety of difficulty levels (Easy, Medium, Hard) to differentiate instruction for different student abilities.
  \item Note Problem 14 (Part 1) as the ``Induction Crossover'' demonstrating overlap with HSC-Induction collection.
\end{itemize}

\subsection{Proof Primer}

\subsubsection*{Essential Number Theory Concepts}

\textbf{1. Divisibility:} $a | b$ (read ``$a$ divides $b$'') means there exists an integer $k$ such that $b = ak$.

\textbf{2. Division Algorithm:} For integers $a$ and $b > 0$, there exist unique integers $q$ (quotient) and $r$ (remainder) such that:
\[ a = bq + r, \quad 0 \le r < b \]
This is essential for case-based proofs (e.g., classifying integers as $3k$, $3k+1$, or $3k+2$).

\textbf{3. Modular Arithmetic:} $a \equiv b \pmod{m}$ (read ``$a$ is congruent to $b$ modulo $m$'') means $m | (a-b)$. Equivalently, $a$ and $b$ have the same remainder when divided by $m$.

\textbf{Key Properties:}
\begin{itemize}
    \item If $a \equiv b \pmod{m}$ and $c \equiv d \pmod{m}$, then $a+c \equiv b+d \pmod{m}$
    \item If $a \equiv b \pmod{m}$ and $c \equiv d \pmod{m}$, then $ac \equiv bd \pmod{m}$
    \item If $a \equiv b \pmod{m}$, then $a^n \equiv b^n \pmod{m}$ for any positive integer $n$
\end{itemize}

\textbf{4. Parity:} An integer is \textit{even} if it can be written as $2k$ for some integer $k$. An integer is \textit{odd} if it can be written as $2k+1$ for some integer $k$.

\textbf{5. Prime Numbers:} An integer $p > 1$ is prime if its only positive divisors are 1 and $p$. Key property: If prime $p$ divides $ab$, then $p | a$ or $p | b$.

\subsubsection*{Essential Proof Techniques}

\textbf{1. Direct Proof:} Start with given information and use logical steps to reach the desired conclusion.
\begin{itemize}
    \item Example: To prove $a | b$ and $b | c$ implies $a | c$, write $b = ak_1$ and $c = bk_2$, then $c = (ak_1)k_2 = a(k_1k_2)$.
\end{itemize}

\textbf{2. Proof by Contradiction (Reductio ad Absurdum):}
\begin{itemize}
    \item Assume the negation of what you want to prove
    \item Use logical reasoning to derive a statement that contradicts a known fact, the hypothesis, or a mathematical truth
    \item Conclude that the assumption must be false, hence the original statement is true
    \item Example: Proving $\sqrt{2}$ is irrational by assuming $\sqrt{2} = \frac{p}{q}$ and deriving a contradiction
\end{itemize}

\textbf{3. Mathematical Induction:} To prove $P(n)$ is true for all integers $n \geq n_0$:
\begin{itemize}
    \item \textit{Base Case:} Prove $P(n_0)$ is true
    \item \textit{Inductive Hypothesis:} Assume $P(k)$ is true for some arbitrary $k \geq n_0$
    \item \textit{Inductive Step:} Prove $P(k+1)$ is true using the assumption that $P(k)$ is true
    \item \textit{Conclusion:} By the principle of mathematical induction, $P(n)$ is true for all $n \geq n_0$
\end{itemize}

\textbf{4. Proof by Cases:} When the hypothesis can be divided into distinct cases:
\begin{itemize}
    \item Identify all possible cases (ensure they are exhaustive and mutually exclusive)
    \item Prove the statement for each case separately
    \item Conclude that the statement holds in all cases
    \item Example: Proving $n^2 \equiv 0$ or $1 \pmod{4}$ by checking cases $n$ even and $n$ odd
\end{itemize}

\textbf{5. Biconditional Proof (If and Only If):} To prove ``$P$ if and only if $Q$'' (written $P \iff Q$):
\begin{itemize}
    \item Prove the forward direction: $P \implies Q$
    \item Prove the reverse direction: $Q \implies P$
    \item Both directions must be proven independently
\end{itemize}

\textbf{6. Inequality Proofs:} Common strategies for proving inequalities:
\begin{itemize}
    \item \textit{Algebraic manipulation:} Transform the inequality to an obviously true statement (show $A \geq B$ by proving $A - B \geq 0$)
    \item \textit{AM-GM Inequality:} For non-negative numbers: $\frac{a_1 + \cdots + a_n}{n} \geq \sqrt[n]{a_1 \cdots a_n}$
    \item \textit{Cauchy-Schwarz:} $(a_1b_1 + \cdots + a_nb_n)^2 \leq (a_1^2 + \cdots + a_n^2)(b_1^2 + \cdots + b_n^2)$
    \item \textit{Induction:} For inequalities involving $n$, establish base case and inductive step
\end{itemize}

\subsubsection*{Induction Variants}

Beyond the standard induction framework, several specialized variants frequently appear:

\textbf{Induction for Series:} Proving closed-form formulas for sums (e.g., $\sum_{k=1}^n k^2 = \frac{n(n+1)(2n+1)}{6}$). In the inductive step, add the $(k+1)$-th term to both sides of $P(k)$ and simplify algebraically.

\textbf{Induction for Divisibility:} Proving statements like ``$n | f(n)$ for all $n \geq n_0$''. Use the inductive hypothesis to express $f(k+1)$ in terms of $f(k)$ and factor to show divisibility.

\textbf{Induction for Inequalities:} Proving statements like ``$2^n > n^2$ for $n \geq 5$''. The inductive step requires showing $f(k+1) > g(k+1)$ using the assumption $f(k) > g(k)$, often with careful estimation.

\textbf{Induction for Recursive Sequences:} For recursively defined sequences $a_{n+1} = f(a_n)$, prove properties by expressing $a_{k+1}$ using the recurrence and applying the inductive hypothesis to $a_k$.

\textbf{Strong Induction:} Assume $P(n_0), P(n_0+1), \ldots, P(k)$ all true, then prove $P(k+1)$. Essential when the $(k+1)$-th case depends on multiple previous cases.

\subsubsection*{Common Proof Strategies}

\textbf{Contrapositive:} To prove $P \implies Q$, instead prove $\neg Q \implies \neg P$ (logically equivalent).

\textbf{Consecutive Integers:} Among $k$ consecutive integers, exactly one is divisible by $k$. Among 2 consecutive integers, exactly one is even.

\textbf{Irrationality Proofs:} Standard template for proving $\sqrt{n}$ irrational (where $n$ is not a perfect square):
\begin{enumerate}
    \item Assume $\sqrt{n} = \frac{p}{q}$ in lowest terms ($\gcd(p,q) = 1$)
    \item Square both sides: $n = \frac{p^2}{q^2}$, so $p^2 = nq^2$
    \item Show $n | p^2$ implies $n | p$ (using prime factorization if $n$ is prime)
    \item Substitute $p = nk$, derive $q^2 = nk^2$, showing $n | q$
    \item Contradict $\gcd(p,q) = 1$
\end{enumerate}

\newpage

\section{Part 1: Problems with Detailed Solutions}

\subsection*{Easy Problems (5 problems)}

\input{problems/part1-01}
\newpage

\input{problems/part1-02}
\newpage

\input{problems/part1-03}
\newpage

\input{problems/part1-04}
\newpage

\input{problems/part1-05}
\newpage

\subsection*{Medium Problems (5 problems)}

\input{problems/part1-06}
\newpage

\input{problems/part1-07}
\newpage

\input{problems/part1-08}
\newpage

\input{problems/part1-09}
\newpage

\input{problems/part1-10}
\newpage

\subsection*{Hard Problems (5 problems)}

\input{problems/part1-11}
\newpage

\input{problems/part1-12}
\newpage

\input{problems/part1-13}
\newpage

\input{problems/part1-14}
\newpage

\input{problems/part1-15}
\newpage

\section{Part 2: Problems with Hints and Sketches}

\textit{The following problems provide strategic hints and solution sketches. Attempt each problem independently before consulting the hint. The sketches outline key steps but leave details for you to complete---use them to check your approach and fill in any gaps in your reasoning.}

\vspace{1cm}

\subsection*{Easy Problems (5 problems)}

\input{problems/part2-01}
\newpage

\input{problems/part2-02}
\newpage

\input{problems/part2-03}
\newpage

\input{problems/part2-04}
\newpage

\input{problems/part2-05}
\newpage

\subsection*{Medium Problems (5 problems)}

\input{problems/part2-06}
\newpage

\input{problems/part2-07}
\newpage

\input{problems/part2-08}
\newpage

\input{problems/part2-09}
\newpage

\input{problems/part2-10}
\newpage

\subsection*{Hard Problems (5 problems)}

\input{problems/part2-11}
\newpage

\input{problems/part2-12}
\newpage

\input{problems/part2-13}
\newpage

\input{problems/part2-14}
\newpage

\input{problems/part2-15}

\vspace{2em}
\noindent\textbf{Contact Information:}\\[0.5em]
LinkedIn: \href{https://www.linkedin.com/in/nguyenvuhung/}{https://www.linkedin.com/in/nguyenvuhung/}\\[0.3em]
GitHub: \href{https://github.com/vuhung16au/}{https://github.com/vuhung16au/}\\[0.3em]
Repository: 
\href{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Proofs}
{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Proofs}


\end{document}
