\documentclass[12pt]{article}

% Basic packages
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage[british]{babel}

% Apply DL101 styles (adapted for article class)
\usepackage{styles/dl101-colors}
\usepackage{styles/dl101-fonts}
\usepackage{styles/dl101-base}

% Create dummy chapter counter and commands for compatibility (article class doesn't have chapters)
\newcounter{chapter}
\renewcommand{\thechapter}{}
\newcommand{\chaptermark}[1]{}  % Dummy chaptermark for header-footer compatibility

\usepackage{styles/dl101-boxes}
\usepackage{styles/dl101-theorems-article}
\usepackage{styles/dl101-header-footer}

% Note: Boxes are defined with chapter numbering, but since we're in article class,
% the chapter counter is empty, so numbering will work correctly within sections

% TikZ libraries for diagrams
\usetikzlibrary{intersections, calc, shapes.geometric, arrows.meta, positioning}

% Algorithm support (for pseudocode)
\usepackage{algorithm}
\usepackage{algpseudocode}

\author{Vu Hung Nguyen}
\title{Rational Numbers: Everywhere but Nowhere\\From Ancient Mathematics to Modern Cryptography}
\date{\today}

\begin{document}

\maketitle

% Abstract
\input{sections/00-abstract}

\tableofcontents
\newpage

% Main content sections
\input{sections/01-introduction}
\input{sections/02-density}
\input{sections/03-countability}
\input{sections/04-measure-zero}
\input{sections/05-paradox}
\input{sections/06-probability}
\input{sections/07-crypto-overview}
\input{sections/08-rng}
\input{sections/09-lattice-crypto}
\input{sections/10-wiener-attack}
\input{sections/11-problems}
\input{sections/12-conclusion}

% References
\begin{thebibliography}{99}
\bibitem{chalktalk}
Chalk Talk, ``The Rational Numbers Are Not So 'Rational' | Everywhere but Nowhere,'' 
YouTube video, \url{https://www.youtube.com/watch?v=G8dLWja-bY0}

\bibitem{wiener1990}
M. J. Wiener, ``Cryptanalysis of short RSA secret exponents,'' 
\textit{IEEE Transactions on Information Theory}, vol. 36, no. 3, pp. 553--558, 1990.

\bibitem{hardy2008}
G. H. Hardy and E. M. Wright, \textit{An Introduction to the Theory of Numbers}, 
6th ed. Oxford University Press, 2008.

\bibitem{rosen2010}
K. H. Rosen, \textit{Elementary Number Theory and Its Applications}, 
6th ed. Pearson, 2010.

\bibitem{regev2009}
O. Regev, ``On lattices, learning with errors, random linear codes, and cryptography,'' 
\textit{Journal of the ACM}, vol. 56, no. 6, pp. 1--40, 2009.

\bibitem{peikert2016}
C. Peikert, ``A decade of lattice cryptography,'' 
\textit{Foundations and Trends in Theoretical Computer Science}, vol. 10, no. 4, pp. 283--424, 2016.
\end{thebibliography}

\end{document}

