\begin{problem}[Vector to Cartesian Line Equation]
What is the Cartesian equation of the line $\mathbf{r} = \begin{pmatrix} 1 \\ 3 \end{pmatrix} + \lambda \begin{pmatrix} 2 \\ 4 \end{pmatrix}$?
\end{problem}

\begin{hint}
Eliminate the parameter $\lambda$ by expressing it from one equation and substituting into the other.
\end{hint}

\begin{solution}[Sketch]
Write parametric equations: $x = 1 + 2\lambda$, $y = 3 + 4\lambda$. From the first equation, $\lambda = \frac{x-1}{2}$. Substitute into second: $y = 3 + 4 \cdot \frac{x-1}{2} = 3 + 2(x-1) = 2x + 1$. Therefore $y = 2x + 1$ or $2x - y + 1 = 0$.
\end{solution}

\vspace{1em}

\begin{problem}[Sketch 3D Helix]
Sketch the curve described by $\mathbf{r}(t) = -5\cos t\,\mathbf{i} + 5\sin t\,\mathbf{j} + t\,\mathbf{k}$ for $0 \le t \le 4\pi$.
\end{problem}

\begin{hint}
The $x$ and $y$ components trace a circle while $z$ increases linearly with $t$.
\end{hint}

\begin{solution}[Sketch]
In the $xy$-plane, $x^2 + y^2 = 25\cos^2 t + 25\sin^2 t = 25$, which is a circle of radius 5. As $t$ increases from 0 to $4\pi$, the point moves around the circle twice while rising from $z=0$ to $z=4\pi$. This creates a helix spiraling upward around the $z$-axis.
\end{solution}

\vspace{1em}

\begin{problem}[Angle Between 3D Vectors]
Find the angle between vectors $\mathbf{a} = \begin{pmatrix} 2 \\ 0 \\ 4 \end{pmatrix}$ and $\mathbf{b} = \begin{pmatrix} -3 \\ 1 \\ 2 \end{pmatrix}$ to the nearest degree.
\end{problem}

\begin{hint}
Use $\cos \theta = \frac{\mathbf{a} \cdot \mathbf{b}}{|\mathbf{a}||\mathbf{b}|}$.
\end{hint}

\begin{solution}[Sketch]
$\mathbf{a} \cdot \mathbf{b} = (2)(-3) + (0)(1) + (4)(2) = -6 + 8 = 2$. $|\mathbf{a}| = \sqrt{4 + 16} = \sqrt{20} = 2\sqrt{5}$, $|\mathbf{b}| = \sqrt{9 + 1 + 4} = \sqrt{14}$. Therefore $\cos \theta = \frac{2}{2\sqrt{5}\sqrt{14}} = \frac{1}{\sqrt{70}}$, so $\theta = \cos^{-1}(1/\sqrt{70}) \approx 83^\circ$.
\end{solution}

\vspace{1em}

\begin{problem}[Unit Vector in Direction]
Given $\mathbf{u} = 2\mathbf{i} - 2\mathbf{j} - 5\mathbf{k}$, find a vector parallel to $\mathbf{u}$ with length 5 units.
\end{problem}

\begin{hint}
Find the unit vector in the direction of $\mathbf{u}$, then scale by 5.
\end{hint}

\begin{solution}[Sketch]
$|\mathbf{u}| = \sqrt{4 + 4 + 25} = \sqrt{33}$. Unit vector: $\hat{\mathbf{u}} = \frac{1}{\sqrt{33}}(2\mathbf{i} - 2\mathbf{j} - 5\mathbf{k})$. Required vector: $5\hat{\mathbf{u}} = \frac{5}{\sqrt{33}}(2\mathbf{i} - 2\mathbf{j} - 5\mathbf{k})$.
\end{solution}

\vspace{1em}

\begin{problem}[Distance to Plane and Axis]
Find the distance from point $P(1, 4, 3)$ to: (a) the $xz$-plane, (b) the $y$-axis.
\end{problem}

\begin{hint}
(a) Distance to $xz$-plane is the $y$-coordinate. (b) Distance to $y$-axis uses $x$ and $z$ coordinates.
\end{hint}

\begin{solution}[Sketch]
\textbf{Part (a): Distance from $P(1, 4, 3)$ to the $xz$-plane}

The $xz$-plane consists of all points where $y = 0$. The perpendicular distance from point $P(1, 4, 3)$ to this plane is simply the absolute value of the $y$-coordinate:
$$\text{Distance} = |y| = |4| = 4 \text{ units}$$

The closest point on the $xz$-plane to $P$ is $(1, 0, 3)$, which we can verify: $\sqrt{(1-1)^2 + (4-0)^2 + (3-3)^2} = 4$.

\vspace{0.5em}

\textbf{Part (b): Distance from $P(1, 4, 3)$ to the $y$-axis}

The $y$-axis consists of all points of the form $(0, y, 0)$ where $y \in \mathbb{R}$. To find the closest point on the $y$-axis to $P(1, 4, 3)$, we note that this point has the same $y$-coordinate as $P$. Therefore, the closest point is $(0, 4, 0)$.

The distance from $P(1, 4, 3)$ to $(0, 4, 0)$ is:
$$\text{Distance} = \sqrt{(1-0)^2 + (4-4)^2 + (3-0)^2} = \sqrt{1 + 0 + 9} = \sqrt{10} \text{ units}$$

\textit{Geometric interpretation:} The distance to the $y$-axis is the radius of the circle in a plane perpendicular to the $y$-axis passing through $P$. This radius involves only the $x$ and $z$ coordinates: $\sqrt{x^2 + z^2} = \sqrt{1^2 + 3^2} = \sqrt{10}$.
\end{solution}

\vspace{1em}

\begin{problem}[Unit Vector Perpendicular to Two Vectors]
Find a unit vector perpendicular to both $\mathbf{u} = \mathbf{i} + \mathbf{j}$ and $\mathbf{v} = \mathbf{i} + \mathbf{k}$, where $\mathbf{i}$, $\mathbf{j}$, and $\mathbf{k}$ are the standard unit vectors along the $x$-, $y$-, and $z$-axes respectively.
\end{problem}

\begin{hint}
Use the cross product $\mathbf{u} \times \mathbf{v}$, then normalize.
\end{hint}

\begin{solution}[Sketch]
$\mathbf{u} \times \mathbf{v} = \begin{vmatrix} \mathbf{i} & \mathbf{j} & \mathbf{k} \\ 1 & 1 & 0 \\ 1 & 0 & 1 \end{vmatrix} = \mathbf{i}(1) - \mathbf{j}(1) + \mathbf{k}(-1) = \mathbf{i} - \mathbf{j} - \mathbf{k}$. Magnitude: $|\mathbf{w}| = \sqrt{1 + 1 + 1} = \sqrt{3}$. Unit vector: $\frac{1}{\sqrt{3}}(\mathbf{i} - \mathbf{j} - \mathbf{k})$.
\end{solution}

\vspace{1em}

\begin{problem}[Point on Line]
Consider the line passing through points $A(0, 2, 1)$ and $B(2, 7, 4)$. Find the values of $a$ and $b$ such that the point $P(a, 1, b)$ also lies on this line.
\end{problem}

\begin{hint}
Find parametric equations using direction vector $(2, 5, 3)$, then solve for the parameter when $y = 1$.
\end{hint}

\begin{solution}[Sketch]
\textbf{Step 1: Find the direction vector}

The direction vector of the line is:
$$\vec{AB} = B - A = (2, 7, 4) - (0, 2, 1) = (2, 5, 3)$$

\vspace{0.3em}

\textbf{Step 2: Write parametric equations}

Using point $A(0, 2, 1)$ and direction vector $(2, 5, 3)$, the parametric equations of the line are:
\begin{align*}
x &= 0 + 2t = 2t \\
y &= 2 + 5t \\
z &= 1 + 3t
\end{align*}
where $t$ is the parameter.

\vspace{0.3em}

\textbf{Step 3: Find the parameter when $y = 1$}

Since point $P$ has $y$-coordinate equal to 1, we substitute into the $y$ equation:
$$1 = 2 + 5t \Rightarrow 5t = -1 \Rightarrow t = -\frac{1}{5}$$

\vspace{0.3em}

\textbf{Step 4: Find $a$ and $b$}

Substituting $t = -\frac{1}{5}$ into the $x$ and $z$ equations:
$$a = x = 2\left(-\frac{1}{5}\right) = -\frac{2}{5}$$
$$b = z = 1 + 3\left(-\frac{1}{5}\right) = 1 - \frac{3}{5} = \frac{2}{5}$$

Therefore, $a = -\dfrac{2}{5}$ and $b = \dfrac{2}{5}$, giving the point $P\left(-\dfrac{2}{5}, 1, \dfrac{2}{5}\right)$.

\vspace{0.3em}

\textit{Verification:} We can verify by checking that this point satisfies the line equation with $t = -\frac{1}{5}$.
\end{solution}

\vspace{1em}

\begin{problem}[Equal Magnitude Perpendicular Vectors]
Prove that if $\mathbf{u}$ and $\mathbf{v}$ are non-zero vectors of equal magnitude, then $\mathbf{u} - \mathbf{v}$ is perpendicular to $\mathbf{u} + \mathbf{v}$.
\end{problem}

\begin{hint}
Show that $(\mathbf{u} - \mathbf{v}) \cdot (\mathbf{u} + \mathbf{v}) = 0$.
\end{hint}

\begin{solution}[Sketch]
To prove perpendicularity, we need to show that the dot product equals zero.

Expand the dot product using distributivity:
\begin{align*}
(\mathbf{u} - \mathbf{v}) \cdot (\mathbf{u} + \mathbf{v}) &= \mathbf{u} \cdot \mathbf{u} + \mathbf{u} \cdot \mathbf{v} - \mathbf{v} \cdot \mathbf{u} - \mathbf{v} \cdot \mathbf{v}
\end{align*}

Since the dot product is commutative ($\mathbf{u} \cdot \mathbf{v} = \mathbf{v} \cdot \mathbf{u}$), the middle terms cancel:
\begin{align*}
&= \mathbf{u} \cdot \mathbf{u} - \mathbf{v} \cdot \mathbf{v} \\
&= |\mathbf{u}|^2 - |\mathbf{v}|^2
\end{align*}

Given that $|\mathbf{u}| = |\mathbf{v}|$, we have:
$$|\mathbf{u}|^2 - |\mathbf{v}|^2 = 0$$

Therefore, $(\mathbf{u} - \mathbf{v}) \perp (\mathbf{u} + \mathbf{v})$.
\end{solution}

\begin{takeaways}
\item This result has a beautiful geometric interpretation: if vectors $\mathbf{u}$ and $\mathbf{v}$ have equal magnitude, then the vectors $\mathbf{u} + \mathbf{v}$ (the diagonal) and $\mathbf{u} - \mathbf{v}$ (the other diagonal) form perpendicular diagonals.

\item \textbf{Rectangle interpretation:} The vectors $\mathbf{u}$, $\mathbf{v}$, $\mathbf{u} + \mathbf{v}$, and $\mathbf{u} - \mathbf{v}$ form a \emph{rhombus} (not a rectangle in general), since $|\mathbf{u}| = |\mathbf{v}|$ makes all four sides equal length. The diagonals of this rhombus are perpendicular, which is what we just proved.

\item Special case: If $\mathbf{u} \perp \mathbf{v}$ \emph{and} $|\mathbf{u}| = |\mathbf{v}|$, then the rhombus becomes a \emph{square}, and the diagonals are both perpendicular and equal in length.

\item This identity is useful in many geometric proofs and demonstrates the power of algebraic vector manipulation to prove geometric relationships.
\end{takeaways}

\vspace{1em}

\begin{problem}[Direction Cosines]
Prove that $\cos^2\alpha + \cos^2\beta + \cos^2\gamma = 1$ for a position vector $\mathbf{r} = a\mathbf{i} + b\mathbf{j} + c\mathbf{k}$ making angles $\alpha, \beta, \gamma$ with the positive $x$-, $y$-, and $z$-axes respectively. (The quantities $\cos\alpha$, $\cos\beta$, and $\cos\gamma$ are called the \emph{direction cosines} of the vector $\mathbf{r}$.)
\end{problem}

\begin{hint}
Use dot products with unit vectors $\mathbf{i}, \mathbf{j}, \mathbf{k}$ to find the direction cosines.
\end{hint}

\begin{solution}[Sketch]
$\cos\alpha = \frac{a}{|\mathbf{r}|}$, $\cos\beta = \frac{b}{|\mathbf{r}|}$, $\cos\gamma = \frac{c}{|\mathbf{r}|}$. Therefore $\cos^2\alpha + \cos^2\beta + \cos^2\gamma = \frac{a^2 + b^2 + c^2}{|\mathbf{r}|^2} = \frac{|\mathbf{r}|^2}{|\mathbf{r}|^2} = 1$.
\end{solution}

\vspace{1em}

\begin{problem}[Line Intersection by Components]
Find the intersection point of lines $\mathbf{r} = \begin{pmatrix} 3 \\ -1 \\ 7 \end{pmatrix} + \lambda_1 \begin{pmatrix} 1 \\ 2 \\ 1 \end{pmatrix}$ and $\mathbf{r} = \begin{pmatrix} 3 \\ -6 \\ 2 \end{pmatrix} + \lambda_2 \begin{pmatrix} -2 \\ 1 \\ 3 \end{pmatrix}$.
\end{problem}

\begin{hint}
Equate components and solve the system for $\lambda_1$ and $\lambda_2$.
\end{hint}

\begin{solution}[Sketch]
Equating: $3 + \lambda_1 = 3 - 2\lambda_2$, $-1 + 2\lambda_1 = -6 + \lambda_2$, $7 + \lambda_1 = 2 + 3\lambda_2$. From first equation: $\lambda_1 = -2\lambda_2$. Substitute into second: $-1 - 4\lambda_2 = -6 + \lambda_2 \Rightarrow \lambda_2 = 1$, so $\lambda_1 = -2$. Check third equation: $7 - 2 = 5 = 2 + 3$ $\checkmark$. Intersection point: $(1, -5, 5)$.
\end{solution}

\vspace{1em}

\begin{problem}[Multiple Choice: Cartesian Equation]
What is the Cartesian equation of $\mathbf{r} = \begin{pmatrix} 1 \\ 3 \end{pmatrix} + \lambda \begin{pmatrix} -2 \\ 4 \end{pmatrix}$? 

Options: \\
A. $2y + x = 7$, \\
B. $y - 2x = -5$, \\
C. $y + 2x = 5$, \\ 
D. $2y - x = -1$.
\end{problem}

\begin{hint}
Eliminate $\lambda$ from $x = 1 - 2\lambda$ and $y = 3 + 4\lambda$.
\end{hint}

\begin{solution}[Sketch]
From $x = 1 - 2\lambda$, get $\lambda = \frac{1-x}{2}$. Substitute: $y = 3 + 4 \cdot \frac{1-x}{2} = 3 + 2(1-x) = 5 - 2x$, giving $y + 2x = 5$. Answer: C.
\end{solution}

\vspace{1em}

\begin{problem}[Closest Point on Line to Origin]
Find the point on line $\mathbf{r} = \begin{pmatrix} 1 \\ 4 \\ 6 \end{pmatrix} + \lambda \begin{pmatrix} 2 \\ 1 \\ 1 \end{pmatrix}$ closest to the origin. That is, find the value of $\lambda$ such that the distance from the point $\mathbf{r}(\lambda)$ to the origin $O(0, 0, 0)$ is minimized.
\end{problem}

\begin{hint}
The closest point occurs when $\mathbf{r}$ is perpendicular to the direction vector.
\end{hint}

\begin{solution}[Sketch]
Let $\mathbf{r} = \begin{pmatrix} 1 + 2\lambda \\ 4 + \lambda \\ 6 + \lambda \end{pmatrix}$. For closest point: $\mathbf{r} \cdot \begin{pmatrix} 2 \\ 1 \\ 1 \end{pmatrix} = 0$. This gives $2(1+2\lambda) + (4+\lambda) + (6+\lambda) = 0 \Rightarrow 6\lambda + 12 = 0 \Rightarrow \lambda = -2$. Point: $(-3, 2, 4)$.
\end{solution}

\vspace{1em}

\begin{problem}[Unit Vector Perpendicular to Two]
Find a unit vector perpendicular to $\begin{bmatrix} 3 \\ -2 \\ 1 \end{bmatrix}$ and $\begin{bmatrix} 1 \\ 4 \\ -1 \end{bmatrix}$.
\end{problem}

\begin{hint}
Compute the cross product and normalize the result.
\end{hint}

\begin{solution}[Sketch]
Cross product: $\mathbf{a} \times \mathbf{b} = \begin{bmatrix} (-2)(-1) - (1)(4) \\ (1)(1) - (3)(-1) \\ (3)(4) - (-2)(1) \end{bmatrix} = \begin{bmatrix} -2 \\ 4 \\ 14 \end{bmatrix}$. Magnitude: $\sqrt{4 + 16 + 196} = 6\sqrt{6}$. Unit vector: $\frac{1}{6\sqrt{6}}\begin{bmatrix} -2 \\ 4 \\ 14 \end{bmatrix} = \frac{1}{3\sqrt{6}}\begin{bmatrix} -1 \\ 2 \\ 7 \end{bmatrix}$.
\end{solution}

\vspace{1em}

\begin{problem}[Perpendicular Dot Product Proof]
Prove that for non-zero vectors $\mathbf{a}, \mathbf{b}$, the equation $(\mathbf{a} + \mathbf{b}) \cdot (\mathbf{a} + \mathbf{b}) = |\mathbf{a}|^2 + |\mathbf{b}|^2$ holds only if $\mathbf{a} \perp \mathbf{b}$.
\end{problem}

\begin{hint}
Expand the left side and compare with the right side.
\end{hint}

\begin{solution}[Sketch]
Expanding: $(\mathbf{a} + \mathbf{b}) \cdot (\mathbf{a} + \mathbf{b}) = |\mathbf{a}|^2 + 2\mathbf{a} \cdot \mathbf{b} + |\mathbf{b}|^2$. Comparing with RHS: $|\mathbf{a}|^2 + 2\mathbf{a} \cdot \mathbf{b} + |\mathbf{b}|^2 = |\mathbf{a}|^2 + |\mathbf{b}|^2 \Rightarrow 2\mathbf{a} \cdot \mathbf{b} = 0 \Rightarrow \mathbf{a} \cdot \mathbf{b} = 0$, which means $\mathbf{a} \perp \mathbf{b}$.
\end{solution}

\vspace{1em}
