% ============================================================
% Part 1 — Medium Worked Problems
% Topics: polynomial division, systematic factoring,
%         Vieta's with manipulation, double/triple roots
% ============================================================

\begin{problem}[Parity and Derivatives]
Suppose $P(x)$ is a polynomial. Prove or disprove:
\[
P(x)\text{ is even if and only if }P'(x)\text{ is odd.}
\]
\end{problem}

\begin{hintbox}
Prove each direction separately. For the reverse direction, examine $P(x)-P(-x)$.
\end{hintbox}

\begin{solution}
If $P$ is even, then $P(-x)=P(x)$. Differentiating gives
\[
-P'(-x)=P'(x),
\]
so
\[
P'(-x)=-P'(x).
\]
Hence $P'$ is odd.

Conversely, suppose $P'$ is odd, so $P'(-x)=-P'(x)$. Let
\[
F(x)=P(x)-P(-x).
\]
Then
\[
F'(x)=P'(x)+P'(-x)=0.
\]
So $F$ is constant. But $F(0)=0$, hence $F(x)=0$ for all $x$. Therefore
\[
P(x)=P(-x),
\]
so $P$ is even.

The statement is \textbf{true}.
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Differentiating a symmetry identity is often the quickest proof.
\item If a polynomial has derivative $0$, it must be constant.
\end{itemize}
\end{takeaways}

\begin{problem}[Behaviour Near Repeated Zeroes]
Let
\[
P(x)=(x+1)^3x^2(x-1)^2.
\]
Discuss the behaviour of $P(x)$ around $-\infty$, $+\infty$, $-1$, $0$ and $1$. Summarise it in a table and hence sketch the graph roughly.
\end{problem}

\begin{hintbox}
Use the degree and leading coefficient for end behaviour, and multiplicity for the local shape at each zero.
\end{hintbox}

\begin{solution}
The degree is $7$ and the leading coefficient is positive, so
\[
x\to-\infty \Rightarrow P(x)\to-\infty,\qquad
x\to+\infty \Rightarrow P(x)\to+\infty.
\]
The real zeroes are
\[
x=-1 \text{ (multiplicity }3),\quad x=0 \text{ (multiplicity }2),\quad x=1 \text{ (multiplicity }2).
\]
Hence
\[
\begin{array}{c|c}
\text{Point} & \text{Behaviour}\\ \hline
x\to-\infty & P(x)\to-\infty\\
x=-1 & \text{crosses with flattening}\\
x=0 & \text{touches and turns}\\
x=1 & \text{touches and turns}\\
x\to+\infty & P(x)\to+\infty
\end{array}
\]

\begin{center}
\begin{tikzpicture}
\begin{axis}[
  axis lines=middle,
  xmin=-2.2,xmax=2.2,
  ymin=-8,ymax=8,
  samples=300,
  width=0.8\linewidth,
  height=6cm,
  xtick={-1,0,1},
  ytick=\empty
]
\addplot[thick,bookpurple,domain=-2.2:2.2]{(x+1)^3*x^2*(x-1)^2};
\end{axis}
\end{tikzpicture}
\end{center}
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Odd multiplicity means the graph crosses the axis.
\item Even multiplicity means the graph touches and turns.
\end{itemize}
\end{takeaways}

\begin{problem}[Same Axes, Different Quadratics]
Sketch the following graphs on the same $xy$-diagram, using a different colour for each:
\[
P(x)=x^2,\qquad Q(x)=(x+1)(x-2),\qquad R(x)=(x-2)^2,\qquad S(x)=4-x^2.
\]
\end{problem}

\begin{hintbox}
Find the vertex and intercepts of each quadratic first.
\end{hintbox}

\begin{solution}
Useful features:
\[
P(x)=x^2 \text{ has vertex }(0,0),
\]
\[
Q(x)=x^2-x-2 \text{ has roots }-1,2 \text{ and vertex }\left(\frac12,-\frac94\right),
\]
\[
R(x)=(x-2)^2 \text{ has vertex }(2,0),
\]
\[
S(x)=4-x^2 \text{ has vertex }(0,4)\text{ and roots }\pm2.
\]

\begin{center}
\begin{tikzpicture}
\begin{axis}[
  axis lines=middle,
  xmin=-3.2,xmax=3.2,
  ymin=-3.5,ymax=5,
  samples=300,
  width=0.85\linewidth,
  height=7cm,
  legend style={draw=none,fill=none,at={(0.98,0.98)},anchor=north east}
]
\addplot[thick,blue,domain=-3.2:3.2]{x^2};
\addlegendentry{$x^2$}
\addplot[thick,red,domain=-3.2:3.2]{(x+1)*(x-2)};
\addlegendentry{$(x+1)(x-2)$}
\addplot[thick,teal!70!black,domain=-3.2:3.2]{(x-2)^2};
\addlegendentry{$(x-2)^2$}
\addplot[thick,orange!90!black,domain=-3.2:3.2]{4-x^2};
\addlegendentry{$4-x^2$}
\end{axis}
\end{tikzpicture}
\end{center}
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Related quadratics are easiest to compare on the same set of axes.
\item Vertex form and factor form both give quick sketching information.
\end{itemize}
\end{takeaways}

\begin{problem}[An Odd Cubic from Two Conditions]
$P(x)$ is an odd polynomial of degree $3$. It has $x+2$ as a factor, and when divided by $x-3$ the remainder is $2$. Find $P(x)$.
\end{problem}

\begin{hintbox}
An odd cubic has the form $ax^3+bx$.
\end{hintbox}

\begin{solution}
Let
\[
P(x)=ax^3+bx.
\]
Since $x+2$ is a factor,
\[
P(-2)=0 \Rightarrow -8a-2b=0 \Rightarrow 4a+b=0.
\]
Since the remainder on division by $x-3$ is $2$,
\[
P(3)=2 \Rightarrow 27a+3b=2.
\]
Using $b=-4a$,
\[
27a-12a=2 \Rightarrow 15a=2 \Rightarrow a=\frac{2}{15}.
\]
Then
\[
b=-\frac{8}{15}.
\]
So
\[
\boxed{P(x)=\frac{2}{15}x^3-\frac{8}{15}x=\frac{2}{15}x(x+2)(x-2).}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Symmetry conditions reduce the number of unknown coefficients.
\item Factor and remainder conditions become linear equations.
\end{itemize}
\end{takeaways}

\begin{problem}[Quadratic Roots and Symmetric Expressions]
If $\alpha$ and $\beta$ are the roots of
\[
x^2-5x+1=0,
\]
find:
\begin{enumerate}[label=(\roman*)]
\item $\left(\alpha+\frac{2}{\alpha}\right)\left(\beta+\frac{2}{\beta}\right)$
\item $\alpha^2+\beta^2$
\item $\alpha^3+\beta^3$
\end{enumerate}
\end{problem}

\begin{hintbox}
Use $\alpha+\beta=5$ and $\alpha\beta=1$.
\end{hintbox}

\begin{solution}
By Vieta,
\[
\alpha+\beta=5,\qquad \alpha\beta=1.
\]

\textbf{(ii)}
\[
\alpha^2+\beta^2=(\alpha+\beta)^2-2\alpha\beta=25-2=23.
\]

\textbf{(iii)}
\[
\alpha^3+\beta^3=(\alpha+\beta)^3-3\alpha\beta(\alpha+\beta)=125-15=110.
\]

\textbf{(i)}
\begin{align*}
\left(\alpha+\frac{2}{\alpha}\right)\left(\beta+\frac{2}{\beta}\right)
&=\alpha\beta+2\frac{\alpha}{\beta}+2\frac{\beta}{\alpha}+\frac{4}{\alpha\beta}\\
&=1+2(\alpha^2+\beta^2)+4\\
&=1+46+4=51.
\end{align*}

Hence
\[
\boxed{51,\quad 23,\quad 110.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Vieta is the fastest route for symmetric expressions in the roots.
\item Compute the simpler expressions first and reuse them.
\end{itemize}
\end{takeaways}

\begin{problem}[A Cubic Root Product Sum]
Let $\alpha,\beta,\gamma$ be the roots of
\[
x^3-5x^2+5x-1=0.
\]
Find
\begin{enumerate}[label=(\roman*)]
\item $\alpha^2\beta\gamma+\alpha\beta^2\gamma+\alpha\beta\gamma^2$
\item $\alpha^3\beta\gamma+\alpha\beta^3\gamma+\alpha\beta\gamma^3$
\end{enumerate}
\end{problem}

\begin{hintbox}
Factor out $\alpha\beta\gamma$. Then use $\alpha+\beta+\gamma$ and $\alpha^2+\beta^2+\gamma^2$.
\end{hintbox}

\begin{solution}
\textbf{(i)}
\[
\alpha^2\beta\gamma+\alpha\beta^2\gamma+\alpha\beta\gamma^2
=\alpha\beta\gamma(\alpha+\beta+\gamma).
\]
By Vieta,
\[
\alpha+\beta+\gamma=5,\qquad \alpha\beta\gamma=1.
\]
So the value is
\[
\boxed{5.}
\]

\textbf{(ii)}
\[
\alpha^3\beta\gamma+\alpha\beta^3\gamma+\alpha\beta\gamma^3
=\alpha\beta\gamma(\alpha^2+\beta^2+\gamma^2).
\]
Now
\[
\alpha^2+\beta^2+\gamma^2=(\alpha+\beta+\gamma)^2-2(\alpha\beta+\beta\gamma+\gamma\alpha).
\]
By Vieta for
\[
\alpha+\beta+\gamma=5,\qquad \alpha\beta+\beta\gamma+\gamma\alpha=5,\qquad \alpha\beta\gamma=1.
\]
So
\[
\alpha^2+\beta^2+\gamma^2=25-10=15.
\]
Hence
\[
\alpha^3\beta\gamma+\alpha\beta^3\gamma+\alpha\beta\gamma^3
=1\cdot 15
=\boxed{15.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Factor the target expression before applying Vieta.
\item Many cubic root-sum questions become one line after the right factorisation.
\item Symmetric root expressions are usually the ones that give a single definite answer.
\end{itemize}
\end{takeaways}

\begin{problem}[The Opposite Roots]
Consider the cubic polynomial
\[
P(x)=x^3-4x^2+kx+16,
\]
where $k$ is a real constant. It is known that the sum of two of the roots of $P(x)=0$ is zero.
\begin{enumerate}[label=(\roman*)]
\item Find the value of $k$.
\item Hence find all three roots of the polynomial.
\end{enumerate}
\end{problem}

\begin{hintbox}
If two roots sum to zero, write them as $\alpha$ and $-\alpha$. Then use the sum of roots first.
\end{hintbox}

\begin{solution}
Let the roots be $\alpha,\beta,\gamma$, with
\[
\alpha+\beta=0.
\]
So we may write the roots as
\[
\alpha,\ -\alpha,\ \gamma.
\]

\textbf{(i)} By Vieta, the sum of the roots is
\[
\alpha+(-\alpha)+\gamma=4,
\]
so
\[
\gamma=4.
\]
Hence $x=4$ is a root, and therefore
\[
P(4)=0.
\]
Substitute into the polynomial:
\[
4^3-4\cdot 4^2+4k+16=0.
\]
This gives
\[
64-64+4k+16=0 \Rightarrow 4k=-16 \Rightarrow \boxed{k=-4.}
\]

\textbf{(ii)} Now
\[
P(x)=x^3-4x^2-4x+16.
\]
Factor by grouping:
\[
P(x)=x^2(x-4)-4(x-4)=(x^2-4)(x-4).
\]
So
\[
P(x)=(x-2)(x+2)(x-4).
\]
Hence the three roots are
\[
\boxed{4,\ 2,\ -2.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item The phrase ``two roots sum to zero'' should immediately suggest a pair of opposite roots.
\item Once one root is identified, the factor theorem is often the fastest way to find the parameter.
\item Before using division, always check whether the cubic factors neatly by grouping.
\end{itemize}
\end{takeaways}

\begin{problem}[The Shifting Exponents]
Consider the polynomial
\[
P(x)=x^n+2x^{n-1}+k,
\]
where $n$ is an integer greater than $1$, and $k$ is a real constant.
\begin{enumerate}[label=(\roman*)]
\item If $n$ is odd, find the value of $k$ such that $(x+1)$ is a factor of $P(x)$.
\item Suppose instead that $n$ is even and $k=-3$. Find the remainder when $P(x)$ is divided by $x^2-1$.
\end{enumerate}
\end{problem}

\begin{hintbox}
Use $x=-1$ for part (i). For part (ii), write the remainder as $ax+b$ and use the roots $x=1$ and $x=-1$ of $x^2-1$.
\end{hintbox}

\begin{solution}
\textbf{(i)} If $(x+1)$ is a factor, then by the factor theorem
\[
P(-1)=0.
\]
So
\[
(-1)^n+2(-1)^{n-1}+k=0.
\]
Since $n$ is odd,
\[
(-1)^n=-1,\qquad (-1)^{n-1}=1.
\]
Hence
\[
-1+2+k=0 \Rightarrow 1+k=0,
\]
so
\[
\boxed{k=-1.}
\]

\textbf{(ii)} Now let $n$ be even and $k=-3$, so
\[
P(x)=x^n+2x^{n-1}-3.
\]
When dividing by $x^2-1$, the remainder has the form
\[
R(x)=ax+b.
\]
Since $x^2-1=(x-1)(x+1)$, we use $x=1$ and $x=-1$.

At $x=1$:
\[
P(1)=1+2-3=0,
\]
so
\[
R(1)=a+b=0.
\]

At $x=-1$: since $n$ is even, $(-1)^n=1$ and $(-1)^{n-1}=-1$, so
\[
P(-1)=1+2(-1)-3=-4.
\]
Thus
\[
R(-1)=-a+b=-4.
\]
Solving
\[
a+b=0,\qquad -a+b=-4
\]
gives
\[
a=2,\qquad b=-2.
\]
Therefore the remainder is
\[
\boxed{2x-2.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item When substituting $x=-1$, the parity of the exponent completely controls the sign.
\item For division by a quadratic, use a general linear remainder.
\item If the divisor factors nicely, substituting its roots is usually the fastest way to determine the remainder.
\end{itemize}
\end{takeaways}

\begin{problem}[The Parity of Powers]
Consider the polynomial
\[
P(x)=x^n+c^n,
\]
where $n$ is a positive integer and $c$ is a non-zero real constant.
\begin{enumerate}[label=(\roman*)]
\item By considering the cases where $n$ is odd and even separately, determine whether $(x+c)$ and $(x-c)$ are factors of $P(x)$.
\item Suppose $n$ is odd. Find an expression for the remainder when $P(x)$ is divided by $x^2-c^2$, in terms of $c$ and $n$.
\end{enumerate}
\end{problem}

\begin{hintbox}
Test $(x-c)$ using $P(c)$ and test $(x+c)$ using $P(-c)$. For part (ii), let the remainder be $Ax+B$.
\end{hintbox}

\begin{solution}
\textbf{(i)} To test $(x-c)$, evaluate
\[
P(c)=c^n+c^n=2c^n.
\]
Since $c\ne0$, this is not zero. So
\[
\boxed{(x-c)\text{ is never a factor}.}
\]

To test $(x+c)$, evaluate
\[
P(-c)=(-c)^n+c^n.
\]
If $n$ is even, then $(-c)^n=c^n$, so
\[
P(-c)=2c^n\ne0.
\]
If $n$ is odd, then $(-c)^n=-c^n$, so
\[
P(-c)=0.
\]
Therefore
\[
\boxed{(x+c)\text{ is a factor if and only if }n\text{ is odd}.}
\]

\textbf{(ii)} Now suppose $n$ is odd. When dividing by $x^2-c^2$, the remainder has the form
\[
R(x)=Ax+B.
\]
Since
\[
x^2-c^2=(x-c)(x+c),
\]
we use $x=c$ and $x=-c$.

At $x=-c$, part (i) gives
\[
P(-c)=0,
\]
so
\[
R(-c)=-Ac+B=0.
\]
Hence
\[
B=Ac.
\]

At $x=c$,
\[
P(c)=2c^n,
\]
so
\[
R(c)=Ac+B=2c^n.
\]
Substitute $B=Ac$:
\[
2Ac=2c^n.
\]
Since $c\ne0$,
\[
A=c^{n-1}.
\]
Then
\[
B=Ac=c^n.
\]
So the remainder is
\[
\boxed{R(x)=c^{n-1}x+c^n=c^{n-1}(x+c).}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Powers of $-c$ must always be split into even and odd cases.
\item Once the divisor factors, its roots give the fastest route to the linear remainder.
\item Part (ii) becomes easier because the result from part (i) gives one of the remainder equations immediately.
\end{itemize}
\end{takeaways}

\begin{problem}[The Cascading Powers]
The cubic equation
\[
x^3-3x^2+4x-5=0
\]
has roots $\alpha,\beta,\gamma$.
\begin{enumerate}[label=(\roman*)]
\item Find the exact value of $\alpha^2+\beta^2+\gamma^2$.
\item By noting that $P(\alpha)=0$, or otherwise, find the value of $\alpha^3+\beta^3+\gamma^3$.
\item Hence evaluate $\alpha^4+\beta^4+\gamma^4$.
\item Let
\[
S_k=\alpha^k+\beta^k+\gamma^k.
\]
Express $S_k$ in terms of $S_{k-1}$, $S_{k-2}$ and $S_{k-3}$.
\end{enumerate}
\end{problem}

\begin{hintbox}
Start with Vieta. For higher powers, rewrite the equation as a recurrence for powers of a root.
\end{hintbox}

\begin{solution}
By Vieta,
\[
\alpha+\beta+\gamma=3,\qquad
\alpha\beta+\beta\gamma+\gamma\alpha=4,\qquad
\alpha\beta\gamma=5.
\]

\textbf{(i)} Use
\[
\alpha^2+\beta^2+\gamma^2=(\alpha+\beta+\gamma)^2-2(\alpha\beta+\beta\gamma+\gamma\alpha).
\]
So
\[
\alpha^2+\beta^2+\gamma^2=3^2-2(4)=9-8=\boxed{1.}
\]

\textbf{(ii)} Since $\alpha$ is a root,
\[
\alpha^3-3\alpha^2+4\alpha-5=0,
\]
so
\[
\alpha^3=3\alpha^2-4\alpha+5.
\]
Similarly,
\[
\beta^3=3\beta^2-4\beta+5,\qquad
\gamma^3=3\gamma^2-4\gamma+5.
\]
Adding,
\[
\alpha^3+\beta^3+\gamma^3
=3(\alpha^2+\beta^2+\gamma^2)-4(\alpha+\beta+\gamma)+15.
\]
Hence
\[
\alpha^3+\beta^3+\gamma^3=3(1)-4(3)+15=\boxed{6.}
\]

\textbf{(iii)} Multiply the original equation by $x$:
\[
x^4-3x^3+4x^2-5x=0.
\]
So for each root,
\[
\alpha^4=3\alpha^3-4\alpha^2+5\alpha,
\]
and similarly for $\beta,\gamma$. Summing,
\[
\alpha^4+\beta^4+\gamma^4
=3(\alpha^3+\beta^3+\gamma^3)-4(\alpha^2+\beta^2+\gamma^2)+5(\alpha+\beta+\gamma).
\]
Therefore
\[
\alpha^4+\beta^4+\gamma^4=3(6)-4(1)+5(3)=18-4+15=\boxed{29.}
\]

\textbf{(iv)} Since each root satisfies
\[
x^3-3x^2+4x-5=0,
\]
we have, for any root $r\in\{\alpha,\beta,\gamma\}$,
\[
r^3=3r^2-4r+5.
\]
Multiply by $r^{k-3}$:
\[
r^k=3r^{k-1}-4r^{k-2}+5r^{k-3}.
\]
Now do this for $r=\alpha,\beta,\gamma$ and add. Then
\[
S_k=3S_{k-1}-4S_{k-2}+5S_{k-3}.
\]
So the required recurrence is
\[
\boxed{S_k=3S_{k-1}-4S_{k-2}+5S_{k-3}\qquad (k\ge 3).}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Vieta gives the starting symmetric sums.
\item A polynomial equation satisfied by each root can be reused to generate higher powers.
\item When summing three root equations, remember to add the constant term three times.
\item The same idea gives a recurrence for all higher power sums $S_k$.
\end{itemize}
\end{takeaways}

\begin{problem}[Sketching a Cubic with Derivatives]
Sketch the graph of
\[
f(x)=x^3-2x^2-3x.
\]

Start by:
\begin{itemize}[leftmargin=*]
\item finding $f'(x)$ and solving $f'(x)=0$,
\item analysing the sign of $f'(x)$ and the turning points,
\item summarising the behaviour in a table,
\item then drawing the graph.
\end{itemize}
\end{problem}

\begin{hintbox}
Factor both $f(x)$ and $f'(x)$. The derivative tells you where the graph is increasing or decreasing.
\end{hintbox}

\begin{solution}
Differentiate:
\[
f'(x)=3x^2-4x-3.
\]
Solve
\[
3x^2-4x-3=0.
\]
Using the quadratic formula,
\[
x=\frac{4\pm\sqrt{16+36}}{6}
=\frac{4\pm2\sqrt{13}}{6}
=\frac{2\pm\sqrt{13}}{3}.
\]

Since $f'(x)$ is an upward-opening quadratic,
\[
f'(x)>0 \text{ for } x<\frac{2-\sqrt{13}}{3}
\quad\text{and}\quad
x>\frac{2+\sqrt{13}}{3},
\]
and
\[
f'(x)<0 \text{ for } \frac{2-\sqrt{13}}{3}<x<\frac{2+\sqrt{13}}{3}.
\]
Therefore:
\begin{itemize}[leftmargin=*]
\item $f$ is increasing, then decreasing, then increasing;
\item there is a local maximum at $x=\dfrac{2-\sqrt{13}}{3}$;
\item there is a local minimum at $x=\dfrac{2+\sqrt{13}}{3}$.
\end{itemize}

Also
\[
f(x)=x(x-3)(x+1),
\]
so the $x$-intercepts are
\[
x=-1,\ 0,\ 3.
\]

The behaviour table is:
\[
\begin{array}{c|ccccc}
x & -\infty & \frac{2-\sqrt{13}}{3} &  & \frac{2+\sqrt{13}}{3} & +\infty\\ \hline
f'(x) & + & 0 & - & 0 & +\\ \hline
f(x) & \nearrow & \text{local max} & \searrow & \text{local min} & \nearrow
\end{array}
\]

A rough sketch is:
\begin{center}
\begin{tikzpicture}
\pgfmathsetmacro{\xone}{(2 - sqrt(13))/3}
\pgfmathsetmacro{\xtwo}{(2 + sqrt(13))/3}
\pgfmathsetmacro{\yone}{(\xone)^3 - 2*(\xone)^2 - 3*(\xone)}
\pgfmathsetmacro{\ytwo}{(\xtwo)^3 - 2*(\xtwo)^2 - 3*(\xtwo)}
\begin{axis}[
  axis lines=middle,
  xmin=-2.2,xmax=4.2,
  ymin=-10,ymax=8,
  samples=400,
  width=0.78\linewidth,
  height=6.2cm,
  xtick={-1,0,3},
  ytick=\empty,
  xlabel={$x$},
  ylabel={$y$}
]
\addplot[thick,bookpurple,domain=-2.2:4.2]{x^3-2*x^2-3*x};
\addplot[dashed,gray] coordinates {(\xone,-10) (\xone,8)};
\addplot[dashed,gray] coordinates {(\xtwo,-10) (\xtwo,8)};
\addplot[dashed,gray] coordinates {(-2.2,\yone) (4.2,\yone)};
\addplot[dashed,gray] coordinates {(-2.2,\ytwo) (4.2,\ytwo)};
\addplot[only marks,mark=*,mark size=2.5pt,bookred] coordinates {(\xone,\yone)};
\addplot[only marks,mark=*,mark size=2.5pt,bookred] coordinates {(\xtwo,\ytwo)};

% \node[
%   font=\small,
%   anchor=south east,
%   fill=white,
%   draw=gray!50,
%   rounded corners=1pt,
%   inner sep=3pt,
%   text width=4.3cm,
%   align=center
% ] at (axis cs:\xone,\yone) {local max at $x\approx -0.53518$};
% \node[
%   font=\small,
%   anchor=north west,
%   fill=white,
%   draw=gray!50,
%   rounded corners=1pt,
%   inner sep=3pt,
%   text width=4.3cm,
%   align=center
% ] at (axis cs:\xtwo,\ytwo) {local min at $x\approx 1.86852$};

\end{axis}
\end{tikzpicture}
\end{center}
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item To sketch a cubic accurately, use both the intercepts and the derivative.
\item The sign of $f'(x)$ tells you where the graph is increasing or decreasing.
\item A local maximum occurs where $f'(x)$ changes from positive to negative; a local minimum occurs where it changes from negative to positive.
\end{itemize}
\end{takeaways}

\begin{problem}[Divisibility by $(x+1)^2$]
Consider the polynomial
\[
P(x)=ax^8+bx^7+1,
\]
where $a$ and $b$ are non-zero real numbers. Find the values of $a$ and $b$ for which $P(x)$ is divisible by $(x+1)^2$.
\end{problem}

\begin{hintbox}
If $(x+1)^2$ divides $P(x)$, then $x=-1$ is a repeated root. So use $P(-1)=0$ and $P'(-1)=0$.
\end{hintbox}

\begin{solution}
For $(x+1)^2$ to divide $P(x)$, we need
\[
P(-1)=0\qquad\text{and}\qquad P'(-1)=0.
\]
Now
\[
P(-1)=a-b+1,
\]
so
\[
a-b+1=0. \tag{1}
\]

Also
\[
P'(x)=8ax^7+7bx^6,
\]
hence
\[
P'(-1)=-8a+7b=0. \tag{2}
\]

From (2),
\[
7b=8a \Rightarrow b=\frac{8a}{7}.
\]
Substitute into (1):
\[
a-\frac{8a}{7}+1=0
\Rightarrow -\frac{a}{7}+1=0
\Rightarrow a=7.
\]
Then
\[
b=\frac{8a}{7}=8.
\]
Therefore
\[
\boxed{a=7,\qquad b=8.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item If $(x-r)^2$ is a factor, then both $P(r)$ and $P'(r)$ must be zero.
\item Repeated-factor questions usually become a pair of simultaneous equations in the unknown coefficients.
\end{itemize}
\end{takeaways}
