% ============================================================
% Part 1 — Basic Worked Problems
% Topics: language of polynomials, simple factor/remainder theorem,
%         straightforward Vieta's, basic graph sketching
% ============================================================

\begin{problem}[Identically Equal Polynomials]
Find $a$, $b$ and $c$ given that
\[
a(x-2)^2+b(x-2)+c=x^2+x+1
\quad\text{for all }x.
\]
\end{problem}

\begin{hintbox}
Expand the left-hand side and compare coefficients.
\end{hintbox}

\begin{solution}
\[
a(x-2)^2+b(x-2)+c=ax^2+(-4a+b)x+(4a-2b+c).
\]
Comparing with $x^2+x+1$ gives
\[
a=1,\qquad -4a+b=1,\qquad 4a-2b+c=1.
\]
Hence
\[
\boxed{a=1,\quad b=5,\quad c=7.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Polynomial identities are solved by coefficient comparison.
\item Expanding around $(x-2)$ is still just an ordinary polynomial identity.
\end{itemize}
\end{takeaways}

\begin{problem}[Odd Polynomial Coefficients]
Suppose
\[
P(x)=ax^5+bx^4+cx^3+dx^2+ex+f
\]
is an odd function, so $P(-x)=-P(x)$ for all $x$. Show that $b=d=f=0$.
\end{problem}

\begin{hintbox}
Write down $P(-x)$ and compare it with $-P(x)$ term by term.
\end{hintbox}

\begin{solution}
\[
P(-x)=-ax^5+bx^4-cx^3+dx^2-ex+f,
\]
while
\[
-P(x)=-ax^5-bx^4-cx^3-dx^2-ex-f.
\]
Since these are identical,
\[
b=-b,\qquad d=-d,\qquad f=-f.
\]
So
\[
\boxed{b=d=f=0.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Odd polynomials contain only odd powers of $x$.
\item The constant term of an odd polynomial must be $0$.
\end{itemize}
\end{takeaways}

\begin{problem}[Central Coefficient]
What is the coefficient of $x^n$ in
\[
G(x)=\left(1+x+x^2+\cdots+x^n\right)^2?
\]
\end{problem}

\begin{hintbox}
Count the pairs $(i,j)$ with $0\le i,j\le n$ and $i+j=n$.
\end{hintbox}

\begin{solution}
Write
\[
G(x)=\left(\sum_{i=0}^{n}x^i\right)\left(\sum_{j=0}^{n}x^j\right).
\]
The term $x^n$ appears whenever $x^i\cdot x^j=x^n$, that is, whenever $i+j=n$.
The possible pairs are
\[
(0,n),(1,n-1),\ldots,(n,0),
\]
which gives $n+1$ choices. Therefore the coefficient is
\[
\boxed{n+1.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Coefficient questions in products often reduce to counting exponent pairs.
\item The coefficient of $x^r$ is the number of ways to split $r$ as a sum of exponents.
\end{itemize}
\end{takeaways}

\begin{problem}[Easy Zeroes First]
Solve
\[
x^3-x^2-4x+4=0
\]
by first factoring the left-hand side.
\end{problem}

\begin{hintbox}
Factor by grouping.
\end{hintbox}

\begin{solution}
\[
x^3-x^2-4x+4=x^2(x-1)-4(x-1)=(x-1)(x^2-4).
\]
So
\[
(x-1)(x-2)(x+2)=0.
\]
Hence
\[
\boxed{x=1,\ 2,\ -2.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Before using heavier tools, try grouping or simple integer roots.
\item Cubics designed for students often hide an obvious factor.
\end{itemize}
\end{takeaways}

\begin{problem}[Factors of $x^n+1$]
Is either $x+1$ or $x-1$ a factor of $x^n+1$, where $n$ is a positive integer?
\end{problem}

\begin{hintbox}
Use the factor theorem at $x=1$ and $x=-1$.
\end{hintbox}

\begin{solution}
Let $P(x)=x^n+1$.

For $x-1$ to be a factor we need $P(1)=0$, but
\[
P(1)=2\ne 0.
\]
So $x-1$ is never a factor.

For $x+1$ to be a factor we need $P(-1)=0$:
\[
P(-1)=(-1)^n+1.
\]
If $n$ is odd, this is $0$. If $n$ is even, this is $2$.

Therefore
\[
\boxed{x+1\text{ is a factor iff }n\text{ is odd, and }x-1\text{ is never a factor}.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item The factor theorem turns factor questions into substitutions.
\item Powers of $-1$ almost always split into odd and even cases.
\end{itemize}
\end{takeaways}

\begin{problem}[Simple Partial Fractions]
Given
\[
\frac{1}{x^3-1}=\frac{A}{x-1}+\frac{Bx+C}{x^2+x+1},
\]
find $A$, $B$ and $C$.
\end{problem}

\begin{hintbox}
Use $x^3-1=(x-1)(x^2+x+1)$, then clear denominators.
\end{hintbox}

\begin{solution}
Since
\[
x^3-1=(x-1)(x^2+x+1),
\]
we have
\[
1=A(x^2+x+1)+(Bx+C)(x-1).
\]
Expand:
\[
1=(A+B)x^2+(A-B+C)x+(A-C).
\]
Comparing coefficients gives
\[
A+B=0,\qquad A-B+C=0,\qquad A-C=1.
\]
Solving,
\[
\boxed{A=\frac13,\qquad B=-\frac13,\qquad C=-\frac23.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Clear denominators before comparing coefficients.
\item Partial fractions are another good place to use polynomial identities.
\end{itemize}
\end{takeaways}

\begin{problem}[Partial Fractions with a Repeated Linear Factor]
Let
\[
P(x)=x^2(x-1)(x+1).
\]
Find constants $A,B,C,D$ such that
\[
\frac{1}{P(x)}=\frac{A}{x}+\frac{B}{x^2}+\frac{C}{x-1}+\frac{D}{x+1}
\]
for all $x$ for which the expression is defined.
\end{problem}

\begin{hintbox}
Multiply through by $x^2(x-1)(x+1)$, then use convenient values such as $x=0,1,-1$ before comparing coefficients.
\end{hintbox}

\begin{solution}
We want
\[
\frac{1}{x^2(x-1)(x+1)}=\frac{A}{x}+\frac{B}{x^2}+\frac{C}{x-1}+\frac{D}{x+1}.
\]
Multiply by $x^2(x-1)(x+1)$:
\[
1=Ax(x-1)(x+1)+B(x-1)(x+1)+Cx^2(x+1)+Dx^2(x-1).
\]
Now substitute convenient values:
\[
x=0 \Rightarrow 1=-B \Rightarrow B=-1,
\]
\[
x=1 \Rightarrow 1=2C \Rightarrow C=\frac12,
\]
\[
x=-1 \Rightarrow 1=2D \Rightarrow D=\frac12.
\]
Use $x=2$:
\[
1=6A+3B+12C+4D.
\]
Substituting $B=-1$, $C=D=\frac12$ gives
\[
1=6A-3+6+2=6A+5,
\]
so
\[
A=-\frac23.
\]
Hence
\[
\boxed{A=-\frac23,\quad B=-1,\quad C=\frac12,\quad D=\frac12.}
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item A repeated linear factor $(x-a)^2$ contributes two terms: $\dfrac{A}{x-a}+\dfrac{B}{(x-a)^2}$.
\item Substituting the roots of the denominator is usually faster than comparing every coefficient immediately.
\end{itemize}
\end{takeaways}

\begin{problem}[Partial Fractions with an Irreducible Quadratic]
Let
\[
Q(x)=(x^2+1)(x-1)^2.
\]
Find constants $A,B,C,D$ such that
\[
\frac{1}{Q(x)}=\frac{Ax+B}{x^2+1}+\frac{C}{x-1}+\frac{D}{(x-1)^2}
\]
for all $x$ for which the expression is defined.
\end{problem}

\begin{hintbox}
Clear denominators first. Use $x=1$, then compare coefficients or substitute more values.
\end{hintbox}

\begin{solution}
We require
\[
\frac{1}{(x^2+1)(x-1)^2}=\frac{Ax+B}{x^2+1}+\frac{C}{x-1}+\frac{D}{(x-1)^2}.
\]
Multiply by $(x^2+1)(x-1)^2$:
\[
1=(Ax+B)(x-1)^2+C(x^2+1)(x-1)+D(x^2+1).
\]
Expand:
\[
1=(Ax+B)(x^2-2x+1)+C(x^3-x^2+x-1)+D(x^2+1).
\]
So
\[
\begin{aligned}
1={}&(A+C)x^3+(-2A+B-C+D)x^2\\
&+(A-2B+C)x+(B-C+D).
\end{aligned}
\]
Compare coefficients:
\[
\begin{cases}
A+C=0,\\
-2A+B-C+D=0,\\
A-2B+C=0,\\
B-C+D=1.
\end{cases}
\]
From $A+C=0$, we get $C=-A$. Then
\[
\;A-2B+C=0 \Rightarrow -2B=0 \Rightarrow B=0.
\]
\[
\;-2A+B-C+D=0 \Rightarrow -A+D=0 \Rightarrow D=A.
\]
Finally,
\[
\;B-C+D=1 \Rightarrow A+A=1 \Rightarrow A=\frac12.
\]
So
\[
\boxed{A=\frac12,\quad B=0,\quad C=-\frac12,\quad D=\frac12.}
\]
\[
\frac{1}{(x^2+1)(x-1)^2}
=\frac{x}{2(x^2+1)}-\frac{1}{2(x-1)}+\frac{1}{2(x-1)^2}.
\]
\end{solution}

\begin{takeaways}
\begin{itemize}[leftmargin=*]
\item Integrals of the form $\int \dfrac{1}{(x^2+1)(x-1)^2}\,dx$ can be evaluated using partial fractions.
\item An irreducible quadratic factor $(ax^2+bx+c)$ contributes a numerator of the form $Mx+N$ in the fully general case.
\item General rule: if
\[
P(x)=\prod (L_i(x))^{r_i}\prod (Q_j(x))^{s_j},
\]
where each $L_i$ is linear and each $Q_j$ is an irreducible quadratic, then
\[
\frac{1}{P(x)}
\]
is expanded using
\[
\frac{A_1}{L_i}+\frac{A_2}{L_i^2}+\cdots+\frac{A_{r_i}}{L_i^{r_i}}
\]
for each repeated linear factor, and
\[
\frac{M_1x+N_1}{Q_j}+\frac{M_2x+N_2}{Q_j^2}+\cdots+\frac{M_{s_j}x+N_{s_j}}{Q_j^{s_j}}
\]
for each repeated irreducible quadratic factor.
\end{itemize}
\end{takeaways}
