\documentclass[12pt]{article}

% Core math packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{enumitem}

% Custom styles shared with HSC-Collections
\usepackage{styles/dl101-colors}
\usepackage{styles/dl101-hints}
\usepackage{styles/dl101-hsc-problems}

\geometry{
  a4paper,
  total={170mm,257mm},
  left=20mm,
  top=20mm,
}

\author{Vu Hung Nguyen}
\title{HSC Math Extension 2: Induction Mastery}
\date{}

\begin{document}

\maketitle

\begingroup\small\noindent This work is licensed under CC BY 4.0, see the LICENSE file on the github page for more info.\par\endgroup

\section{Introduction}

\subsection{Project Overview}
This booklet compiles high quality mathematical induction problems curated specifically for the HSC Mathematics Extension~2 syllabus. Every task can be attacked with induction (sometimes combined with algebraic manipulation, inequalities, recurrence relations, or integrals). Detailed reasoning showcases common techniques such as weak and strong induction, structural induction on algebraic forms, and inductive proofs for inequalities.

\subsection{Target Audience}
The explanations are crafted for Extension~2 students aiming to deepen their proof-writing skills. Each solution explicitly states the induction hypothesis, justification of the base case, and reasoning for the inductive step so that high-school learners can follow every transition.

\subsection{How to Use This Booklet}
\begin{itemize}[leftmargin=*]
  \item Read the overview and induction primer before attempting the problems.
  \item Attempt problems in Part~1 without hints; compare against the detailed solutions to understand model reasoning.
  \item Use the upside-down hints in Part~2 only after making a genuine attempt.
  \item Revisit problems after a few days and try to re-derive the arguments without notes to reinforce technique.
\end{itemize}

\subsection{Induction Primer}
Mathematical induction is a two-step method:
\begin{enumerate}[leftmargin=*]
  \item \textbf{Base Case:} Verify the statement for the initial value (often $n=1$ or $n=0$), ensuring the proposition is anchored.
  \item \textbf{Inductive Step:} Assume the statement is true for $n=k$ (or all values up to $k$ for strong induction) and show it holds for $n=k+1$. Clearly state the hypothesis and highlight algebraic transformations used to bridge from $k$ to $k+1$.
\end{enumerate}
We will reference these two steps in every solution so students can trace the logic effortlessly.

\section{Part 1: Problems and Solutions (Detailed)}
Part~1 contains three sets of problems---basic, medium, and advanced. Each set provides five problems. For every problem we present a comprehensive induction-based solution without any hints so that learners focus on the full reasoning trail.

\subsection{Basic Induction Problems}
\input{problems/part1-basic}
\input{problems/13}

\subsection{Medium Induction Problems}
\input{problems/part1-medium}
% additionally include the new weighted geometric-sum problem
\input{problems/06}
% additionally include newly added problems 07 and 08
\input{problems/07}
\input{problems/08}
% additionally include problems 09 and 10
\input{problems/09}
\input{problems/10}
% additionally include problems 11 and 12
\input{problems/11}
\input{problems/12}

\subsection{Advanced Induction Problems}
\input{problems/part1-advanced}

\section{Part 2: Problems and Solutions (Concise + Hints)}
Part~2 revisits three new sets of problems. Solutions are intentionally briefer to encourage student ownership, and every problem includes an upside-down hint.

\subsection{Basic Induction Problems}
\input{problems/part2-basic}
% New problem added: logarithmic derivative identity (induction)
% Difficulty: Basic (easy) — included in Part 2 Basic problems
\input{problems/14}

\subsection{Medium Induction Problems}
\input{problems/part2-medium}

\subsection{Advanced Induction Problems}
\input{problems/part2-advanced}

\section{Conclusion}
Induction is a core reasoning tool in the HSC Mathematics Extension~2 course. Mastery comes from repeated, reflective practice. Use these problems to sharpen the ability to hypothesize, algebraically manipulate, and communicate complete proofs. Best of luck with your studies and future competitions!

\vspace{2em}
\noindent\textbf{Contact Information:}\\[0.5em]
LinkedIn: \href{https://www.linkedin.com/in/nguyenvuhung/}{https://www.linkedin.com/in/nguyenvuhung/}\\[0.3em]
GitHub: \href{https://github.com/vuhung16au/}{https://github.com/vuhung16au/}\\[0.3em]
Repository: \href{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Induction}{https://github.com/vuhung16au/math-olympiad-ml/tree/main/HSC-Induction}

\clearpage
\section{Cross-Links to Other Booklets}

\begingroup\small
\begin{itemize}
  \item \textbf{Last Resorts (Problem 16)}\\
  \href{https://bit.ly/HSC-ext2-LastResort}{https://bit.ly/HSC-ext2-LastResort}\\
  This project curates the most challenging problems from HSC Mathematics Extension 2 examinations---the notorious Problem 16, known as "Last Resorts." Every problem represents the pinnacle of high school mathematics, combining multiple advanced topics and requiring sophisticated problem-solving techniques.

  \item \textbf{Collections (Hard Problems)}\\
  \href{https://bit.ly/HSC-ext2-Collections}{https://bit.ly/HSC-ext2-Collections}\\
  This repository contains a curated collection of challenging problems from the HSC Mathematics Extension 2 curriculum, designed to test deep understanding, creative problem-solving skills, and the ability to synthesize multiple mathematical concepts.

  \item \textbf{Complex Numbers}\\
  \href{https://bit.ly/HSC-ext2-ComplexNumbers}{https://bit.ly/HSC-ext2-ComplexNumbers}\\
  This project curates complex numbers problems tailored to HSC Mathematics Extension 2 students. Every problem is solved with comprehensive explanations written so motivated high-school learners can follow each step.

  \item \textbf{Mathematical Induction}\\
  \href{https://bit.ly/HSC-ext2-Induction}{https://bit.ly/HSC-ext2-Induction}\\
  This project curates mathematical induction problems tailored to HSC Mathematics Extension 2 students. Every problem is solved with induction and written so motivated high-school learners can follow each step.

  \item \textbf{Inequalities}\\
  \href{https://bit.ly/HSC-ext2-Inequalities}{https://bit.ly/HSC-ext2-Inequalities}\\
  This project curates inequality problems tailored to HSC Mathematics Extension 2 students. Each problem explores fundamental inequality techniques including AM-GM, Cauchy-Schwarz, triangle inequality, integration-based inequalities, and inequalities via mathematical induction.

  \item \textbf{Integrals}\\
  \href{https://bit.ly/HSC-ext2-Integrals}{https://bit.ly/HSC-ext2-Integrals}\\
  This project curates integration problems tailored to HSC Mathematics Extension 2 students. Every problem demonstrates essential integration techniques with solutions written so motivated high-school learners can follow each step.

  \item \textbf{Mechanics}\\
  \href{https://bit.ly/hsc-ext2-Mechanics}{https://bit.ly/hsc-ext2-Mechanics}\\
  This project presents a comprehensive collection of mechanics problems for HSC Mathematics Extension 2 students. Every problem demonstrates key mathematical techniques including integration methods, differential equations, Newton's laws applications, and limiting behavior analysis.

  \item \textbf{Polynomials}\\
  \href{https://bit.ly/hsc-ext2-Polynomials}{https://bit.ly/hsc-ext2-Polynomials}\\
  This project curates polynomial problems tailored to HSC Mathematics Extension 2 students. Every problem covers essential polynomial techniques including complex numbers, roots of unity, Vieta's formulas, De Moivre's theorem, and transformations of roots.

  \item \textbf{Proofs}\\
  \href{https://bit.ly/hsc-ext2-proofs}{https://bit.ly/hsc-ext2-proofs}\\
  This project presents a comprehensive collection of mathematical proof problems for HSC Mathematics Extension 2 students. Every problem demonstrates key proof techniques including direct proof, proof by contradiction, mathematical induction, and proof.

  \item \textbf{Vectors}\\
  \href{https://bit.ly/hsc-ext2-vectors}{https://bit.ly/hsc-ext2-vectors}\\
  This project curates high-quality vector problems tailored to HSC Mathematics Extension 2 students. Every problem covers essential 3D vector concepts with solutions that demonstrate both algebraic rigor and geometric intuition.
\end{itemize}
\endgroup

\end{document}
