\subsection{Differential equations and solutions}
A differential equation links an unknown function to its derivatives. Here we focus on first-order equations (highest derivative is $y'$).

\begin{itemize}[leftmargin=*]
  \item A \textbf{general solution} contains an arbitrary constant such as $C$.
  \item A \textbf{particular solution} is found after using a condition such as $y(0)=3$.
  \item An \textbf{initial condition} picks one curve from the family.
\end{itemize}

\textbf{Example:} $\frac{dy}{dx}=2x \Rightarrow y=x^2+C$, and $y(0)=3 \Rightarrow y=x^2+3$.

\subsection{Direction fields and equilibrium solutions}
For $\frac{dy}{dx}=F(x,y)$, a slope field draws a short segment at each point $(x,y)$ with slope $F(x,y)$.

\begin{itemize}[leftmargin=*]
  \item If the slope is positive, solution curves rise from left to right.
  \item If the slope is negative, solution curves fall from left to right.
  \item If the slope is zero along $y=k$, then $y=k$ is an equilibrium solution.
\end{itemize}

\textbf{Example:} $\frac{dy}{dx}=y-2$ has equilibrium $y=2$.

\subsection{Separable structure}
Common separable forms are:
\[
\frac{dy}{dx} = f(x), \qquad
\frac{dy}{dx} = g(y), \qquad
\frac{dy}{dx} = f(x)g(y).
\]

If variables can be separated, use this workflow:
\begin{enumerate}[leftmargin=*]
  \item Rearrange so all $y$ terms are with $dy$ and all $x$ terms are with $dx$.
  \item Integrate both sides.
  \item Apply any initial condition.
  \item Interpret the result in context.
\end{enumerate}

\textbf{Example:} $\frac{dy}{dx}=xy \Rightarrow \frac{1}{y}dy=x\,dx$.

\subsection{Logistic-style models}
The logistic equation is
\[
\frac{dy}{dx} = ky(M-y),
\]
equivalently
\[
\frac{dy}{dx} = ry\left(1-\frac{y}{M}\right),
\]
where $M$ is the carrying capacity.

\begin{itemize}[leftmargin=*]
  \item $y=0$ and $y=M$ are equilibrium values.
  \item If $0<y<M$, growth is positive.
  \item If $y>M$, the model predicts decay back toward the carrying capacity.
\end{itemize}

\textbf{Example:} If $M=100$ and $y=20$, then $\left(1-\frac{y}{M}\right)=0.8$ so growth is positive.
